### 2024-07-05 Manuscript_draft/Draft.Rmd
First version draft (just the core skeleton)

---
title: "Draft"
author: "Lorenzo Menichetti"
date: "2024-07-04"
output: html_document
  # pdf_document:
  #   citation_package: natbib
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Abstract

# Introduction

# Material and methods

## Study sites
The study sites are located in Gamiz, Northern Spain (42°49'2.32"N; 2°37'10.67"W), Saint-Christol-d’Albion, South of France (xxx) and Dumbravita, East Romania (42°49'2.32"N; 2°37'10.67"W). According to Köppen-Geiger climatic classification, the climate in Gamiz is Oceanic, with cold and rainy winters and warm summers. In St. Christol the climate is Mediterranean, with mild winters and warm and dry summers. In Dumbravita, the climate is humid continental with warm to hot summers, and cold and snowy winters with precipitation usually distributed throughout the year (Table 1). In the three study sites, the dominant vegetation are deciduous species of Quercus. Specifically, Quercus faginea Lam. in Gamiz, Quercus pubescens Willd. in St. Christol and Quercus robur L. in Dumbravita (Table 1). In St.Christol and Dumbravita the stands are Quercus monodominant, while in Gamiz, Quercus faginea appears together with scattered Acer campestris L. and Fraxinus excelsior L., and an understory dominated by Crataegus monogyna Jacq., Juniperus communis L. and Viburnum lantana L. More characteristics of the study sites are found in Table 1.
The experimental sites have an approximate total area of 1.5 ha. In each site, three stands of 0.12 ha each and replicated in four blocks were delimited. The stands corresponded to a control and two canopy disturbance treatments, i.e., 50% thinning and clear-cut (Fig. S1). The stands from disturbed treatments (thinning and clear-cut) were in turn divided in two: on half of the stands the slash resulting from the logging (leaves and twigs) was crashed and added on top of the soil and on the other half, the woody debris resulting from logging was removed. This made a total of 5 treatments (control, clear-cut with and without slash, thinning with and without slash) per block (Fig. S1). In each of the 5 treatments, two plots were established per block making a total of 40 plots (4 blocks x 5 treatments x 2 replicates per block; Fig S1). In turn, in each plot a transect of 2m length was established for monitoring (Fig. S1). Both the canopy disturbance and the slash addition treatments together with the fencing of the entire area were carried out in October 2021 in Gamiz, April 2022 in St. Christol and March 2022 in Dumbravita.

## Soil respiration 
	To measure soil CO2 efflux (i.e., soil respiration), three rings of 10 cm diameter and 8 cm high were put along the 2 m transect in each of the monitoring plots (Fig. S1). Respiration was measured by a mobile infrared gas analyser (EGM-5, PP Systems, Hitchin, UK) and an attached soil respiration chamber (SRC-1). Soil CO2 efflux (g CO2 m-2 h-1) was calculated as linear increase in CO2 concentrations within the chamber headspace. The total volume used for calculation was the volume of the chamber plus the volume of the specific ring. Measurements were done monthly in the 120 collars in each site. Measurement in Gamiz started in November 2021, in St. Christol in February 2023 and in Dumbravita in May 2022.
Forest stand characteristics
Within each of the control and thinned stands (n=8), we recorded the number of trees to calculate tree density (No. tree per ha) and measured tree diameter at breast height (DBH; 1.3 m) to calculate basal area.

## Microclimatic variables
We measured soil moisture and temperature using TMS data loggers version 4 (Wild et al., 2019), located at the extreme of the 2 m transect in each of the monitoring plots (Fig S1). These probes have three temperature sensors positioned at +15 cm, 2 cm and -6 cm relative to the soil surface, and a sensor that measures volumetric soil moisture based on time domain transmission to a depth of 6 cm. To convert the raw TMS measurements of soil moisture to volumetric soil water content, we used a calibration curve developed by Kopecký et al. (2021). For further modelling and since these sensors measure every 15 min, we extracted the soil temperature and soil moisture measurements closer to the time we measured soil respiration in the field.
Soil physical-chemical analyses
Chemical analyses were measured on 5 soil cores of 4 cm diameter separated 0.5 m to a depth of 10 cm along the 2m monitoring transect. The five soil cores were mixed into a composite sample per plot. The composite sample was sieved through a 5-mm mesh sieve in the field, stored at −20°C and then freeze-dried prior to laboratory analyses. The pH was measured in distilled water 1:10 (w/v). The carbon (C) and nitrogen (N) content of the soil was measured using an elemental analyser: C content was measured by sulfochromic oxidation (ISO 14235), and N content was estimated by sulfuric acid mineralization with the addition of selenium and sodium sulphate and conversion to ammonium ions (ISO 11261), which were measured with the Segmented Flow Analyzer (SFA) Skalar. Available phosphorus (P) was determined after extraction with malachite green according to the protocol by Ohno & Zibilske (1991). Bulk density was measured in each of the plots by collecting 3 soil cores of 6 cm diameter and 10 cm depth around the 2m transect that were mixed into a composite sample. Soil bulk density was determined by dividing the dry mass of the composite sample by the volume of the 3 cores, after removing the stones. Soil C stocks (g m2) were calculated by multiplying C concentration by bulk density and sampling depth.
Microbial biomass
The same five soil cores used for chemical analyses were used to estimate soil microbial biomass. To do that, we extracted Phospholipid fatty acids (PLFAs) following Šnajdr et al. (2008). Lipids were extracted using approximately 1g of dry soil with a mixture of chloroform:methanol:phosphate buffer (1:2:0.8) and separated by solid-phase extraction cartridges (HyperSep Silica SPE columns 200 mg/3 mL, Thermo Fisher Scientific). PLFAs were eluted using 2 mL of methanol. The selected fractions underwent to mild alkaline methanolysis and the methyl esters of PLFAs were analyzed by gas chromatography-mass spectrometry (GC-MS; 450-GC, 240-MS ion trap detector, Varian, Walnut Creek, CA). In the polar PLFA fraction, fungal biomass was quantified based on 18:2ω6,9 (fungal PLFA) content; bacterial biomass (total bacterial PLFA) was quantified as a sum of i14:0, i15:0, a15:0, 16:1ω5, 16:1ω7, 16:1ω9, 10Me-16:0, i16:0, i17:0, a17:0, cy17:0, 17:0, 10Me-17:0, 18:1ω7, 10Me-18:0, cy19:0 (Actinomycetes 10Me-16:0, 10Me-17:0, 10Me-18:0, Gram-positive i14:0, i15:0, a15:0, i16:0, i17:0, a17:0 and Gram-negative 16:1ω7, 16:1ω9, 18:1ω7, cy17:0, cy19:0). The sum of all of the selected PLFAs together with 16:0 and 18:1w9 was used as an estimate of total microbial biomass (total PLFA).

## Data analysis

## Validation
In order to ensure a robust validation, all models have been validated with a split approach, keeping two completely independent data sets (80% of the points for training and 20% validation).

## Parametric model
We defined a parametric model to consider all the three effects on respiration we hypothesized. The first is the influence of temperature, then the influence of soil moisture, and then a seasonal component which is assumed to be roughly proportional to the interactions of autotrophic respiration (which we hypothesized following a seasonal pattern).
Temperature and moisture scaling interact linearly, while seasonality was assumed additive:
$$
Rh(t) =  \xi_{temp(t,j)} \cdot \xi_{moist(t,j)} \right) + \xi_{sin(t,j)}
$$

Where the index $j$ corresponds to the treatments and the index $k$ corresponds to the plots (minimal statistical unit), while $(t)$ denotes the variation over time of each forcing variable (temperature, moisture, or days of the year)

## Model calibration
The parametric model was calibrated in a stratified Bayesian framework, where we assigned independent parameters depending on different groping of the data.  
The model was written in Stan [@carpenter2017stan; @standev2018stancore], run from its R interface [@Rstan]

### Moisture 
The moisture scaling function had to be taken from a simplified version of $Moyano.2013 in the SoilR package [@SierraMueller].
$$
\xi_{moist(t,k)}  =   a_j \cdot \theta_{v}(t) - b_j \cdot (\theta_{v}(t))^2
$$
Variance is by treatment.

### Temperature
The temperature scaling function assumes an independent activation energy $E_a$ varying by treatments $j$:
$$
\xi_{temp(t,j)} = A_{k} \cdot e^{\frac{-Ea_j}{(T(t)+273.15)-T_0)}}
$$
The two parameters $E_a$ and $T_0$ are represented by Bayesian priors.

### Seasonality
The seasonality scaling functions assume a sine function with period = 365 days, with both amplitude $ampl_j$ (the magnitude) and peak day $peak_j$ (when the function is at its only annual peak) varying as a function of treatments:
$$
\begin{aligned}
    \xi_{sin(t,k)} = ampl_k \cdot sin \left(\frac{2 \cdot \pi}{365}\right) \cdot day(t) \\
    +\left(\frac{2 \cdot \pi}{365}\right) \cdot (peak_k - 1) - \frac{\pi}{2})
\end{aligned}
$$
Both the terms $ampl_j$ and $peak_j$ are described by Bayesian priors. Variance is by treatment.

## Model benchmark

## Model extrapolation
Once we calibrated our parametric model, we could test the consequences of an increase in 5&deg;C

## Machine learning variance decomposition
On a more limited subset (only Spain), we tested

# Results

## Sites microclimate

## Benchmark models and parametric model fit

![Alt text](../Figures/fit_ML_validation.png)

![Alt text](../Figures/fit_model_validation.png)

## Parametric model posteriors

![Alt text](../Figures/posteriors_boxplots.png)

![Alt text](../Figures/posteriors_boxplots_A.png)

## Machine learning variance decomposition

![Alt text](../Figures/variance_decomposition.png)

# Conclusions

# Appendix

![Alt text](../Figures/posteriors_temp.png)

![Alt text](../Figures/posteriors_seasonality.png)

![Alt text](../Figures/posteriors_moisture.png)

## Model extrapolations

![Alt text](../Figures/scenario_extrapolation.png)

# References
\bibliography{references}

### 2024-07-05 bayesian_correction_MoySOilR.R

@@ -1,0 +1,717 @@
+
+
+library(rstan)
+library(paletteer)
+library(prettyGraphs)
+library(randomForest)
+library(dplyr)
+library(lubridate)
+library(readxl)
+library(caret)
+
+#TODO microclimatic plots
+#TODO is there an interaction between texture and the TMS calibration?
+#TODO distance between control and treatments
+#TODO plot the treatments for each A
+#TODO unit of A
+#TODO Q10?
+#TODO extrapolation to future?
+
+palette_treat <-  c(paletteer::paletteer_c("ggthemes::Orange-Gold", n=5),
+                    paletteer::paletteer_c("ggthemes::Green", n=5),
+                    paletteer::paletteer_c("ggthemes::Blue-Teal", n=5))
+
+palette_plot <-  c(paletteer::paletteer_c("ggthemes::Orange-Gold", n=length(unique(data[data$site=="france",]$plot_ID))),
+                   paletteer::paletteer_c("ggthemes::Green", n=length(unique(data[data$site=="romania",]$plot_ID))),
+                   paletteer::paletteer_c("ggthemes::Blue-Teal", n=length(unique(data[data$site=="spain",]$plot_ID))))
+
+
+
+edaphics <- read.csv("edaphics.csv")
+#data <- read.csv("Database_co2_28_05_2024.csv")
+data <- read_xlsx("Database_co2_28_05_2024.xlsx", sheet=1)
+data$treatment <- as.factor(data$treatment)
+data$site <- as.factor(data$site)
+data$Soil_moist <- as.numeric(data$Soil_moist)/100
+data$T1_soil <- as.numeric(data$T1_soil)
+data$CO2_flux_hour <- as.numeric(data$CO2_flux_hour)
+data$seq <- 1:dim(data)[1]
+
+
+data$Date <- as.Date(data$Date, format = "%d.%m.%Y")
+
+
+
+processed_data <- data.frame(ID = data$seq,
+                             ID_old = data$ID,
+                             CO2_flux_hour = data$CO2_flux_hour, T1_soil = data$T1_soil,
+                             Soil_moist = data$Soil_moist,
+                             treatment = interaction(data$treatment,as.factor(data$site)),
+                             site = data$site,
[…]

### 2024-07-05 temp_moist_season_model.R

@@ -1,0 +1,22 @@
+
+temp_moist_season <- function(temp, M, day_year, a, b, Ea, A, peak_day, amplitude, treatment, plot_id){
+
+  N = length(temp)
+  xi_moist <- c()
+  xi_temp <- c()
+  sine_wave <- c()
+  model_resp <- c()
+
+  T_0 = 227.13
+
+  for (i in 1:N) {
+    xi_moist[i] =   a[treatment[i]] * M[i] - b[treatment[i]] * M[i]^2;
+    xi_temp[i] = A[plot_id[i]] * (exp(-Ea[treatment[i]] / ((temp[i] + 273.15) - T_0)));
+    sine_wave[i] = amplitude[treatment[i]] * cos((2 * pi / 365) * day_year[i] + (2 * pi / 365) * (peak_day[treatment[i]] - 1) - pi / 2);
+    model_resp[i] = sine_wave[i] +  (xi_temp[i] * xi_moist[i]);
+  }
+
+return(model_resp)
+
+}
+

### 2024-07-05 bayesian_correction_MoySOilR.R
Added climatic variables plot

@@ -15,7 +15,6 @@ library(caret)
 #TODO plot the treatments for each A
 #TODO unit of A
 #TODO Q10?
-#TODO extrapolation to future?
 
 palette_treat <-  c(paletteer::paletteer_c("ggthemes::Orange-Gold", n=5),
                     paletteer::paletteer_c("ggthemes::Green", n=5),
@@ -67,6 +66,65 @@ dim(processed_data)
 
 dropped_levels<-levels(processed_data_filtered_preprocess$plot_id)[!levels(processed_data_filtered_preprocess$plot_id) %in% levels(droplevels(processed_data_filtered_preprocess$plot_id))]
 processed_data_filtered_preprocess$plot_id <- droplevels(processed_data_filtered_preprocess$plot_id)
+
+
+
+
+########### site micrometeorology
+
+png("./Figures/climatedata.png", height=2500, width = 2000, res=300)
+
+data = data.frame(Date = processed_data_filtered_preprocess$date,
+                  Site= processed_data_filtered_preprocess$site,
+                  Temperature = processed_data_filtered_preprocess$T1_soil,
+                  Moisture = processed_data_filtered_preprocess$Soil_moist * 100)
+
+sites <- levels(data_long$Site)
+capitalized_sites <- paste0(toupper(substr(sites, 1, 1)), substr(sites, 2, nchar(sites)))
+data$DateNumeric <- as.numeric(data$Date)
+extended_date_range <- seq(min(data$DateNumeric) - 30, max(data$DateNumeric) + 30, length.out = 300)
+
+par(mfrow=c(3,2), mar=c(2,5,1,1))
+for(i in 1:length(sites)){
+subset<-data[data$Site== sites[i],]
+# Remove NAs
+data_clean <- subset %>% filter(!is.na(Temperature))
+data_clean$DateNumeric <- as.numeric(data_clean$Date)
+
+# Calculate the 2D kernel density estimate
+dens_temp <- kde2d(data_clean$DateNumeric, data_clean$Temperature, n = 300, lims = c(range(extended_date_range), c(0,30)))
+dens_moist <- kde2d(data_clean$DateNumeric, data_clean$Moisture, n = 300, lims = c(range(extended_date_range), c(0,60)))
+
+custom_palette_temp <- colorRampPalette(c("white", "goldenrod1", "orangered", "red4"))(100)
+custom_palette_moist <- colorRampPalette(c("white", "cadetblue2", "dodgerblue", "darkblue"))(100)
+
[…]

### 2024-07-08 Manuscript_draft/Draft.Rmd
Added climatic variables plot
Minor update to the draft

@@
 $$
-Rh(t) =  \xi_{temp(t,j)} \cdot \xi_{moist(t,j)} \right) + \xi_{sin(t,j)}
+Rh(t) =  \xi_{temp(t,j)} \cdot \xi_{moist(t,j)} + \xi_{sin(t,j)}
 $$
@@
 ### Moisture 
-The moisture scaling function had to be taken from a simplified version of $Moyano.2013 in the SoilR package [@SierraMueller].
+The moisture scaling function had to be taken from a simplified version of Moyano (2013) in the SoilR package [@SierraMueller].
 $$
@@
 ## Sites microclimate
+![Climatic parameters measured at the sites, plotted as data density of all measurements. Not all times series cover the same period of time.](../Figures/climatedata.png){width=60%}
 
@@
 
-![Alt text](../Figures/fit_ML_validation.png)
+![Alt text](../Figures/fit_ML_validation.png){width=80%}
 
-![Alt text](../Figures/fit_model_validation.png)
+![Alt text](../Figures/fit_model_validation.png){width=60%}
 
@@
 
-![Alt text](../Figures/posteriors_boxplots.png)
+![Alt text](../Figures/posteriors_boxplots.png){width=60%}
 
-![Alt text](../Figures/posteriors_boxplots_A.png)
+![Alt text](../Figures/posteriors_boxplots_A.png){width=60%}
 
@@
 
-![Alt text](../Figures/variance_decomposition.png)
+![Alt text](../Figures/variance_decomposition.png){width=60%}
+
+## Model extrapolations
+
+![Alt text](../Figures/scenario_extrapolation.png){width=60%}
 
 # Conclusions
+
+# References
+\bibliography{references}
 
@@
 ![Alt text](../Figures/posteriors_moisture.png)
-
-## Model extrapolations
-
-![Alt text](../Figures/scenario_extrapolation.png)
-
-# References
-\bibliography{references}

### 2024-08-01 Manuscript_draft/Draft.Rmd
Updated the draft, switched to pdf

--- 2024-07-05 Manuscript_draft/Draft.Rmd
@@
 date: "2024-07-04"
-output: html_document
-  # pdf_document:
+output: pdf_document
   #   citation_package: natbib
 bibliography: references.bib
+header-includes:
+   - \usepackage[utf8]{inputenc}
+   - \usepackage{textcomp}
+   - \usepackage{amsmath}
+   - \DeclareUnicodeCharacter{03C9}{$\omega$}  # Declare the Unicode character for omega
 ---
@@
 The five soil cores were mixed into a composite sample per plot.
-The composite sample was sieved through a 5-mm mesh sieve in the field, stored at −20°C and then freeze-dried prior to laboratory analyses.
+The composite sample was sieved through a 5-mm mesh sieve in the field, stored at -20°C and then freeze-dried prior to laboratory analyses.
 The pH was measured in distilled water 1:10 (w/v).
@@
 
-## Data analysis
+## Data pretreatment
 
@@
 In order to ensure a robust validation, all models have been validated with a split approach, keeping two completely independent data sets (80% of the points for training and 20% validation).
+Given the size of the dataset, we considered this approach more reliable than any crossvalidation approach.
 
@@
 $$
-Rh(t) =  \xi_{temp(t,j)} \cdot \xi_{moist(t,j)} \right) + \xi_{sin(t,j)}
+Rh(t) =  \xi_{temp(t,j, k)} \cdot \xi_{moist(t,j)} + \xi_{sin(t,j)}
 $$
 
-Where the index $j$ corresponds to the treatments and the index $k$ corresponds to the plots (minimal statistical unit), while $(t)$ denotes the variation over time of each forcing variable (temperature, moisture, or days of the year)
+Where the index $j$ corresponds to the treatments and the index $k$ corresponds to the plots (minimal statistical unit), while $(t)$ denotes the variation over time of each forcing variable (temperature, moisture, or days of the year).
 
@@
 The parametric model was calibrated in a stratified Bayesian framework, where we assigned independent parameters depending on different groping of the data.   
-The model was written in Stan [@carpenter2017stan; @standev2018stancore], run from its R interface [@Rstan]
+The model was written in Stan [@carpenter2017stan; @standev2018stancore], run from its R interface [@Rstan].
+The likelihood function was defined as:  
+$$
+    Rh_{meas} \sim \mathcal{N}(Rh_{sim}, \sigma)
+$$
+Where the parameter $\sigma$ was defined as the variance of the whole dataset (0.27).
 
 ### Moisture 
-The moisture scaling function had to be taken from a simplified version of $Moyano.2013 in the SoilR package [@SierraMueller].
+The moisture scaling function was taken from a simplified version of Moyano (2013) in the SoilR package [@SierraMueller].
 $$
@@
 $$
-Variance is by treatment.
+We assumed that the variability in moisture sensitivity is driven by treatments ($j$). 
+The prior distribution for the term $a$ was assumed as $\sim \mathcal{N}(3.11, 0.1)$ and for the term $b$ was assumed as $\sim \mathcal{N}(2.42, 0.1)$, utilizing as mean values the values in @SierraMueller and setting up conservatives standard deviations.
 
 ### Temperature
-The temperature scaling function assumes an independent activation energy $E_a$ varying by treatments $j$:
+The temperature scaling function we utilized to define the relationship is the one defined by @Lloyd.1994l68 and it is derived from the original Arrhenius equation [Arrhenius.1889]:
 $$
@@
 $$
-The two parameters $E_a$ and $T_0$ are represented by Bayesian priors.
+The two parameters $A_{k}$ and $E_a$ are represented by Bayesian priors.  
+For the temperature scaling we assumed that the activation energy term $E_a$ (which defines roughly the slope of the relationship) is driven by treatments ($j$) while the linear rescaling term $A_{k}$ is driven by the substrate at the single plot level ($k$).
+This because substrate (organic matter) amount is usually highly variable in soil even at small spatial scales, while its quality, which defines its temperature sensitivity, is assumed to follow the inputs quality which are relatively uniform over each treatments.
+The effect of the rescaling term $A_{k}$ is included in the function for temperature scaling to keep the convention of the original formulation of the equation, but it could be brought outside since it just a linear interaction.
+The prior distribution for the term $A$ was assumed as $\sim \mathcal{N}(300, 100)$ and for the term $E_a$ was assumed as $\sim \mathcal{N}(398.5, 50)$, utilizing as mean values the values  in @Lloyd.1994l68 and setting up conservatives standard deviations.
 
 ### Seasonality
+In order to correct for eventual seasonality patters, we introduce also a seasonality scaling.
 The seasonality scaling functions assume a sine function with period = 365 days, with both amplitude $ampl_j$ (the magnitude) and peak day $peak_j$ (when the function is at its only annual peak) varying as a function of treatments:
@@
 Both the terms $ampl_j$ and $peak_j$ are described by Bayesian priors. Variance is by treatment.
+We assumed that the variability in seaslonalityis driven by treatments ($j$). 
+The prior distribution for the term $ampl$ was assumed as $\sim \mathcal{N}(0, 0.5)$, so assuming as mean no seasonality effect, and for the term $peak$ was assumed as $\sim \mathcal{N}(175, 50)$, so assuming as mean for the peak the 24th of June (Midsummer day).
 
 ## Model benchmark
+The parametric model has been compared against two nonparametric machine learning models.
+The firs was a more conventional multiple linear regression model (default R package `lm`, @R-Core-Team_2023), and the second a random forest model (command `randomForest` from package `randomForest`, @randomForest_2021).
+Both models had the structure as:
+$$
+Rh(t) \sim f(T(t), theta_v(t), treatment)
+$$
+The rationale behind these choices is to offer two extremes to compare the fitness of our parametric model with, assuming that a linear regression model will perform more poorly than the parametric model given its limited capabilities to represent relationships more complex than linear (as we assume temperature and moisture relationships with respiration), and that a random forest model will perform better than the parametric model given its plasticity.
+The random forest model was in this case fit with default parameters, without any other optimization.
 
-## Model extrapolation
-Once we calibrated our parametric model, we could test the consequences of an increase in 5&deg;C
+## Model extrapolation of increased temperature effect on decomposition
+Once we calibrated our parametric model, we could test the consequences of an increase in temperature on decompiosition by driving the model with the same dataset except temperature was increased by 5&deg;C.
+This resulted in a more immediate way of visualizing the impact of a different temperature sensitivity of organic material.
 
 ## Machine learning variance decomposition
-On a more limited subset (only Spain), we tested
+On a more limited subset (only Spain), we tested a more extensive random forest model in order to test the predictive power of other variables on the respiration. The model had structure as:  
+$$
+\begin{aligned}
+Rh(t) \sim f(trees, DBH, BD, por, pH, P, N_{tot}, C{tot}, biom_{fungi}, biom_{bact}, \\
+biom_{act}, biom_{gram+}, biom_{gram-}, biom_{tot}, T(t), theta_v(t), treatment)
+\end{aligned}
+$$
+We included in this model tree density in plants ha$^{-1}$ ($trees$), average diameter at breast height ($DBH$), bulk density ($BD$), porosity ($por$), pH, phosophorous ($P$), total nitrogen ($N_{tot}$), total carbon ($C_{tot}$), fungal biomass ($biom_{fungi}$), bacterial biomass ($biom_{bact}$), actinomycetes biomass ($biom_{act}$), gram+ bacterial biomass ($biom_{gram+}$), gram- bacterial biomass ($biom_{gram-}$), total biomass ($biom_{tot}$), as well as variables mentioned above.
+The aim of this model was to test the predictive potential of any of these additional variables, to explore eventual correlations and possible related processes.
+The choice of a random forest model over a more conventional linear model is motivated by the capacity of the former to consider more complex relationships and interactions, on top of the convenience of being able to assess quantitatively the predictive potential of each variable by estimating relative variable importance.
 
@@
 ## Sites microclimate
+![Climatic parameters measured at the sites, plotted as data density of all measurements.
+Please notice thot not every time series covers the same period of time.](../Figures/climatedata.png){width=60%}
 
@@
 
-![Alt text](../Figures/fit_ML_validation.png)
+![Fit of the two benchmark machine learning models, linear (a) and random forest (b)](../Figures/fit_ML_validation.png){width=80%}
 
-![Alt text](../Figures/fit_model_validation.png)
+![Fit of the parametric model](../Figures/fit_model_validation.png){width=60%}
 
@@
 
-![Alt text](../Figures/posteriors_boxplots.png)
+![Posteriors of the parameters varying as a function of treatments](../Figures/posteriors_boxplots.png){width=60%}
 
-![Alt text](../Figures/posteriors_boxplots_A.png)
+![Posteriors of the scaling parameter, varying as a function of single plots](../Figures/posteriors_boxplots_A.png){width=60%}
 
@@
 
-![Alt text](../Figures/variance_decomposition.png)
+![Fit of the detailed random forest model (a) and variable importance of the same model (b)](../Figures/variance_decomposition.png){width=60%}
+
+## Model extrapolations
+
+![Sensitivity of the organic matter in different treatments to a 5&deg;C increase in temperature, expressed as increase in mineralization from the baseline temperature](../Figures/scenario_extrapolation.png){width=60%}
+
+# Discussion
 
@@
 
+# References
+\bibliography{references}
+
 # Appendix
 
-![Alt text](../Figures/posteriors_temp.png)
+![Posteriors of the parameters of the temperature function, plotted as density distributions](../Figures/posteriors_temp.png)
 
-![Alt text](../Figures/posteriors_seasonality.png)
+![Posteriors of the parameters of the moisture function, plotted as density distributions](../Figures/posteriors_seasonality.png)
 
-![Alt text](../Figures/posteriors_moisture.png)
-
-## Model extrapolations
-
-![Alt text](../Figures/scenario_extrapolation.png)
-
-# References
-\bibliography{references}
+![Posteriors of the parameters of the seasonality function, plotted as density distributions](../Figures/posteriors_moisture.png)

### 2024-08-01 bayesian_correction_MoySOilR.R

@@ -166,6 +166,7 @@ summary(lm)
 abline(a=0, b=1, col="black", lty=2)
 legend("topright", levels(processed_data$treatment), bty="n", pch=1:15, col = palette_treat)
 legend("bottomright", paste("R^2 =", round(summary(lm)$r.squared,3)), bty="n", pch=NA, col = NA)
+legend("bottomleft", "a)", bty="n", pch=NA, col = NA)
 
 plot(predict(linear_model, newdata = validation_data), validation_data$CO2_flux_hour, ylab="predicted", xlab="observed", main="Multiple linear model",
     col=palette_treat[as.numeric(validation_data$treatment)], pch=as.numeric(validation_data$treatment), xlim=c(0,3), ylim=c(0,3))
@@ -173,6 +174,8 @@ lm<-lm(validation_data$CO2_flux_hour ~ predict(linear_model, newdata = validatio
 summary(lm)
 abline(a=0, b=1, col="black", lty=2)
 legend("bottomright", paste("R^2 =", round(summary(lm)$r.squared,3)), bty="n", pch=NA, col = NA)
+legend("bottomleft", "b)", bty="n", pch=NA, col = NA)
+
 dev.off()
 
 

### 2024-08-01 Manuscript_draft/Draft.Rmd
Updated draft

@@
 ---
-title: "Draft"
-author: "Lorenzo Menichetti"
+title: "The impact of forest management on the temperature sensitivity of organic matter"
+author: "Leticia Perez-Izquiero, Lorenzo Menichetti, Jorge Curiel-Juste, etc"
 date: "2024-07-04"
-output: pdf_document
+output:
+  bookdown::pdf_document2: default
+  bookdown::html_document2: default
+  #output: pdf_document
   #   citation_package: natbib
@@
 We included in this model tree density in plants ha$^{-1}$ ($trees$), average diameter at breast height ($DBH$), bulk density ($BD$), porosity ($por$), pH, phosophorous ($P$), total nitrogen ($N_{tot}$), total carbon ($C_{tot}$), fungal biomass ($biom_{fungi}$), bacterial biomass ($biom_{bact}$), actinomycetes biomass ($biom_{act}$), gram+ bacterial biomass ($biom_{gram+}$), gram- bacterial biomass ($biom_{gram-}$), total biomass ($biom_{tot}$), as well as variables mentioned above.
-The aim of this model was to test the predictive potential of any of these additional variables, to explore eventual correlations and possible related processes.
+This model was tested only on a subset of the dataset (only the sites in Spain), for which all the additional variables were available.
+The aim of this model was to test the predictive potential of any of these additional variables, in order to to explore eventual correlations and possible related processes.
 The choice of a random forest model over a more conventional linear model is motivated by the capacity of the former to consider more complex relationships and interactions, on top of the convenience of being able to assess quantitatively the predictive potential of each variable by estimating relative variable importance.
@@
 ## Machine learning variance decomposition
-
-![Fit of the detailed random forest model (a) and variable importance of the same model (b)](../Figures/variance_decomposition.png){width=60%}
+The fitness of the detailed random forest model (fitted only on data from Spain) was the highest of all the models we tested, unsurprisingly given the huge amount of predictors compared with all other models.
+Soil temperature and moisture resulted the most important variables in predicting respiration from each site by far, followed by edaphic charachrteristics of the site (density and porosity), then by treatment, and then again by edaphic site characteristics (namely pH and total C).
+Only afterwards the first variable related to aboveground vegetation.
+Variables related to biomass, total and specific to microorganisms classes, come only later and did not seem particularly important.
+![Fit of the detailed random forest model (a) and variable importance of the same model (b)](../Figures/variance_decomposition.png){width=50%}
 
@@
 # Discussion
+
+## Role of temperature and moisture in conditioning decomposition
+Temperature and moisture seem the most important components in explaining the variance of the soil respiration.
+The seasonality scaling, although apparently differentiating slightly between sites, presents always very low amplitude posteriors, weighting on average only a 5% of the total measured respiration.
+It was more important in France (between 4.5% and 12.2%) and much less in Spain and Romania (between 2.0% and 3.2%).
+This seasonality might represent many different processes, but the most likely are seasonal processes connected with the above-ground vegetation and their inputs to the soil, which in turn would control microbial communities.
+
+## Activation energy of different organic material
+
+## Role of edaphic, ecological and microbiological variables in conditioning decomposition
+Any other variable than soil temperature and moisture did not seem to play a pivotal role in explaining the variance of the soil respiration.
+
+## Consequences of different temperature sensitivity of organic material in reacting to warming
 
@@
 ![Posteriors of the parameters of the seasonality function, plotted as density distributions](../Figures/posteriors_moisture.png)
+
+![Ratio (expressed as %) between the mean amplitude parameter and the mean measured respiration by treatment](../Figures/Amplitude_ratio.png)

### 2024-08-01 bayesian_correction_MoySOilR.R

@@ -9,11 +9,8 @@ library(lubridate)
 library(readxl)
 library(caret)
 
-#TODO microclimatic plots
 #TODO is there an interaction between texture and the TMS calibration?
 #TODO distance between control and treatments
-#TODO plot the treatments for each A
-#TODO unit of A
 #TODO Q10?
 
 palette_treat <-  c(paletteer::paletteer_c("ggthemes::Orange-Gold", n=5),
@@ -311,7 +308,7 @@ par(mfrow=c(1,2))
 E_0_densities_indA_Moy <- list()
 E_0_box_indA_Moy <- list()
 plot( density(post_bytreat_indA_Moy$Ea[,1]), xlim=c(range(post_bytreat_indA_Moy$Ea)[1]*0.9, range(post_bytreat_indA_Moy$Ea)[2]*1.1),
-      ylim=c(0,0.03), xlab = expression('E'[0]), col=NA, main="Temperature scaling")
+      ylim=c(0,0.07), xlab = expression('E'[0]), col=NA, main="Temperature scaling")
 for(i in 1:stan_data$Tr){
   E_0_densities_indA_Moy[[i]] <- density(post_bytreat_indA_Moy$Ea[,i])
   E_0_box_indA_Moy[[i]] <- E_0_densities_indA_Moy[[i]]$x
@@ -358,7 +355,7 @@ for(i in 1:stan_data$Tr){
 amplitude_densities_indA_Moy <- list()
 amplitude_box_indA_Moy <- list()
 plot( density(post_bytreat_indA_Moy$amplitude[,1]), xlim=c(range(post_bytreat_indA_Moy$amplitude)[1]*0.9, range(post_bytreat_indA_Moy$amplitude)[2]*1.1),
-      ylim=c(0, 20), xlab = expression('amplitude'), col=NA, main="Seasonality scaling")
+      ylim=c(0, 70), xlab = expression('amplitude'), col=NA, main="Seasonality scaling")
 for(i in 1:stan_data$Tr){
   amplitude_densities_indA_Moy[[i]] <- density(post_bytreat_indA_Moy$amplitude[,i])
   polygon(amplitude_densities_indA_Moy[[i]], col=add.alpha(palette_treat[i],0.25), border = add.alpha(palette_treat[i],0.6))
@@ -377,7 +374,7 @@ par(mfrow=c(1,2))
 a_densities_indA_Moy <- list()
[…]

### 2024-08-06 Manuscript_draft/Draft.Rmd
Just a test
Updated Fig 1, with climate data. I made only one figure instead of one per site, it's busy but we would reduce the nr of figs if it worked.

@@
 output:
+  bookdown::html_document2: default
   bookdown::pdf_document2: default
-  bookdown::html_document2: default
   #output: pdf_document
@@
 # Abstract
+TEST CHANGE
 
 # Introduction
+
+## Soil C turnover is controlled by temperature and moisture
+
+## Climate change will influence soil C turnover globally
 
@@
 ## Sites microclimate
-![Climatic parameters measured at the sites, plotted as data density of all measurements.
-Please notice thot not every time series covers the same period of time.](../Figures/climatedata.png){width=60%}
+<!-- ![Climatic parameters measured at the sites, plotted as data density of all measurements.
+Please notice thot not every time series covers the same period of time.](../Figures/climatedata.png){width=60%} -->
+![Climatic parameters and respirations measured at the sites (mean values after loess interpolation with span = 0.5).
+Please notice thot not every time series covers the same period of time.
+Red shades = France, green shades = Romania, blue shades = Spain](../Figures/climatedata_means.png){width=60%}
 

### 2024-08-06 bayesian_correction_MoySOilR.R

@@ -8,10 +8,22 @@ library(dplyr)
 library(lubridate)
 library(readxl)
 library(caret)
+library(MASS)
 
-#TODO is there an interaction between texture and the TMS calibration?
-#TODO distance between control and treatments
+#TODO DISCUSSION is there an interaction between texture and the TMS calibration?
 #TODO Q10?
+
+#TODO normalization of respiration by SOC when data are there
+
+#DONE interpolated line of temperature by treatment (average)
+#DONE add the same graph for respiration
+#DONE try to make them side by side by site (3 plots)
+
+#TODO plot the reduction functions in appendix
+
+#TODO reorder the treatments with control first
+#TODO treatments control thinning clearcut with different shades, and slah with 45 degrees lines
+
 
 palette_treat <-  c(paletteer::paletteer_c("ggthemes::Orange-Gold", n=5),
                     paletteer::paletteer_c("ggthemes::Green", n=5),
@@ -35,7 +47,6 @@ data$seq <- 1:dim(data)[1]
 
 
 data$Date <- as.Date(data$Date, format = "%d.%m.%Y")
-
 
 
 processed_data <- data.frame(ID = data$seq,
@@ -70,18 +81,18 @@ processed_data_filtered_preprocess$plot_id <- droplevels(processed_data_filtered
 ########### site micrometeorology
 
 png("./Figures/climatedata.png", height=2500, width = 2000, res=300)
+par(mfrow=c(3,2), mar=c(2,5,1,1))
 
 data = data.frame(Date = processed_data_filtered_preprocess$date,
                   Site= processed_data_filtered_preprocess$site,
                   Temperature = processed_data_filtered_preprocess$T1_soil,
                   Moisture = processed_data_filtered_preprocess$Soil_moist * 100)
 
-sites <- levels(data_long$Site)
+sites <- levels(data$Site)
 capitalized_sites <- paste0(toupper(substr(sites, 1, 1)), substr(sites, 2, nchar(sites)))
 data$DateNumeric <- as.numeric(data$Date)
 extended_date_range <- seq(min(data$DateNumeric) - 30, max(data$DateNumeric) + 30, length.out = 300)
 
-par(mfrow=c(3,2), mar=c(2,5,1,1))
 for(i in 1:length(sites)){
 subset<-data[data$Site== sites[i],]
 # Remove NAs
[…]

### 2024-08-13 bayesian_correction_MoySOilR.R
Updated the draft and figures as we said last time

@@ -9,29 +9,23 @@ library(lubridate)
 library(readxl)
 library(caret)
 library(MASS)
+library(parallel)
 
 #TODO DISCUSSION is there an interaction between texture and the TMS calibration?
 #TODO Q10?
 
 #TODO normalization of respiration by SOC when data are there
+#TODO convert volumetric water content to theta_s?
 
 #DONE interpolated line of temperature by treatment (average)
 #DONE add the same graph for respiration
 #DONE try to make them side by side by site (3 plots)
 
-#TODO plot the reduction functions in appendix
+#DONE plot the reduction functions in appendix
 
-#TODO reorder the treatments with control first
-#TODO treatments control thinning clearcut with different shades, and slah with 45 degrees lines
+#DONE reorder the treatments with control first
+#DONE in the boxplots treatments control thinning clearcut with different shades, and slah with 45 degrees lines
 
-
-palette_treat <-  c(paletteer::paletteer_c("ggthemes::Orange-Gold", n=5),
-                    paletteer::paletteer_c("ggthemes::Green", n=5),
-                    paletteer::paletteer_c("ggthemes::Blue-Teal", n=5))
-
-palette_plot <-  c(paletteer::paletteer_c("ggthemes::Orange-Gold", n=length(unique(data[data$site=="france",]$plot_ID))),
-                   paletteer::paletteer_c("ggthemes::Green", n=length(unique(data[data$site=="romania",]$plot_ID))),
-                   paletteer::paletteer_c("ggthemes::Blue-Teal", n=length(unique(data[data$site=="spain",]$plot_ID))))
 
 
 
@@ -47,6 +41,25 @@ data$seq <- 1:dim(data)[1]
 
 
 data$Date <- as.Date(data$Date, format = "%d.%m.%Y")
+
+#rearrange the levels putting control first
+data$treatment <- factor(data$treatment, levels = c("control", "clear_cut_no_slash", "clear_cut_slash", "thinning_no_slash", "thinning_slash"))
+
+
+
+
+palette_treat <-  c(paletteer::paletteer_c("ggthemes::Orange-Gold", n=5),
+                    paletteer::paletteer_c("ggthemes::Green", n=5),
+                    paletteer::paletteer_c("ggthemes::Blue-Teal", n=5))
+palette_treat_simplified <- c(palette_treat[1], palette_treat[3], palette_treat[3], palette_treat[5], palette_treat[5],
[…]

### 2024-08-13 functions_plotting.R

@@ -1,0 +1,45 @@
+
+
+png("./Figures/Temp.png", width = 2000, height = 2000, res=300)
+temp = seq(0,45)
+xi_temp = exp(-311.9924/ ((temp+ 273.15) - 227.13));
+
+plot(temp, xi_temp, type = "l", main = "Loyd-Taylor",
+     xlab = "Temperature", ylab = "Activity", col = "firebrick")
+
+dev.off()
+
+
+png("./Figures/Moist.png", width = 2000, height = 2000, res=300)
+
+### From Sierra SoilR
+fW.Moyano <- function(theta, a = 3.11, b = 2.42) {
+  a * theta - b * theta^2
+}
+
+# Example usage
+th <- seq(0, 1, 0.01)
+xi <- fW.Moyano(theta = th)
+
+plot(th, xi, type = "l", main = "Moyano 2013",
+     xlab =expression("Volumetric soil water content (cm"^3~"cm"^-3~")"),  ylab = "Activity", col = "dodgerblue3")
+dev.off()
+
+
+png("./Figures/Season.png", width = 2000, height = 2000, res=300)
+
+day_year = seq(1:365)
+sine_wave =  0.1183339 * cos((2 * pi / 365) * day_year + (2 * pi/ 365) * (176.3293 - 1) - pi / 2);
+
+plot(day_year, sine_wave, type = "l", main = "Sine seasonality",
+     xlab = "Day of the year", ylab = "Additive seasonality pattern", col = "darkorange")
+
+dev.off()
+
+ampl=0
+peak=175
+
+
+ampl * cos((2 * pi / 365) * day_year + (2 * pi / 365) * (peak - 1) - pi / 2)
+
+

### 2024-08-20 Manuscript_draft/Draft.Rmd
Updated the draft and figures as we said last time
Changes in M&M 20/08/2024

@@
 Christol in February 2023 and in Dumbravita in May 2022.
-Forest stand characteristics
+
+## Forest stand characteristics
 Within each of the control and thinned stands (n=8), we recorded the number of trees to calculate tree density (No.
@@
 For further modelling and since these sensors measure every 15 min, we extracted the soil temperature and soil moisture measurements closer to the time we measured soil respiration in the field.
-Soil physical-chemical analyses
+
+## Soil physical-chemical analyses
 Chemical analyses were measured on 5 soil cores of 4 cm diameter separated 0.5 m to a depth of 10 cm along the 2m monitoring transect.
@@
 Soil C stocks (g m2) were calculated by multiplying C concentration by bulk density and sampling depth.
-Microbial biomass
+
+## Microbial biomass
 The same five soil cores used for chemical analyses were used to estimate soil microbial biomass.
@@
 ![Ratio (expressed as %) between the mean amplitude parameter and the mean measured respiration by treatment](../Figures/Amplitude_ratio.png)
+
+![Temperature scaling function](../Figures/Temp.png){width=50%}
+
+![Moisture scaling function](../Figures/Moist.png){width=50%}
+
+![Seasonality scaling function](../Figures/Season.png){width=50%}

### 2024-08-21 Manuscript_draft/Draft.Rmd
Changes on introduction to see

@@
 ---
-title: "The impact of forest management on the temperature sensitivity of organic matter"
-author: "Leticia Perez-Izquiero, Lorenzo Menichetti, Jorge Curiel-Juste, etc"
+title: "The impact of forest management on the temperature sensitivity of organic matter turnover"
+author: "Leticia Pérez-Izquiero, Lorenzo Menichetti, Jorge Curiel-Juste, etc"
 date: "2024-07-04"
@@
 # Introduction
+The soil is one of the largest terrestrial reservoirs of carbon, containing approximately three times more carbon than the atmosphere or the vegetation on Earth (IPCC, 2021).
+Soil carbon is released to the atmosphere as carbon dioxide (CO₂) through decomposition of organic matter by microbes (heterotrophic respiration) and root respiration (autotrophic respiration).
+This process, i.e., soil respiration, constitutes the main carbon efflux into the atmosphere (ref).
+Under the new climate change scenarios increases of Co2 efflux from soils are expected (Schlesinger and Andrews, 2000), in turn this can imply a positive feedback to future warming (ref).
+Because soil respiration is highly sensitive to soil temperature and soil moisture (ref), understanding the responses of soil respiration to these microclimatic conditions, which are affected by global change factors as climate change and forest management is critical for predicting future atmospheric CO2 concentrations.
 
@@
 
+## Forest management changes the soil C turnover sensitivity to microclimate
+
 ## Climate change will influence soil C turnover globally
+
+In this study, during three years we have monthly recorded data on soil respiration and microclimate in three replicated Quercus forests across Europe subjected to different forest management strategies.
+We have analysed the sensitivity of soil respiration to microclimate (i.e., soil temperature and soil moisture) by constructing a mechanistic model…
 
@@
 ## Model extrapolation of increased temperature effect on decomposition
-Once we calibrated our parametric model, we could test the consequences of an increase in temperature on decompiosition by driving the model with the same dataset except temperature was increased by 5&deg;C.
+Once we calibrated our parametric model, we could test the consequences of an increase in temperature on decomposition by driving the model with the same dataset except temperature was increased by 5&deg;C.
 This resulted in a more immediate way of visualizing the impact of a different temperature sensitivity of organic material.
@@
 The fitness of the detailed random forest model (fitted only on data from Spain) was the highest of all the models we tested, unsurprisingly given the huge amount of predictors compared with all other models.
-Soil temperature and moisture resulted the most important variables in predicting respiration from each site by far, followed by edaphic charachrteristics of the site (density and porosity), then by treatment, and then again by edaphic site characteristics (namely pH and total C).
+Soil temperature and moisture resulted the most important variables in predicting respiration from each site by far, followed by edaphic characteristics of the site (density and porosity), then by treatment, and then again by edaphic site characteristics (namely pH and total C).
 Only afterwards the first variable related to aboveground vegetation.

### 2024-08-23 Manuscript_draft/Draft.Rmd
Minor commit after ouor talk, syncing just in case

@@
 This process, i.e., soil respiration, constitutes the main carbon efflux into the atmosphere (ref).
-Under the new climate change scenarios increases of Co2 efflux from soils are expected (Schlesinger and Andrews, 2000), in turn this can imply a positive feedback to future warming (ref).
+Under the new climate change scenarios increases of CO₂ efflux from soils are expected (Schlesinger and Andrews, 2000), in turn this can imply a positive feedback to future warming (ref).
 Because soil respiration is highly sensitive to soil temperature and soil moisture (ref), understanding the responses of soil respiration to these microclimatic conditions, which are affected by global change factors as climate change and forest management is critical for predicting future atmospheric CO2 concentrations.

### 2024-08-23 bayesian_correction_MoySOilR.R

@@ -11,11 +11,14 @@ library(caret)
 library(MASS)
 library(parallel)
 
-#TODO DISCUSSION is there an interaction between texture and the TMS calibration?
-#TODO Q10?
 
 #TODO normalization of respiration by SOC when data are there
-#TODO convert volumetric water content to theta_s?
+
+#TODO test variable E_a over time, remove seasonality scaling and use E_a by time
+
+#TODO Enzyme relationships
+
+#TODO convert volumetric water content to theta_s? When data come
 
 #DONE interpolated line of temperature by treatment (average)
 #DONE add the same graph for respiration

### 2024-08-29 Manuscript_draft/Draft.Rmd
Some ideas on introduction

@@
 The soil is one of the largest terrestrial reservoirs of carbon, containing approximately three times more carbon than the atmosphere or the vegetation on Earth (IPCC, 2021).
-Soil carbon is released to the atmosphere as carbon dioxide (CO₂) through decomposition of organic matter by microbes (heterotrophic respiration) and root respiration (autotrophic respiration).
-This process, i.e., soil respiration, constitutes the main carbon efflux into the atmosphere (ref).
-Under the new climate change scenarios increases of CO₂ efflux from soils are expected (Schlesinger and Andrews, 2000), in turn this can imply a positive feedback to future warming (ref).
-Because soil respiration is highly sensitive to soil temperature and soil moisture (ref), understanding the responses of soil respiration to these microclimatic conditions, which are affected by global change factors as climate change and forest management is critical for predicting future atmospheric CO2 concentrations.
+Soil carbon is released to the atmosphere as carbon dioxide (CO2) through decomposition of organic matter by microbes (heterotrophic respiration) and root respiration (autotrophic respiration).
+This process (i.e., soil respiration) constitutes the main carbon efflux into the atmosphere (ref).
+Under the new climate change scenarios increases of CO2 efflux from soils are expected (Schlesinger and Andrews, 2000), this can imply a positive feedback to future warming if stored carbon is mobilized and carbon input to soil does not increase (Lamberty&Thomson, 2010).
+Because soil respiration is highly sensitive to soil temperature and soil moisture (Nissan et al.
+2022), understanding the responses of soil respiration to these microclimatic conditions, which are affected by global change factors as climate change and forest management, is critical for predicting future atmospheric CO2 concentrations.
+Temperature affects respiration rates positively (Lloyd and Taylor, 1994), by increasing the metabolic rates of soil microorganisms and enzyme efficiency (ref).
+However, the relationship with soil moisture is non-linear.
+Low water content limit nutrient diffusion and enzymes and exert drought-stress on microbes while high water contents limits oxygen (O2) diffusion, creating anaerobic conditions that suppress aerobic microbial respiration (ref).
+Something here about modelling?
 
-## Soil C turnover is controlled by temperature and moisture
+Forest management have the potential to affect respiration rates by modifying the microclimate.
+Removal of trees by forest logging increases the exposure of soil to temperature fluctuations and alter the soil water content either by decreasing water through soil evaporation or by increasing it by reducing the transpiration of water by trees (Zhang et al 2018).
+The degree of the response of soil will be dependent on the severity of the forest intervention.
+Thus, more severe forestry intervention such as clear-cuts may increase soil moisture, enhance heat and frost damage (Kovaks et al 2019; Zhang et al 2018) and it seems to reduce respiration rates (Yang et el 2022; Zhang et al 2018).
+Less severe interventions such as thinning, have shown to increase respiration in broadleaf forests in a meta-analysis (Zhang et al 2018).
+Negative effects of forest management on soil could be buffered by the addition of forest residues on soil (i.e., slash addition), which can improve the microclimatic conditions, nutrient fertility and biodiversity (ref).
+We do not know whether the response of soil CO2 effluxes to the microclimatic conditions (i.e.
+soil temperature and moisture) differs among different forest management strategies.
+Understanding the sensitivity of respiration to microclimate under different forest management strategies is crucial in a context of climate-smart forestry (CSF) development.
 
-## Forest management changes the soil C turnover sensitivity to microclimate
+Here, we hypothesized that logging of trees will decrease the respiration rates because of the strong decrease in vegetation activity and microbiota associated.
+In turn, addition of slash will provide labile organic matter to decompose increasing the respiration rates.
+We also hypothesized that the variation in soil respiration would be mainly dependent on microclimatic variables (soil temperature and moisture), especially in the management treatments where vegetation was removed, since heterotrophic respiration will prevail over autotrophic respiration.
+Moreover, the sensitivity of soil respiration to microclimatic conditions will depend on the type of climate/forest, being the Mediterranean the most sensitive as the microbial communities will be better adapted to the strong seasonal changes in water availability.
 
-## Climate change will influence soil C turnover globally
-
-In this study, during three years we have monthly recorded data on soil respiration and microclimate in three replicated Quercus forests across Europe subjected to different forest management strategies.
+To test these hypotheses, we have monthly recorded data on soil respiration and microclimate in three replicated Quercus forests across Europe subjected to different forest management strategies during three years.
 We have analysed the sensitivity of soil respiration to microclimate (i.e., soil temperature and soil moisture) by constructing a mechanistic model…

### 2024-09-06 Manuscript_draft/Draft.Rmd
Restructuring the code and trying to fit normalized resp

@@
 
-## Machine learning variance decomposition
+## Machine learning variance decomposition: predicting the whole dataset based on ecological variables
 The fitness of the detailed random forest model (fitted only on data from Spain) was the highest of all the models we tested, unsurprisingly given the huge amount of predictors compared with all other models.
@@
 ![Fit of the detailed random forest model (a) and variable importance of the same model (b)](../Figures/variance_decomposition.png){width=50%}
+
+## Machine learning variance decomposition: predicting a substet based on soil enzymatic data
+
+![Fit of the detailed random forest model (a) and variable importance of the same model (b)](../Figures/variance_decomposition_enzyme.png){width=50%}
 

### 2024-09-06 bayesian_correction_MoySOilR.R

@@ -11,6 +11,19 @@ library(caret)
 library(MASS)
 library(parallel)
 
+set.seed(123)
+
+#TODO Use basal area instead of DBH in the RF model
+
+#TODO NORMALIZE THE TEMPERATURE SCENARIOS!!
+
+#TODO: E_o and E_a are mixed up in the manuscript
+
+#TODO plot SOC stocks by site
+
+#TODO: correlation. between E_a and enzyme activities
+
+#NOTE: CUE and aciivation energy: in reality they relate, modelwise no, it's a shortcoming of having a model not distinguishing for C quality
 
 #TODO normalization of respiration by SOC when data are there
 
@@ -32,9 +45,13 @@ library(parallel)
 
 
 
-edaphics <- read.csv("edaphics.csv")
+# edaphics <- read.csv("edaphics.csv")
 #data <- read.csv("Database_co2_28_05_2024.csv")
-data <- read_xlsx("Database_co2_28_05_2024.xlsx", sheet=1)
+data <- read_xlsx("Database_co2_03_09_2024_format_Lorenzo.xlsx", sheet=1)
+str(data)
+soil_data <- read_xlsx("Soil_data_Lorenzo_22_08_2024_enzymes.xlsx", sheet=1)
+
+
 data$treatment <- as.factor(data$treatment)
 data$site <- as.factor(data$site)
 data$Soil_moist <- as.numeric(data$Soil_moist)/100
@@ -42,14 +59,179 @@ data$T1_soil <- as.numeric(data$T1_soil)
 data$CO2_flux_hour <- as.numeric(data$CO2_flux_hour)
 data$seq <- 1:dim(data)[1]
 
+soil_data
 
 data$Date <- as.Date(data$Date, format = "%d.%m.%Y")
+soil_data$date  <- as.Date(soil_data$date, format = "%d.%m.%Y")
 
 #rearrange the levels putting control first
 data$treatment <- factor(data$treatment, levels = c("control", "clear_cut_no_slash", "clear_cut_slash", "thinning_no_slash", "thinning_slash"))
 
 
+boxplot(soil_data$bulk_den ~soil_data$site)
+boxplot(soil_data$Ctot ~soil_data$site)
+boxplot(soil_data$C_stocks ~soil_data$site)
 
+boxplot(data$CO2_flux_hour ~ data$site)
+plot(data[data$site=="romania",]$Date,
[…]

### 2024-09-06 data_analysis_sections/ML_model_benchmarks.R

@@ -1,0 +1,70 @@
+
+
+## stratified data split for validation
+set.seed(123) # Set seed for reproducibility
+
+# Create stratification group
+strat_group <- interaction(processed_data_filtered_preprocess$gen_treatment, processed_data_filtered_preprocess$country)
+
+# Create stratified partition
+train_indices <- createDataPartition(strat_group, p = 0.8, list = FALSE)
+
+# Split the data into training and validation sets
+processed_data_filtered <- processed_data_filtered_preprocess[train_indices, ]
+validation_data <- processed_data_filtered_preprocess[-train_indices, ]
+
+dim(validation_data)
+dim(processed_data_filtered)
+
+
+#Regression Models
+linear_model = lm(CO2_flux_norm ~ T1_soil * Soil_moist * treatment, processed_data_filtered)
+summary(linear_model)$r.squared
+plot(predict(linear_model, newdata = validation_data), validation_data$CO2_flux_norm)
+
+rf_model = randomForest(CO2_flux_norm ~ T1_soil * Soil_moist * treatment, processed_data_filtered)
+summary(rf_model)
+rf_fit <- lm(predict(rf_model) ~ processed_data_filtered$CO2_flux_norm)
+plot(predict(rf_model), processed_data_filtered$CO2_flux_hour)
+summary(rf_fit)$r.squared
+
+
+
+
+
+png("./Figures/fit_ML_validation.png", height = 2000, width = 4000, res=300)
+
+range_plot = range(c(predict(rf_model, newdata = validation_data),
+                     validation_data$CO2_flux_norm,
+                     predict(linear_model, newdata = validation_data)))
+
+par(mfrow=c(1,2))
+plot(predict(rf_model, newdata = validation_data), validation_data$CO2_flux_norm, ylab="predicted", xlab="observed", main="RF model",
+     col=palette_treat[as.numeric(validation_data$treatment)], pch=as.numeric(validation_data$treatment), xlim=range_plot, ylim=range_plot)
+lm<-lm(validation_data$CO2_flux_norm ~ predict(rf_model, newdata = validation_data))
+summary(lm)
+abline(a=0, b=1, col="black", lty=2)
+legend("topright", levels(processed_data$treatment), bty="n", pch=1:15, col = palette_treat)
+legend("bottomright", paste("R^2 =", round(summary(lm)$r.squared,3)), bty="n", pch=NA, col = NA)
[…]

### 2024-09-06 data_analysis_sections/ML_variance_decomposition_enzymes.R

@@ -1,0 +1,94 @@
+
+
+predictor_list = c("No_trees_ha", "treatment",
+                   "Basal_area","bulk_den", "Ntot", "C_stocks",
+                   "beta_gluco", "acid_phos", "beta_xylo", "chitise", #"phosphorous",  "pH",
+                   "cellobiohydrolase", "beta_galacto", "alpha_gluco", "lipase", "CO2_flux_norm",
+                   "T1_soil", "Soil_moist" ,"month" )
+
+dim(na.omit(soil_data[,predictor_list]))
+dim(soil_data[,predictor_list])
+soil_data$CO2_flux_hour
+
+# Create stratification group
+soil_strat_group <- interaction(soil_data$treatment, soil_data$site)
+
+# Create stratified partition
+train_indices <- createDataPartition(soil_strat_group, p = 0.8, list = FALSE)
+
+# Split the data into training and validation sets
+soil_data_train <- soil_data[train_indices, ]
+soil_data_valid <- soil_data[-train_indices, ]
+
+rf_soil_data_train <- na.omit(soil_data_train[,predictor_list])
+rf_soil_data_valid <- na.omit(soil_data_valid[,predictor_list])
+
+rf_soil_data_train$treatment <- as.factor(rf_soil_data_train$treatment)
+rf_soil_data_valid$treatment <- as.factor(rf_soil_data_valid$treatment)
+
+
+dim(rf_soil_data_train)
+dim(rf_soil_data_valid)
+
+
+
+# Define the grid of hyperparameters to search
+tune_grid <- expand.grid(
+  mtry = c(3, 4, 5, 6) # Adjust based on the number of predictors
+)
+
+# Define cross-validation method
+control <- trainControl(
+  method = "cv",
+  number = 5,
+  search = "grid"
+)
+
+# Perform the tuning
+set.seed(123) # For reproducibility
+tuned_rf <- train(
+  CO2_flux_norm ~ .,
+  data = rf_soil_data_train,
+  method = "rf",
+  trControl = control,
+  tuneGrid = tune_grid,
+  ntree = 500 # or another fixed value
+)
+
+
+rf_model_CO2_flux <- randomForest(CO2_flux_norm ~ . , data = rf_soil_data_train,
+                                  mtry = tuned_rf$bestTune$mtry)
+
[…]

### 2024-09-06 data_analysis_sections/parametric_model_fitting.R

@@ -1,0 +1,428 @@
+library(parallel)
+
+# Set the number of cores to use
+options(mc.cores = parallel::detectCores())
+
+# test a bit of the data
+
+plot(density(processed_data_filtered$CO2_flux_norm))
+polygon(density(processed_data_filtered$CO2_flux_hour))
+range(processed_data_filtered$CO2_flux_norm)
+range(processed_data_filtered$CO2_flux_hour)
+
+
+# Data list for Stan
+stan_data <- list(N = length(processed_data_filtered$CO2_flux_norm),
+                  Tr = length(levels(processed_data_filtered$treatment)),
+                  treatment = as.numeric(processed_data_filtered$treatment),
+                  resp = processed_data_filtered$CO2_flux_norm,
+                  temp = processed_data_filtered$T1_soil,
+                  M = processed_data_filtered$Soil_moist,
+                  Pl = length(levels(processed_data_filtered$plot_id)),
+                  plot_id = as.integer(processed_data_filtered$plot_id),
+                  day_year = yday(as.Date(processed_data_filtered$date))
+)
+
+stan_data_valid <- list(N = length(validation_data$CO2_flux_norm),
+                        Tr = length(levels(validation_data$treatment)),
+                        treatment = as.numeric(validation_data$treatment),
+                        resp = validation_data$CO2_flux_norm,
+                        temp = validation_data$T1_soil,
+                        M = validation_data$Soil_moist,
+                        Pl = length(levels(validation_data$plot_id)),
+                        plot_id = as.integer(validation_data$plot_id),
+                        day_year = yday(as.Date(validation_data$date))
+)
+
+#quick check of the data...
+str(stan_data)
+plot(density(processed_data_filtered$CO2_flux_norm))
+polygon(density(validation_data$CO2_flux_norm))
+
+
[…]

### 2024-09-06 data_analysis_sections/site_micrometeorology.R

@@ -1,0 +1,174 @@
+
+########### site micrometeorology
+
+png("./Figures/climatedata.png", height=2500, width = 2000, res=300)
+par(mfrow=c(3,2), mar=c(2,5,1,1))
+
+data_meteo = data.frame(Date = processed_data_filtered_preprocess$date,
+                  Site= processed_data_filtered_preprocess$site,
+                  Temperature = processed_data_filtered_preprocess$T1_soil,
+                  Moisture = processed_data_filtered_preprocess$Soil_moist * 100)
+
+sites <- levels(data_meteo$Site)
+capitalized_sites <- paste0(toupper(substr(sites, 1, 1)), substr(sites, 2, nchar(sites)))
+data_meteo$DateNumeric <- as.numeric(data_meteo$Date)
+extended_date_range <- seq(min(data_meteo$DateNumeric) - 30, max(data_meteo$DateNumeric) + 30, length.out = 300)
+
+for(i in 1:length(sites)){
+  subset<-data_meteo[data_meteo$Site== sites[i],]
+  # Remove NAs
+  data_clean <- subset %>% filter(!is.na(Temperature))
+  data_clean$DateNumeric <- as.numeric(data_clean$Date)
+
+  # Calculate the 2D kernel density estimate
+  dens_temp <- kde2d(data_clean$DateNumeric, data_clean$Temperature, n = 300, lims = c(range(extended_date_range), c(0,30)))
+  dens_moist <- kde2d(data_clean$DateNumeric, data_clean$Moisture, n = 300, lims = c(range(extended_date_range), c(0,60)))
+
+  custom_palette_temp <- colorRampPalette(c("white", "goldenrod1", "orangered", "red4"))(100)
+  custom_palette_moist <- colorRampPalette(c("white", "cadetblue2", "dodgerblue", "darkblue"))(100)
+
+  average_temperatures <- tapply(data_clean$Temperature, data_clean$DateNumeric, mean, na.rm = TRUE)
+  average_temperatures_df <- data.frame(DateNumeric = as.numeric(names(average_temperatures)), AverageTemperature = as.numeric(average_temperatures))
+  loess_model_temp <- loess(AverageTemperature ~ DateNumeric, data = average_temperatures_df)
+  date_seq_temp <- seq(min(data_clean$DateNumeric), max(data_clean$DateNumeric), by = 1)
+  predicted_temperatures <- predict(loess_model_temp, newdata = data.frame(DateNumeric = date_seq_temp))
[…]

### 2024-09-06 master_data_analysis.R

@@ -1,0 +1,470 @@
+
+
+library(rstan)
+library(paletteer)
+library(prettyGraphs)
+library(randomForest)
+library(dplyr)
+library(lubridate)
+library(readxl)
+library(caret)
+library(MASS)
+library(parallel)
+
+set.seed(123)
+
+#TODO Use basal area instead of DBH in the RF model
+
+#TODO NORMALIZE THE TEMPERATURE SCENARIOS!!
+
+#TODO: E_o and E_a are mixed up in the manuscript
+
+#TODO plot SOC stocks by site
+
+#TODO: correlation. between E_a and enzyme activities
+
+#NOTE: CUE and aciivation energy: in reality they relate, modelwise no, it's a shortcoming of having a model not distinguishing for C quality
+
+#TODO normalization of respiration by SOC when data are there
+
+#TODO test variable E_a over time, remove seasonality scaling and use E_a by time
+
+#TODO Enzyme relationships
+
+#TODO convert volumetric water content to theta_s? When data come
+
+#DONE interpolated line of temperature by treatment (average)
+#DONE add the same graph for respiration
+#DONE try to make them side by side by site (3 plots)
+
+#DONE plot the reduction functions in appendix
+
+#DONE reorder the treatments with control first
+#DONE in the boxplots treatments control thinning clearcut with different shades, and slah with 45 degrees lines
+
+
+
+
+
+# edaphics <- read.csv("edaphics.csv")
+#data <- read.csv("Database_co2_28_05_2024.csv")
+data <- read_xlsx("Database_co2_03_09_2024_format_Lorenzo.xlsx", sheet=1)
+str(data)
+soil_data <- read_xlsx("Soil_data_Lorenzo_03_09_2024_enzymes.xlsx", sheet=1)
+
+
+data$treatment <- as.factor(data$treatment)
+data$site <- as.factor(data$site)
+data$Soil_moist <- as.numeric(data$Soil_moist)/100
+data$T1_soil <- as.numeric(data$T1_soil)
+data$CO2_flux_hour <- as.numeric(data$CO2_flux_hour)
+data$seq <- 1:dim(data)[1]
+
+soil_data
+
+data$Date <- as.Date(data$Date, format = "%d.%m.%Y")
+soil_data$date  <- as.Date(soil_data$date, format = "%d.%m.%Y")
+
+#rearrange the levels putting control first
[…]

### 2024-09-11 bayesian_correction_MoySOilR.R
Finally the parametric fitting works!

@@ -1,1210 +1,0 @@
-
-
-library(rstan)
-library(paletteer)
-library(prettyGraphs)
-library(randomForest)
-library(dplyr)
-library(lubridate)
-library(readxl)
-library(caret)
-library(MASS)
-library(parallel)
-
-set.seed(123)
-
-#TODO Use basal area instead of DBH in the RF model
-
-#TODO NORMALIZE THE TEMPERATURE SCENARIOS!!
-
-#TODO: E_o and E_a are mixed up in the manuscript
-
-#TODO plot SOC stocks by site
-
-#TODO: correlation. between E_a and enzyme activities
-
-#NOTE: CUE and aciivation energy: in reality they relate, modelwise no, it's a shortcoming of having a model not distinguishing for C quality
-
-#TODO normalization of respiration by SOC when data are there
-
-#TODO test variable E_a over time, remove seasonality scaling and use E_a by time
-
-#TODO Enzyme relationships
-
-#TODO convert volumetric water content to theta_s? When data come
-
-#DONE interpolated line of temperature by treatment (average)
-#DONE add the same graph for respiration
-#DONE try to make them side by side by site (3 plots)
-
-#DONE plot the reduction functions in appendix
-
-#DONE reorder the treatments with control first
-#DONE in the boxplots treatments control thinning clearcut with different shades, and slah with 45 degrees lines
-
-
-
-
-# edaphics <- read.csv("edaphics.csv")
-#data <- read.csv("Database_co2_28_05_2024.csv")
-data <- read_xlsx("Database_co2_03_09_2024_format_Lorenzo.xlsx", sheet=1)
-str(data)
-soil_data <- read_xlsx("Soil_data_Lorenzo_22_08_2024_enzymes.xlsx", sheet=1)
-
-
-data$treatment <- as.factor(data$treatment)
-data$site <- as.factor(data$site)
-data$Soil_moist <- as.numeric(data$Soil_moist)/100
-data$T1_soil <- as.numeric(data$T1_soil)
-data$CO2_flux_hour <- as.numeric(data$CO2_flux_hour)
-data$seq <- 1:dim(data)[1]
-
-soil_data
-
-data$Date <- as.Date(data$Date, format = "%d.%m.%Y")
-soil_data$date  <- as.Date(soil_data$date, format = "%d.%m.%Y")
-
-#rearrange the levels putting control first
[…]

### 2024-09-11 data_analysis_sections/ML_model_benchmarks.R

@@ -25,9 +25,6 @@ plot(predict(linear_model, newdata = validation_data), validation_data$CO2_flux_
 rf_model = randomForest(CO2_flux_norm ~ T1_soil * Soil_moist * treatment, processed_data_filtered)
 summary(rf_model)
 rf_fit <- lm(predict(rf_model) ~ processed_data_filtered$CO2_flux_norm)
-plot(predict(rf_model), processed_data_filtered$CO2_flux_hour)
-summary(rf_fit)$r.squared
-
 
 
 

### 2024-09-11 data_analysis_sections/ML_variance_decomposition_enzymes.R

@@ -1,7 +1,7 @@
 
 
 predictor_list = c("No_trees_ha", "treatment",
-                   "Basal_area","bulk_den", "Ntot", "C_stocks",
+                   "Basal_area","bulk_den", #"Ntot", "C_stocks",
                    "beta_gluco", "acid_phos", "beta_xylo", "chitise", #"phosphorous",  "pH",
                    "cellobiohydrolase", "beta_galacto", "alpha_gluco", "lipase", "CO2_flux_norm",
                    "T1_soil", "Soil_moist" ,"month" )

### 2024-09-11 data_analysis_sections/parametric_model_fitting.R

@@ -1,3 +1,5 @@
+
+
 library(parallel)
 
 # Set the number of cores to use
@@ -41,7 +43,7 @@ polygon(density(validation_data$CO2_flux_norm))
 
 
 # define the number of runs of the MCMC
-N_RUNS = 4000
+N_RUNS = 10000
 
 # Determine the number of cores available
 num_cores <- detectCores()
@@ -52,7 +54,7 @@ fit_bytreat_indA_Moy_sin <- stan(
   file = 'temp_moist_season_model.stan',
   data = stan_data,
   iter = N_RUNS,
-  chains = 10,
+  chains = 5,
   cores = num_cores-1,  # Use all available cores
   control = list(adapt_delta = 0.99, max_treedepth = 15)
 )
@@ -315,11 +317,13 @@ max(summary(fit_bytreat_indA_Moy_sin)$summary[,"Rhat"], na.rm = T)
 predicted_means_indA_Moy <- apply(post_bytreat_indA_Moy$model_resp, 2, mean)
 predicted_sds_indA_Moy <- apply(post_bytreat_indA_Moy$model_resp, 2, sd)
 
+length(predicted_means_indA_Moy)
+
 png("./Figures/fit_model.png", height = 2000, width = 2000, res=300)
-plot(predicted_means_indA_Moy, processed_data_filtered$CO2_flux_hour, ylab="observed", xlab="predicted",
+plot(predicted_means_indA_Moy, processed_data_filtered$CO2_flux_norm, ylab="observed", xlab="predicted",
      col=palette_treat[as.numeric(processed_data_filtered$treatment)],
      pch=as.numeric(processed_data_filtered$treatment), xlim=c(0,3), ylim=c(0,3))
-lm<-lm(processed_data_filtered$CO2_flux_hour ~ predicted_means_indA_Moy)
+lm<-lm(processed_data_filtered$CO2_flux_norm ~ predicted_means_indA_Moy)
 summary(lm)
 abline(a=0, b=1, col="black", lty=2)
 #legend("topleft", levels(processed_data$site), bty="n", pch=1:3)
@@ -414,10 +418,10 @@ modeled_validation <- temp_moist_season(temp = stan_data_valid$temp,
 
 png("./Figures/fit_model_validation.png", height = 2000, width = 2000, res=300)
 # the object range_plot comes from running the ML_model_benchmark.R file first
-plot(modeled_validation, validation_data$CO2_flux_hour, ylab="observed", xlab="predicted",
[…]

### 2024-09-11 functions_plotting.R

@@ -28,8 +28,9 @@ dev.off()
 
 png("./Figures/Season.png", width = 2000, height = 2000, res=300)
 
+amplitude=0.1183339
 day_year = seq(1:365)
-sine_wave =  0.1183339 * cos((2 * pi / 365) * day_year + (2 * pi/ 365) * (176.3293 - 1) - pi / 2);
+sine_wave =  amplitude * cos((2 * pi / 365) * day_year + (2 * pi/ 365) * (176.3293 - 1) - pi / 2);
 
 plot(day_year, sine_wave, type = "l", main = "Sine seasonality",
      xlab = "Day of the year", ylab = "Additive seasonality pattern", col = "darkorange")

### 2024-09-11 funtion_testing.R

@@ -1,97 +1,0 @@
-
-beta1 = 0.08;
-beta2 = 0.3;
-beta3 = 0.3;
-beta4 = 0.01;
-beta5 = 0.04;
-beta6 = 0.02;
-beta7 = 0.06;
-beta8 = 0.09;
-
-M=processed_data_filtered$Soil_moist[1]
-BD=1.2
-Clay=0.2
-SOC=0.03
-
-xi_moist  =  beta1*M + beta2*M^2 + beta3*M^3 + beta4*BD + beta5*M*BD + beta6*Clay + beta7*M*Clay + beta8*SOC
-
-Ea = 400
-logA=1
-
-model_resp_temp = exp(logA - Ea / (R * temp))
-
-temp=15
-Ea = 300
-T_0 =seq(from=210.13, to=220, by=0.1)
-A = 1
-model_resp_temp = A*exp(-Ea/((temp+273.15)-T_0))
-plot(T_0, model_resp_temp)
-
-
-
-k_b = 1.380649*10^-23 #kJ K^-1
-h=6.62607015*10^-34 #kJ S^-1
-R=8.314462618 # mol^-1 K^-1
-Dg_cat=51*10^3 #kJ
-DH_eq=92*10^3 #kJ
-T_eq=25+273 #K
-temp=seq(-20:100)+273 #Kß
-E_0=1
-
-K_cat = ((k_b*temp)/h)*exp(-(Dg_cat/(R*temp)));
-plot(temp-273, K_cat)
-
-K_eq = exp((DH_eq/R)*((1/T_eq)-(1/temp)));
-
-model_resp_temp = (K_cat * E_0)/(1+K_eq);
-
-plot(temp-273, model_resp_temp)
-plot(temp-273, K_cat)
-plot(temp-273, K_eq)
-
-
-
-
-Dg_cat=51*10^3 #kJ
-DH_eq=92*10^3 #kJ
-T_eq=25+273 #K
-temp=seq(-20:100)+273 #Kß
-E_0=1
-A=0.0001
-
-K_cat = ((k_b*temp)/h)*exp(-(Dg_cat/(R*temp)));
-K_eq = exp((DH_eq/R)*((1/T_eq)-(1/temp)));
-
-model_resp_temp = A* (K_cat * E_0)/(1+K_eq);
-
-plot(temp-273, model_resp_temp)
-
-
-# Constants
-h <- 6.62607015e-34  # Planck's constant, Joule*seconds
[…]

### 2024-09-11 master_data_analysis.R

@@ -25,7 +25,7 @@ set.seed(123)
 
 #NOTE: CUE and aciivation energy: in reality they relate, modelwise no, it's a shortcoming of having a model not distinguishing for C quality
 
-#TODO normalization of respiration by SOC when data are there
+#DONE normalization of respiration by SOC when data are there
 
 #TODO test variable E_a over time, remove seasonality scaling and use E_a by time
 
@@ -84,7 +84,40 @@ palette_plot <-  c(paletteer::paletteer_c("ggthemes::Orange-Gold", n=length(uniq
                    paletteer::paletteer_c("ggthemes::Green", n=length(unique(data[data$site=="romania",]$plot_ID))),
                    paletteer::paletteer_c("ggthemes::Blue-Teal", n=length(unique(data[data$site=="spain",]$plot_ID))))
 
+# palette_treat
+palette_treat <- c("#F4D166FF", "#F6A143FF", "#EE711DFF", "#CD4E22FF", "#9E3A26FF",
+                   "#B3E0A6FF", "#7FC572FF", "#5DA554FF", "#378748FF", "#24693DFF",
+                   "#BCE4D8FF", "#81C3CBFF", "#46A1B8FF", "#347C9FFF", "#2C5985FF")
 
+# palette_plot
+palette_plot <- c("#F4D166FF", "#F5CC63FF", "#F6C760FF", "#F7C25CFF", "#F8BD58FF",
+                  "#F8B855FF", "#F8B352FF", "#F8AE4FFF", "#F7A94BFF", "#F7A447FF",
+                  "#F69F42FF", "#F59B3EFF", "#F49639FF", "#F49235FF", "#F38D31FF",
+                  "#F3882CFF", "#F28228FF", "#F17D24FF", "#F07820FF", "#EF731DFF",
+                  "#EC6E1DFF", "#E96A1CFF", "#E6661DFF", "#E3631EFF", "#E05F1FFF",
+                  "#DD5C1FFF", "#D95820FF", "#D65421FF", "#D25122FF", "#CE4E22FF",
+                  "#C94C22FF", "#C54A22FF", "#C04823FF", "#BB4623FF", "#B64423FF",
+                  "#B14223FF", "#AD4024FF", "#A83E24FF", "#A33C25FF", "#9E3A26FF",
+                  "#B3E0A6FF", "#ACDE9EFF", "#A5DB96FF", "#9FD98FFF", "#99D688FF",
[…]

### 2024-09-11 temp_moist_season_model.R

@@ -14,6 +14,7 @@ temp_moist_season <- function(temp, M, day_year, a, b, Ea, A, peak_day, amplitud
     xi_temp[i] = A[plot_id[i]] * (exp(-Ea[treatment[i]] / ((temp[i] + 273.15) - T_0)));
     sine_wave[i] = amplitude[treatment[i]] * cos((2 * pi / 365) * day_year[i] + (2 * pi / 365) * (peak_day[treatment[i]] - 1) - pi / 2);
     model_resp[i] = sine_wave[i] +  (xi_temp[i] * xi_moist[i]);
+    #model_resp[i] = (xi_temp[i] * xi_moist[i]);
   }
 
 return(model_resp)

### 2024-09-11 testing.R

@@ -1,0 +1,31 @@
+
+dim(post_bytreat_indA_Moy$amplitude)
+dim(post_bytreat_indA_Moy$peak_day)
+Ea = mean(post_bytreat_indA_Moy$Ea)
+A = mean(post_bytreat_indA_Moy$A)
+a = mean(post_bytreat_indA_Moy$a)
+b = mean(post_bytreat_indA_Moy$b)
+T_0= 227.13
+
+moist = validation_data$Soil_moist
+temp =  validation_data$T1_soil
+
+xi_moist =   a * M - b * moist^2;
+xi_temp= A*(exp(-Ea / ((temp + 273.15) - T_0)));
+
+model_resp = (xi_temp * xi_moist);
+
+
+plot(model_resp)
+lines(validation_data$CO2_flux_norm, col="red")
+
+plot(model_resp, validation_data$CO2_flux_norm)
+abline(a=0, b=1, col="black", lty=2)
+
+
+
+
+plot(xi_moist)
+
+lines(validation_data$CO2_flux_hour, col="red")
+

### 2024-09-11 data_analysis_sections/parametric_model_fitting.R
Updated CC vulnerability plot

@@ -146,7 +146,7 @@ par(mfrow=c(1,2))
 E_0_densities_indA_Moy <- list()
 E_0_box_indA_Moy <- list()
 plot( density(post_bytreat_indA_Moy$Ea[,1]), xlim=c(range(post_bytreat_indA_Moy$Ea)[1]*0.9, range(post_bytreat_indA_Moy$Ea)[2]*1.1),
-      ylim=c(0,0.07), xlab = expression('E'[0]), col=NA, main="Temperature scaling")
+      ylim=c(0,0.07), xlab = expression('E'[a]), col=NA, main="")
 for(i in 1:stan_data$Tr){
   E_0_densities_indA_Moy[[i]] <- density(post_bytreat_indA_Moy$Ea[,i])
   E_0_box_indA_Moy[[i]] <- E_0_densities_indA_Moy[[i]]$x
@@ -160,7 +160,7 @@ for(i in 1:stan_data$Tr){
 A_densities_indA_Moy <- list()
 A_box_indA_Moy <- list()
 plot( density(post_bytreat_indA_Moy$A[,1]), xlim=c(range(post_bytreat_indA_Moy$A)[1]*0.9, range(post_bytreat_indA_Moy$A)[2]*1.1),
-      ylim=c(0, 0.03), xlab = expression('A'), col=NA, main="Temperature scaling")
+      ylim=c(0, 0.03), xlab = expression('A'), col=NA, main="")
 for(i in 1:stan_data$Pl){
   A_densities_indA_Moy[[i]] <- density(post_bytreat_indA_Moy$A[,i])
   A_box_indA_Moy[[i]] <- A_densities_indA_Moy[[i]]$x
@@ -179,7 +179,7 @@ par(mfrow=c(1,2))
 peak_densities_indA_Moy <- list()
 peak_box_indA_Moy <- list()
 plot( density(post_bytreat_indA_Moy$peak_day[,1]), xlim=c(range(post_bytreat_indA_Moy$peak_day)[1]*0.9, range(post_bytreat_indA_Moy$peak_day)[2]*1.1),
-      ylim=c(0,0.05), xlab = "peak_day", col=NA, main="Seasonality scaling")
+      ylim=c(0,0.05), xlab = "peak_day", col=NA, main="")
 for(i in 1:stan_data$Tr){
   peak_densities_indA_Moy[[i]] <- density(post_bytreat_indA_Moy$peak_day[,i])
[…]

### 2024-09-11 master_data_analysis.R

@@ -17,7 +17,7 @@ set.seed(123)
 
 #TODO NORMALIZE THE TEMPERATURE SCENARIOS!!
 
-#TODO: E_o and E_a are mixed up in the manuscript
+#DONE: E_o and E_a are mixed up in the manuscript
 
 #TODO plot SOC stocks by site
 
@@ -340,7 +340,7 @@ dev.off()
 ###########################################################################
 
 
-modeled_extrapolations <- temp_moist_season(temp = stan_data$temp+5,
+modeled_extrapolations <- temp_moist_season(temp = stan_data$temp+3,
                                         M = stan_data$M+0.05,
                                         day_year = stan_data$day_year,
                                         a = colMeans(post_bytreat_indA_Moy$a),
@@ -366,9 +366,15 @@ plot(modeled_extrapolations,
      modeled_zero)
 
 
-temp_effect = modeled_extrapolations -  modeled_zero
+
+temp_effect = (modeled_extrapolations -  modeled_zero)/modeled_zero
 png("./Figures/scenario_extrapolation.png", height=1800, width = 2000, res = 300)
 par(mar=c(12,4,2,2))
-bp <- boxplot(temp_effect ~ stan_data$treatment, names = names_treats, las=2, col = palette_treat_simplified, main = "Climate change vulnerability", ylab= "increase in mineralization", xlab="")
+bp <- boxplot(temp_effect*100 ~ stan_data$treatment, names = names_treats, las=2,
+              col = palette_treat_simplified,
+              main = expression("Vulnerability to +3" * degree * " C"),
+              ylab= "Increase in mineralization (% of current climate)",
+              xlab="",
+              ylim=c(0,150))
 add_shading_to_boxplot(bp, density = density_palette, angle = angle_palette)
 dev.off()

### 2024-09-11 Manuscript_draft/Draft.Rmd
Word document format for the draft

@@
 date: "2024-07-04"
-output:
-  bookdown::html_document2: default
-  bookdown::pdf_document2: default
+# output:
+#   bookdown::html_document2: default
+#   bookdown::pdf_document2: default
+output: word_document
   #output: pdf_document
@@
 ## Sites microclimate
-<!-- ![Climatic parameters measured at the sites, plotted as data density of all measurements.
-Please notice thot not every time series covers the same period of time.](../Figures/climatedata.png){width=60%} -->
 ![Climatic parameters and respirations measured at the sites (mean values after loess interpolation with span = 0.5).

### 2024-09-11 master_data_analysis.R

@@ -15,7 +15,7 @@ set.seed(123)
 
 #TODO Use basal area instead of DBH in the RF model
 
-#TODO NORMALIZE THE TEMPERATURE SCENARIOS!!
+#DONE NORMALIZE THE TEMPERATURE SCENARIOS!!
 
 #DONE: E_o and E_a are mixed up in the manuscript
 

### 2024-09-24 Manuscript_draft/Draft.tex
Added descriptie stats about enzyme and microbial data

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% definitions for citeproc citations
\NewDocumentCommand\citeproctext{}{}
\NewDocumentCommand\citeproc{mm}{%
  \begingroup\def\citeproctext{#2}\cite{#1}\endgroup}
\makeatletter
 % allow citations to break across lines
 \let\@cite@ofmt\@firstofone
 % avoid brackets around text for \cite:
 \def\@biblabel#1{}
 \def\@cite#1#2{{#1\if@tempswa , #2\fi}}
\makeatother
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-indent, #2 entry-spacing
 {\begin{list}{}{%
  \setlength{\itemindent}{0pt}
  \setlength{\leftmargin}{0pt}
  \setlength{\parsep}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
   \setlength{\leftmargin}{\cslhangindent}
   \setlength{\itemindent}{-1\cslhangindent}
  \fi
  % set entry spacing
  \setlength{\itemsep}{#2\baselineskip}}}
 {\end{list}}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{\hfill\break\parbox[t]{\linewidth}{\strut\ignorespaces#1\strut}}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{\strut#1\strut}}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\usepackage[utf8]{inputenc}
\usepackage{textcomp}
\usepackage{amsmath}
\DeclareUnicodeCharacter{03C9}{$\omega$}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={The impact of forest management on the temperature sensitivity of organic matter turnover},
  pdfauthor={Leticia Pérez-Izquiero, Lorenzo Menichetti, Jorge Curiel-Juste, etc},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{The impact of forest management on the temperature sensitivity of organic matter turnover}
\author{Leticia Pérez-Izquiero, Lorenzo Menichetti, Jorge Curiel-Juste, etc}
\date{2024-07-04}

\begin{document}
\maketitle

{
\setcounter{tocdepth}{2}
\tableofcontents
}
\section{Abstract}\label{abstract}

TEST CHANGE

\section{Introduction}\label{introduction}

The soil is one of the largest terrestrial reservoirs of carbon, containing approximately three times more carbon than the atmosphere or the vegetation on Earth (IPCC, 2021). Soil carbon is released to the atmosphere as carbon dioxide (CO₂) through decomposition of organic matter by microbes (heterotrophic respiration) and root respiration (autotrophic respiration). This process, i.e., soil respiration, constitutes the main carbon efflux into the atmosphere (ref). Under the new climate change scenarios increases of Co2 efflux from soils are expected (Schlesinger and Andrews, 2000), in turn this can imply a positive feedback to future warming (ref). Because soil respiration is highly sensitive to soil temperature and soil moisture (ref), understanding the responses of soil respiration to these microclimatic conditions, which are affected by global change factors as climate change and forest management is critical for predicting future atmospheric CO2 concentrations.

\subsection{Soil C turnover is controlled by temperature and moisture}\label{soil-c-turnover-is-controlled-by-temperature-and-moisture}

\subsection{Forest management changes the soil C turnover sensitivity to microclimate}\label{forest-management-changes-the-soil-c-turnover-sensitivity-to-microclimate}

\subsection{Climate change will influence soil C turnover globally}\label{climate-change-will-influence-soil-c-turnover-globally}

In this study, during three years we have monthly recorded data on soil respiration and microclimate in three replicated Quercus forests across Europe subjected to different forest management strategies. We have analysed the sensitivity of soil respiration to microclimate (i.e., soil temperature and soil moisture) by constructing a mechanistic model\ldots{}

\section{Material and methods}\label{material-and-methods}

\subsection{Study sites}\label{study-sites}

The study sites are located in Gamiz, Northern Spain (42°49'2.32''N; 2°37'10.67''W), Saint-Christol-d'Albion, South of France (xxx) and Dumbravita, East Romania (42°49'2.32''N; 2°37'10.67''W). According to Köppen-Geiger climatic classification, the climate in Gamiz is Oceanic, with cold and rainy winters and warm summers. In St.~Christol the climate is Mediterranean, with mild winters and warm and dry summers. In Dumbravita, the climate is humid continental with warm to hot summers, and cold and snowy winters with precipitation usually distributed throughout the year (Table 1). In the three study sites, the dominant vegetation are deciduous species of Quercus. Specifically, Quercus faginea Lam. in Gamiz, Quercus pubescens Willd. in St.~Christol and Quercus robur L. in Dumbravita (Table 1). In St.Christol and Dumbravita the stands are Quercus monodominant, while in Gamiz, Quercus faginea appears together with scattered Acer campestris L. and Fraxinus excelsior L., and an understory dominated by Crataegus monogyna Jacq., Juniperus communis L. and Viburnum lantana L. More characteristics of the study sites are found in Table 1.
The experimental sites have an approximate total area of 1.5 ha. In each site, three stands of 0.12 ha each and replicated in four blocks were delimited. The stands corresponded to a control and two canopy disturbance treatments, i.e., 50\% thinning and clear-cut (Fig. S1). The stands from disturbed treatments (thinning and clear-cut) were in turn divided in two: on half of the stands the slash resulting from the logging (leaves and twigs) was crashed and added on top of the soil and on the other half, the woody debris resulting from logging was removed. This made a total of 5 treatments (control, clear-cut with and without slash, thinning with and without slash) per block (Fig. S1). In each of the 5 treatments, two plots were established per block making a total of 40 plots (4 blocks x 5 treatments x 2 replicates per block; Fig S1). In turn, in each plot a transect of 2m length was established for monitoring (Fig. S1). Both the canopy disturbance and the slash addition treatments together with the fencing of the entire area were carried out in October 2021 in Gamiz, April 2022 in St.~Christol and March 2022 in Dumbravita.

\subsection{Soil respiration}\label{soil-respiration}

To measure soil CO2 efflux (i.e., soil respiration), three rings of 10 cm diameter and 8 cm high were put along the 2 m transect in each of the monitoring plots (Fig. S1). Respiration was measured by a mobile infrared gas analyser (EGM-5, PP Systems, Hitchin, UK) and an attached soil respiration chamber (SRC-1). Soil CO2 efflux (g CO2 m-2 h-1) was calculated as linear increase in CO2 concentrations within the chamber headspace. The total volume used for calculation was the volume of the chamber plus the volume of the specific ring. Measurements were done monthly in the 120 collars in each site. Measurement in Gamiz started in November 2021, in St.~Christol in February 2023 and in Dumbravita in May 2022.

\subsection{Forest stand characteristics}\label{forest-stand-characteristics}

Within each of the control and thinned stands (n=8), we recorded the number of trees to calculate tree density (No.~tree per ha) and measured tree diameter at breast height (DBH; 1.3 m) to calculate basal area.

\subsection{Microclimatic variables}\label{microclimatic-variables}

We measured soil moisture and temperature using TMS data loggers version 4 (Wild et al., 2019), located at the extreme of the 2 m transect in each of the monitoring plots (Fig S1). These probes have three temperature sensors positioned at +15 cm, 2 cm and -6 cm relative to the soil surface, and a sensor that measures volumetric soil moisture based on time domain transmission to a depth of 6 cm. To convert the raw TMS measurements of soil moisture to volumetric soil water content, we used a calibration curve developed by Kopecký et al.~(2021). For further modelling and since these sensors measure every 15 min, we extracted the soil temperature and soil moisture measurements closer to the time we measured soil respiration in the field.

\subsection{Soil physical-chemical analyses}\label{soil-physical-chemical-analyses}

Chemical analyses were measured on 5 soil cores of 4 cm diameter separated 0.5 m to a depth of 10 cm along the 2m monitoring transect. The five soil cores were mixed into a composite sample per plot. The composite sample was sieved through a 5-mm mesh sieve in the field, stored at -20°C and then freeze-dried prior to laboratory analyses. The pH was measured in distilled water 1:10 (w/v). The carbon (C) and nitrogen (N) content of the soil was measured using an elemental analyser: C content was measured by sulfochromic oxidation (ISO 14235), and N content was estimated by sulfuric acid mineralization with the addition of selenium and sodium sulphate and conversion to ammonium ions (ISO 11261), which were measured with the Segmented Flow Analyzer (SFA) Skalar. Available phosphorus (P) was determined after extraction with malachite green according to the protocol by Ohno \& Zibilske (1991). Bulk density was measured in each of the plots by collecting 3 soil cores of 6 cm diameter and 10 cm depth around the 2m transect that were mixed into a composite sample. Soil bulk density was determined by dividing the dry mass of the composite sample by the volume of the 3 cores, after removing the stones. Soil C stocks (g m2) were calculated by multiplying C concentration by bulk density and sampling depth.

\subsection{Microbial biomass}\label{microbial-biomass}

The same five soil cores used for chemical analyses were used to estimate soil microbial biomass. To do that, we extracted Phospholipid fatty acids (PLFAs) following Šnajdr et al.~(2008). Lipids were extracted using approximately 1g of dry soil with a mixture of chloroform:methanol:phosphate buffer (1:2:0.8) and separated by solid-phase extraction cartridges (HyperSep Silica SPE columns 200 mg/3 mL, Thermo Fisher Scientific). PLFAs were eluted using 2 mL of methanol. The selected fractions underwent to mild alkaline methanolysis and the methyl esters of PLFAs were analyzed by gas chromatography-mass spectrometry (GC-MS; 450-GC, 240-MS ion trap detector, Varian, Walnut Creek, CA). In the polar PLFA fraction, fungal biomass was quantified based on 18:2ω6,9 (fungal PLFA) content; bacterial biomass (total bacterial PLFA) was quantified as a sum of i14:0, i15:0, a15:0, 16:1ω5, 16:1ω7, 16:1ω9, 10Me-16:0, i16:0, i17:0, a17:0, cy17:0, 17:0, 10Me-17:0, 18:1ω7, 10Me-18:0, cy19:0 (Actinomycetes 10Me-16:0, 10Me-17:0, 10Me-18:0, Gram-positive i14:0, i15:0, a15:0, i16:0, i17:0, a17:0 and Gram-negative 16:1ω7, 16:1ω9, 18:1ω7, cy17:0, cy19:0). The sum of all of the selected PLFAs together with 16:0 and 18:1w9 was used as an estimate of total microbial biomass (total PLFA).

\subsection{Data pretreatment}\label{data-pretreatment}

\subsection{Validation}\label{validation}

In order to ensure a robust validation, all models have been validated with a split approach, keeping two completely independent data sets (80\% of the points for training and 20\% validation). Given the size of the dataset, we considered this approach more reliable than any crossvalidation approach.

\subsection{Parametric model}\label{parametric-model}

We defined a parametric model to consider all the three effects on respiration we hypothesized. The first is the influence of temperature, then the influence of soil moisture, and then a seasonal component which is assumed to be roughly proportional to the interactions of autotrophic respiration (which we hypothesized following a seasonal pattern).
Temperature and moisture scaling interact linearly, while seasonality was assumed additive:\\
\[
Rh(t) =  \xi_{temp(t,j, k)} \cdot \xi_{moist(t,j)} + \xi_{sin(t,j)}
\]

Where the index \(j\) corresponds to the treatments and the index \(k\) corresponds to the plots (minimal statistical unit), while \((t)\) denotes the variation over time of each forcing variable (temperature, moisture, or days of the year).

\subsection{Model calibration}\label{model-calibration}

The parametric model was calibrated in a stratified Bayesian framework, where we assigned independent parameters depending on different groping of the data.\\
The model was written in Stan (Carpenter et al. 2017; Stan Development Team 2018), run from its R interface (Stan Development Team 2024).
The likelihood function was defined as:\\
\[
    Rh_{meas} \sim \mathcal{N}(Rh_{sim}, \sigma)
\]
Where the parameter \(\sigma\) was defined as the variance of the whole dataset (0.27).

\subsubsection{Moisture}\label{moisture}

The moisture scaling function was taken from a simplified version of Moyano (2013) in the SoilR package (Carlos A. Sierra, Markus Mueller, and Susan E. Trumbore 2012).
\[
\xi_{moist(t,k)} = a_j \cdot \theta_{v}(t) - b_j \cdot (\theta_{v}(t))^2
\]
We assumed that the variability in moisture sensitivity is driven by treatments (\(j\)).
The prior distribution for the term \(a\) was assumed as \(\sim \mathcal{N}(3.11, 0.1)\) and for the term \(b\) was assumed as \(\sim \mathcal{N}(2.42, 0.1)\), utilizing as mean values the values in Carlos A. Sierra, Markus Mueller, and Susan E. Trumbore (2012) and setting up conservatives standard deviations.

\subsubsection{Temperature}\label{temperature}

The temperature scaling function we utilized to define the relationship is the one defined by Lloyd and Taylor (1994) and it is derived from the original Arrhenius equation {[}Arrhenius.1889{]}:
\[
\xi_{temp(t,j)} = A_{k} \cdot e^{\frac{-Ea_j}{(T(t)+273.15)-T_0)}}
\]
The two parameters \(A_{k}\) and \(E_a\) are represented by Bayesian priors.\\
For the temperature scaling we assumed that the activation energy term \(E_a\) (which defines roughly the slope of the relationship) is driven by treatments (\(j\)) while the linear rescaling term \(A_{k}\) is driven by the substrate at the single plot level (\(k\)). This because substrate (organic matter) amount is usually highly variable in soil even at small spatial scales, while its quality, which defines its temperature sensitivity, is assumed to follow the inputs quality which are relatively uniform over each treatments.\\
The effect of the rescaling term \(A_{k}\) is included in the function for temperature scaling to keep the convention of the original formulation of the equation, but it could be brought outside since it just a linear interaction.\\
The prior distribution for the term \(A\) was assumed as \(\sim \mathcal{N}(300, 100)\) and for the term \(E_a\) was assumed as \(\sim \mathcal{N}(398.5, 50)\), utilizing as mean values the values in Lloyd and Taylor (1994) and setting up conservatives standard deviations.

\subsubsection{Seasonality}\label{seasonality}

In order to correct for eventual seasonality patters, we introduce also a seasonality scaling.
The seasonality scaling functions assume a sine function with period = 365 days, with both amplitude \(ampl_j\) (the magnitude) and peak day \(peak_j\) (when the function is at its only annual peak) varying as a function of treatments:
\[
\begin{aligned}
    \xi_{sin(t,k)} = ampl_k \cdot sin \left(\frac{2 \cdot \pi}{365}\right) \cdot day(t) \\
    +\left(\frac{2 \cdot \pi}{365}\right) \cdot (peak_k - 1) - \frac{\pi}{2})
\end{aligned}
\]
Both the terms \(ampl_j\) and \(peak_j\) are described by Bayesian priors. Variance is by treatment.
We assumed that the variability in seaslonalityis driven by treatments (\(j\)).
The prior distribution for the term \(ampl\) was assumed as \(\sim \mathcal{N}(0, 0.5)\), so assuming as mean no seasonality effect, and for the term \(peak\) was assumed as \(\sim \mathcal{N}(175, 50)\), so assuming as mean for the peak the 24th of June (Midsummer day).

\subsection{Model benchmark}\label{model-benchmark}

The parametric model has been compared against two nonparametric machine learning models. The firs was a more conventional multiple linear regression model (default R package \texttt{lm}, R Core Team (2023)), and the second a random forest model (command \texttt{randomForest} from package \texttt{randomForest}, Liaw and Wiener (2021)). Both models had the structure as:\\
\[
Rh(t) \sim f(T(t), theta_v(t), treatment)
\]
The rationale behind these choices is to offer two extremes to compare the fitness of our parametric model with, assuming that a linear regression model will perform more poorly than the parametric model given its limited capabilities to represent relationships more complex than linear (as we assume temperature and moisture relationships with respiration), and that a random forest model will perform better than the parametric model given its plasticity. The random forest model was in this case fit with default parameters, without any other optimization.

\subsection{Model extrapolation of increased temperature effect on decomposition}\label{model-extrapolation-of-increased-temperature-effect-on-decomposition}

Once we calibrated our parametric model, we could test the consequences of an increase in temperature on decomposition by driving the model with the same dataset except temperature was increased by 5°C. This resulted in a more immediate way of visualizing the impact of a different temperature sensitivity of organic material.

\subsection{Machine learning variance decomposition}\label{machine-learning-variance-decomposition}

On a more limited subset (only Spain), we tested a more extensive random forest model in order to test the predictive power of other variables on the respiration. The model had structure as:\\
\[
\begin{aligned}
Rh(t) \sim f(trees, DBH, BD, por, pH, P, N_{tot}, C{tot}, biom_{fungi}, biom_{bact}, \\
biom_{act}, biom_{gram+}, biom_{gram-}, biom_{tot}, T(t), theta_v(t), treatment)
\end{aligned}
\]
We included in this model tree density in plants ha\(^{-1}\) (\(trees\)), average diameter at breast height (\(DBH\)), bulk density (\(BD\)), porosity (\(por\)), pH, phosophorous (\(P\)), total nitrogen (\(N_{tot}\)), total carbon (\(C_{tot}\)), fungal biomass (\(biom_{fungi}\)), bacterial biomass (\(biom_{bact}\)), actinomycetes biomass (\(biom_{act}\)), gram+ bacterial biomass (\(biom_{gram+}\)), gram- bacterial biomass (\(biom_{gram-}\)), total biomass (\(biom_{tot}\)), as well as variables mentioned above.\\
This model was tested only on a subset of the dataset (only the sites in Spain), for which all the additional variables were available.
The aim of this model was to test the predictive potential of any of these additional variables, in order to to explore eventual correlations and possible related processes. The choice of a random forest model over a more conventional linear model is motivated by the capacity of the former to consider more complex relationships and interactions, on top of the convenience of being able to assess quantitatively the predictive potential of each variable by estimating relative variable importance.

\section{Results}\label{results}

\subsection{Sites microclimate}\label{sites-microclimate}

\begin{figure}
\centering
\includegraphics[width=0.6\textwidth,height=\textheight]{../Figures/climatedata_means.png}
\caption{Climatic parameters and respirations measured at the sites (mean values after loess interpolation with span = 0.5). Please notice thot not every time series covers the same period of time. Red shades = France, green shades = Romania, blue shades = Spain}
\end{figure}

\subsection{Benchmark models and parametric model fit}\label{benchmark-models-and-parametric-model-fit}

\begin{figure}
\centering
\includegraphics[width=0.8\textwidth,height=\textheight]{../Figures/fit_ML_validation.png}
\caption{Fit of the two benchmark machine learning models, linear (a) and random forest (b)}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.6\textwidth,height=\textheight]{../Figures/fit_model_validation.png}
\caption{Fit of the parametric model}
\end{figure}

\subsection{Parametric model posteriors}\label{parametric-model-posteriors}

\begin{figure}
\centering
\includegraphics[width=0.6\textwidth,height=\textheight]{../Figures/posteriors_boxplots.png}
\caption{Posteriors of the parameters varying as a function of treatments}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.6\textwidth,height=\textheight]{../Figures/posteriors_boxplots_A.png}
\caption{Posteriors of the scaling parameter, varying as a function of single plots}
\end{figure}

\subsection{Machine learning variance decomposition}\label{machine-learning-variance-decomposition-1}

The fitness of the detailed random forest model (fitted only on data from Spain) was the highest of all the models we tested, unsurprisingly given the huge amount of predictors compared with all other models.\\
Soil temperature and moisture resulted the most important variables in predicting respiration from each site by far, followed by edaphic characteristics of the site (density and porosity), then by treatment, and then again by edaphic site characteristics (namely pH and total C). Only afterwards the first variable related to aboveground vegetation. Variables related to biomass, total and specific to microorganisms classes, come only later and did not seem particularly important.\\
\includegraphics[width=0.5\textwidth,height=\textheight]{../Figures/variance_decomposition.png}

\subsection{Model extrapolations}\label{model-extrapolations}

\begin{figure}
\centering
\includegraphics[width=0.6\textwidth,height=\textheight]{../Figures/scenario_extrapolation.png}
\caption{Sensitivity of the organic matter in different treatments to a 5°C increase in temperature, expressed as increase in mineralization from the baseline temperature}
\end{figure}

\section{Discussion}\label{discussion}

\subsection{Role of temperature and moisture in conditioning decomposition}\label{role-of-temperature-and-moisture-in-conditioning-decomposition}

Temperature and moisture seem the most important components in explaining the variance of the soil respiration. The seasonality scaling, although apparently differentiating slightly between sites, presents always very low amplitude posteriors, weighting on average only a 5\% of the total measured respiration. It was more important in France (between 4.5\% and 12.2\%) and much less in Spain and Romania (between 2.0\% and 3.2\%). This seasonality might represent many different processes, but the most likely are seasonal processes connected with the above-ground vegetation and their inputs to the soil, which in turn would control microbial communities.

\subsection{Activation energy of different organic material}\label{activation-energy-of-different-organic-material}

\subsection{Role of edaphic, ecological and microbiological variables in conditioning decomposition}\label{role-of-edaphic-ecological-and-microbiological-variables-in-conditioning-decomposition}

Any other variable than soil temperature and moisture did not seem to play a pivotal role in explaining the variance of the soil respiration.

\subsection{Consequences of different temperature sensitivity of organic material in reacting to warming}\label{consequences-of-different-temperature-sensitivity-of-organic-material-in-reacting-to-warming}

\section{Conclusions}\label{conclusions}

\section{References}\label{references}

\bibliography{references}

\section{Appendix}\label{appendix}

\begin{figure}
\centering
\includegraphics{../Figures/posteriors_temp.png}
\caption{Posteriors of the parameters of the temperature function, plotted as density distributions}
\end{figure}

\begin{figure}
\centering
\includegraphics{../Figures/posteriors_seasonality.png}
\caption{Posteriors of the parameters of the moisture function, plotted as density distributions}
\end{figure}

\begin{figure}
\centering
\includegraphics{../Figures/posteriors_moisture.png}
\caption{Posteriors of the parameters of the seasonality function, plotted as density distributions}
\end{figure}

\begin{figure}
\centering
\includegraphics{../Figures/Amplitude_ratio.png}
\caption{Ratio (expressed as \%) between the mean amplitude parameter and the mean measured respiration by treatment}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth,height=\textheight]{../Figures/Temp.png}
\caption{Temperature scaling function}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth,height=\textheight]{../Figures/Moist.png}
\caption{Moisture scaling function}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth,height=\textheight]{../Figures/Season.png}
\caption{Seasonality scaling function}
\end{figure}

\phantomsection\label{refs}
\begin{CSLReferences}{1}{0}
\bibitem[\citeproctext]{ref-SierraMueller}
Carlos A. Sierra, Markus Mueller, and Susan E. Trumbore. 2012. {``Models of Soil Organic Matter Decomposition: The SoilR Package, Version 1.0.''} \emph{Geoscientific Model Development} 5 (4): 1045--60. \url{https://doi.org/10.5194/gmd-5-1045-2012}.

\bibitem[\citeproctext]{ref-carpenter2017stan}
Carpenter, Bob, Andrew Gelman, Matthew D Hoffman, Daniel Lee, Ben Goodrich, Michael Betancourt, Marcus Brubaker, Jiqiang Guo, Peter Li, and Allen Riddell. 2017. {``Stan: A Probabilistic Programming Language.''} \emph{Journal of Statistical Software} 76 (1).

\bibitem[\citeproctext]{ref-randomForest_2021}
Liaw, Andy, and Matthew Wiener. 2021. \emph{randomForest: Breiman and Cutler's Random Forests for Classification and Regression}. \url{https://CRAN.R-project.org/package=randomForest}.

\bibitem[\citeproctext]{ref-Lloyd.1994l68}
Lloyd, J, and J A Taylor. 1994. {``{On the Temperature Dependence of Soil Respiration}.''} \emph{Functional Ecology} 8 (3): 315. \url{https://doi.org/10.2307/2389824}.

\bibitem[\citeproctext]{ref-R-Core-Team_2023}
R Core Team. 2023. \emph{{R}: A Language and Environment for Statistical Computing}. Vienna, Austria: R Foundation for Statistical Computing. \url{https://www.R-project.org/}.

\bibitem[\citeproctext]{ref-standev2018stancore}
Stan Development Team. 2018. {``{The Stan Core Library}.''} \href{http://mc-stan.org/\%2018}{http://mc-stan.org/ 18}.

\bibitem[\citeproctext]{ref-Rstan}
---------. 2024. {``{RStan}: The {R} Interface to {Stan}.''} \url{https://mc-stan.org/}.

\end{CSLReferences}

\end{document}

### 2024-09-24 Reports/report_17_Jun_2023.Rmd
Added descriptie stats about enzyme and microbial data

---
title: "Differences in activation energy of SOC from different treatments"
author: "Lorenzo Menichetti"
date: "2024-06-17"
output:
  beamer_presentation: default
  ioslides_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Scope of the work

This work was initiated on respiration time series from Spain, France and Bulgaria as a preliminary exploration of how to define and test a set of hypotheses connecting the data streams from field sampling campaigns in Holisoils.

It is ultimately a stratified variance decomposition. With parametric, process based models, Bayesian, but just a **variance decomposition** at the end.

## Variance over time
- The variance in observed respiration caused by **temperature** can be described by a **Arrhenius**-like relationship (LLoyd & Taylor)
- The variance in observed respiration caused by soil **moisture** can be described by a **bitonic** relationship (Moyano)
- Some residual variance can be described by a **seasonal** term ($\cos$)

## Between sites and between treatments variance
We hypothesized that an important component of the observed between sites and between treatments variance can be described by independent parameterizations of the scaling functions, specifically:

- The effect of initial SOC, so substrate availability, can be described by a site-specific intercept (A parameter),
- The effect of **different activation energies**, therefore different SOC qualities caused by treatments, can be described by treatment-specific "slopes" (Activation energy, $E_a$ parameter),  as well as seasonality and moisture interaction

## Model
Temperature and moisture scaling interact linearly, while seasonality is assumed additive. 
$$
Rh(t) = \left(A_{k} \cdot \xi_{temp(t,j)} \cdot \xi_{moist(t,j)} \right) + \xi_{sin(t,j)}
$$

Where the index $j$ corresponds to the treatments and the index $k$ corresponds to the plots (minimal statistical unit), while $(t)$ denotes the variation over time of each forcing variable (temperature, moisture, or days of the year)

## Temperature scaling term
The temperature scaling function assumes an independent activation energy $E_a$ varying by treatments $j$:
$$
\xi_{temp(t,j)} = e^ {\frac{-Ea_j}{(T(t)+273.15)-T_0)}}
$$
The two parameters $E_a$ and $T_0$ are represented by Bayesian priors.

## Temperature scaling term
```{r,  echo=FALSE,  message=FALSE, fig.height=5}

temp = seq(0,45)
xi_temp = exp(-311.9924/ ((temp+ 273.15) - 227.13));

plot(temp, xi_temp, type = "l", main = "Loyd-Taylor",
     xlab = "Temperature", ylab = "Activity")
```

## Moisture scaling term
The moisture scaling function had to be taken from a simplified version of Moyano et al. 2013 in the SoilR package. This was due to exectution time of the function (with more covariates) in Moyano et al. 2012, one or two orders of magnitudes above. I had to drop SOC, BD and Clay as covariates. Didn't lose almost anything in predictive power.
$$
\xi_{moist(t,k)}  =   a_j \cdot \theta_{v}(t) - b_j \cdot (\theta_{v}(t))^2
$$
Variance is by treatment.

## Moisture scaling term
```{r,  echo=FALSE,  message=FALSE, fig.height=5}
### From Sierra SoilR
fW.Moyano <- function(theta, a = 3.11, b = 2.42) {
  a * theta - b * theta^2
}

# Example usage
th <- seq(0, 1, 0.01)
xi <- fW.Moyano(theta = th)

plot(th, xi, type = "l", main = "Moyano 2013",
     xlab = "Volumetric soil water content (cm3 cm-3)", ylab = "Activity")
```

## Seasonality scaling term
The seasonality scaling functions assume a sine function with period = 365 days, with both amplitude $ampl_j$ (the magnitude) and peak day $peak_j$ (when the function is at its only annual peak) varying as a function of treatments:
$$
\begin{aligned}
    \xi_{sin(t,k)} = ampl_k \cdot sin \left(\frac{2 \cdot \pi}{365}\right) \cdot day(t) \\
    +\left(\frac{2 \cdot \pi}{365}\right) \cdot (peak_k - 1) - \frac{\pi}{2})
\end{aligned}
$$
Both the terms $ampl_j$ and $peak_j$ are described by Bayesian priors. Variance is by treatment.

## Seasonality scaling term
```{r,  echo=FALSE,  message=FALSE, fig.height=5}
day_year = seq(1:365)
sine_wave =  0.1183339 * cos((2 * pi / 365) * day_year + (2 * pi/ 365) * (176.3293 - 1) - pi / 2);

plot(day_year, sine_wave, type = "l", main = "Sine seasonality",
     xlab = "Day of the year", ylab = "Additive seasonality pattern")
```

## Bottom line: stratification
- The activation energy $E_a$ of the substrate vary with **treatments**
- The "intercept", explaining the variance due to SOC, varies with **plots**
- The two modifiers of the moisture scaling vary with **treatments**
- The parameters of the seasonal component vary with **treatments**

## Implementation (coarsely)
The model is written in Stan (https://mc-stan.org), which is an environment for statistical (Bayesian) modeling.

Once the model is written, Stan compiles it before running the sampler, so it is relatively good in terms of performances (it runs from R or Python or directly from Bash, in my case I use the R interface)

## Model fit: not too bad
```{r, echo=FALSE, out.width='70%'}
knitr::include_graphics("../fit_indA_Moy.png")
```

## ML Model fit (benchmark)

```{r, echo=FALSE, out.width='100%'}
knitr::include_graphics("../fit_rf.png")
```

Fitting two machine learning models (random forests and multiple linear regression) earns an even worse fit, 0.51 and 0.45 respectively.

## Posteriors and hypothesis testing: temperature
```{r, echo=FALSE, out.width='100%'}
knitr::include_graphics("../E0_indA_Moy.png")
```
Unsurprisingly, there are some significant **differences between the activation energies (or temperature sensitivity) of the treatments**. This might be the main line of our story.

## Posteriors and hypothesis testing: moisture
```{r, echo=FALSE, out.width='100%'}
knitr::include_graphics("../a_b_Moy.png")
```
The parameters of the moisture function also differed by treatment.

## Posteriors and hypothesis testing: seasonality
```{r, echo=FALSE, out.width='100%'}
knitr::include_graphics("../peak_amplitude_Moy.png")
```
Also the seasonality component differs by treatment.  
But adding the seasonality model gained information, form a R^2 of 0.55 to 0.56 so it is not particularly relevant.

## Residuals 
```{r, echo=FALSE, out.width='100%'}
knitr::include_graphics("../residuals_vs_climate.png")
```
Residuals seems distributed in a quite healthy way

## Residual variance to be explained?
I tried a RF model to see if there is some more information to use:
$$
\begin{aligned}
residuals \sim disturbance, density_{trees}, density_{quercus}, \\ 
              Mean DBH, porosity, BD, pH, phosphorous,\\
              Ntot, Ctot, biom_{fungi}, biom_{bact}, \\
              biom_{actinobac}, biom_{gram_pos}, biom_{gram_neg},\\
              biom_{tot}, fungi:bact
\end{aligned}
$$
There's not much. The fitness of the model is $R^2=0.06$.

## Next steps
- Testing for hysteretic phenomena? We might see if including the amplitude of the short term variation in soil moisture can improve the model

### 2024-09-24 bayesian_correction_MoySOilR.R

@@ -1,0 +1,1210 @@
+
+
+library(rstan)
+library(paletteer)
+library(prettyGraphs)
+library(randomForest)
+library(dplyr)
+library(lubridate)
+library(readxl)
+library(caret)
+library(MASS)
+library(parallel)
+
+set.seed(123)
+
+#TODO Use basal area instead of DBH in the RF model
+
+#TODO NORMALIZE THE TEMPERATURE SCENARIOS!!
+
+#TODO: E_o and E_a are mixed up in the manuscript
+
+#TODO plot SOC stocks by site
+
+#TODO: correlation. between E_a and enzyme activities
+
+#NOTE: CUE and aciivation energy: in reality they relate, modelwise no, it's a shortcoming of having a model not distinguishing for C quality
+
+#TODO normalization of respiration by SOC when data are there
+
+#TODO test variable E_a over time, remove seasonality scaling and use E_a by time
+
+#TODO Enzyme relationships
+
+#TODO convert volumetric water content to theta_s? When data come
+
+#DONE interpolated line of temperature by treatment (average)
+#DONE add the same graph for respiration
+#DONE try to make them side by side by site (3 plots)
+
+#DONE plot the reduction functions in appendix
+
+#DONE reorder the treatments with control first
+#DONE in the boxplots treatments control thinning clearcut with different shades, and slah with 45 degrees lines
+
+
+
+
+# edaphics <- read.csv("edaphics.csv")
+#data <- read.csv("Database_co2_28_05_2024.csv")
+data <- read_xlsx("Database_co2_03_09_2024_format_Lorenzo.xlsx", sheet=1)
+str(data)
+soil_data <- read_xlsx("Soil_data_Lorenzo_22_08_2024_enzymes.xlsx", sheet=1)
+
+
+data$treatment <- as.factor(data$treatment)
+data$site <- as.factor(data$site)
+data$Soil_moist <- as.numeric(data$Soil_moist)/100
+data$T1_soil <- as.numeric(data$T1_soil)
+data$CO2_flux_hour <- as.numeric(data$CO2_flux_hour)
+data$seq <- 1:dim(data)[1]
+
+soil_data
+
+data$Date <- as.Date(data$Date, format = "%d.%m.%Y")
+soil_data$date  <- as.Date(soil_data$date, format = "%d.%m.%Y")
+
+#rearrange the levels putting control first
[…]

### 2024-09-24 data_analysis_sections/conventional_variance_decomposition_enzymes.R

@@ -1,0 +1,79 @@
+
+library("multcompView")
+
+
+
+# Function to calculate confidence intervals using quantiles
+calc_ci <- function(data) {
+  lower <- quantile(data, probs = 0.025)  # 2.5% quantile for lower bound
+  upper <- quantile(data, probs = 0.975)  # 97.5% quantile for upper bound
+  return(c(lower, upper))
+}
+
+
+
+
+plot_names = colnames(soil_data)[30:44]
+
+enzyme_means_table = mat.or.vec( length(unique(interaction(soil_data$treatment, soil_data$site))), length(plot_names))
+colnames(enzyme_means_table) = plot_names
+rownames(enzyme_means_table) = levels(interaction(soil_data$treatment, soil_data$site))
+
+png("./Figures/Appendix/enzymes_boxplots.png", height=4000, width = 5000, res = 300)
+par(mar=c(14,5,1,1), mfrow=c(4,4))
+for (i in 1:length(plot_names)) {
+
+  name = plot_names[i]
+
+  boxplot(soil_data[[name]] ~ interaction(soil_data$treatment, soil_data$site), xlab="", ylab="",
+          col = palette_treat,
+          las = 2,
+          main = paste(name))
+
+  # Perform ANOVA
+  anova_model <- aov(soil_data[[name]] ~ interaction(soil_data$treatment, soil_data$site))
+
+  # Get the means and confidence intervals for each group
+  means_table <- aggregate(soil_data[[name]],
+                           by = list(interaction(soil_data$treatment, soil_data$site)),
+                           FUN = mean)
+  colnames(means_table) <- c("Group", "Mean")
+
+  # Apply quantile-based CI calculation for each group
+  ci_table <- aggregate(soil_data[[name]],
+                        by = list(interaction(soil_data$treatment, soil_data$site)),
+                        FUN = calc_ci)
+
+  # Extract lower and upper CIs using apply
+  lower_cis <- apply(ci_table$x, 1, function(row) row[1])  # Extract lower bound (2.5%)
+  upper_cis <- apply(ci_table$x, 1, function(row) row[2])  # Extract upper bound (97.5%)
+  ci_table <- data.frame(Group = ci_table$Group,
+                         Lower = lower_cis,
+                         Upper = upper_cis)
+
+  # Perform ANOVA again (to match grouping)
[…]

### 2024-09-24 data_analysis_sections/site_local_data.R

@@ -1,0 +1,60 @@
+
+png("./Figures/SOC.png", height=1800, width = 2000, res = 300)
+boxplot(soil_data$C_stocks ~ interaction(soil_data$site, soil_data$treatment),names = names_treats,
+        col = palette_treat_simplified,
+        main = "SOC content",
+        ylab = expression(SOC~(Mg~ha^-1)),
+        xlab="", las=2)
+dev.off()
+
+
+
+# define the number of runs of the MCMC
+N_RUNS = 5000
+
+# Determine the number of cores available
+num_cores <- detectCores()
+
+# fitting the model (in multicore mode)
+start <- Sys.time()
+fit_bytreat_indA_indEa_Moy_sin <- stan(
+  file = 'temp_moist_season_model_ind_Ea.stan',
+  data = stan_data,
+  iter = N_RUNS,
+  chains = 4,
+  cores = num_cores-1,  # Use all available cores
+  control = list(adapt_delta = 0.99, max_treedepth = 15)
+)
+end <- Sys.time()
+tot_time <- end - start
+tot_time
+
+post_bytreat_indA_indEa_Moy <- extract(fit_bytreat_indA_indEa_Moy_sin)
+
+Ea_vector<- colMeans(post_bytreat_indA_indEa_Moy$Ea)
+
+
+
+str(soil_data)
+# Calculate the mean of numeric columns by 'treatment', ignoring NAs
+mean_enzymes <- soil_data %>%
+  group_by(unique_ID) %>%  # Group by the factor (e.g., treatment)
+  summarise(across(where(is.numeric), ~ mean(.x, na.rm = TRUE)))  # Apply mean to numeric columns, ignoring NAs
+names(mean_enzymes)
+
+plot(mean_enzymes$CO2_flux_norm, Ea_vector)
+
+
+corr_vectors<-c()
+rsquared_vectors<-c()
+for(i in 1:(36-5)){
+  corr_vectors[i]<- cor(mean_enzymes[,6:36][,i], Ea_vector)
+  rsquared_vectors[i]<-summary(lm(unlist(mean_enzymes[,6:36][,i])~ Ea_vector))$r.squared
+}
+names(corr_vectors) = names(mean_enzymes[,6:36])
+names(rsquared_vectors) = names(mean_enzymes[,6:36])
+
+png("./Figures/Appendix/Ea_corr.png", height=1800, width = 3000, res = 300)
+par(mar=c(10,4,2,2))
+barplot(rsquared_vectors, las=2, ylab="R^2")
+dev.off()

### 2024-09-24 funtion_testing.R

@@ -1,0 +1,97 @@
+
+beta1 = 0.08;
+beta2 = 0.3;
+beta3 = 0.3;
+beta4 = 0.01;
+beta5 = 0.04;
+beta6 = 0.02;
+beta7 = 0.06;
+beta8 = 0.09;
+
+M=processed_data_filtered$Soil_moist[1]
+BD=1.2
+Clay=0.2
+SOC=0.03
+
+xi_moist  =  beta1*M + beta2*M^2 + beta3*M^3 + beta4*BD + beta5*M*BD + beta6*Clay + beta7*M*Clay + beta8*SOC
+
+Ea = 400
+logA=1
+
+model_resp_temp = exp(logA - Ea / (R * temp))
+
+temp=15
+Ea = 300
+T_0 =seq(from=210.13, to=220, by=0.1)
+A = 1
+model_resp_temp = A*exp(-Ea/((temp+273.15)-T_0))
+plot(T_0, model_resp_temp)
+
+
+
+k_b = 1.380649*10^-23 #kJ K^-1
+h=6.62607015*10^-34 #kJ S^-1
+R=8.314462618 # mol^-1 K^-1
+Dg_cat=51*10^3 #kJ
+DH_eq=92*10^3 #kJ
+T_eq=25+273 #K
+temp=seq(-20:100)+273 #Kß
+E_0=1
+
+K_cat = ((k_b*temp)/h)*exp(-(Dg_cat/(R*temp)));
+plot(temp-273, K_cat)
+
+K_eq = exp((DH_eq/R)*((1/T_eq)-(1/temp)));
+
+model_resp_temp = (K_cat * E_0)/(1+K_eq);
+
+plot(temp-273, model_resp_temp)
+plot(temp-273, K_cat)
+plot(temp-273, K_eq)
+
+
+
+
+Dg_cat=51*10^3 #kJ
+DH_eq=92*10^3 #kJ
+T_eq=25+273 #K
+temp=seq(-20:100)+273 #Kß
+E_0=1
+A=0.0001
+
+K_cat = ((k_b*temp)/h)*exp(-(Dg_cat/(R*temp)));
+K_eq = exp((DH_eq/R)*((1/T_eq)-(1/temp)));
+
+model_resp_temp = A* (K_cat * E_0)/(1+K_eq);
+
+plot(temp-273, model_resp_temp)
+
+
+# Constants
+h <- 6.62607015e-34  # Planck's constant, Joule*seconds
[…]

### 2024-09-24 master_data_analysis.R

@@ -20,6 +20,7 @@ set.seed(123)
 #DONE: E_o and E_a are mixed up in the manuscript
 
 #TODO plot SOC stocks by site
+#TODO plot the enzyme data by site (probably for supplement). Maybe ANOVA by site?
 
 #TODO: correlation. between E_a and enzyme activities
 
@@ -28,8 +29,6 @@ set.seed(123)
 #DONE normalization of respiration by SOC when data are there
 
 #TODO test variable E_a over time, remove seasonality scaling and use E_a by time
-
-#TODO Enzyme relationships
 
 #TODO convert volumetric water content to theta_s? When data come
 
@@ -180,6 +179,11 @@ names(soil_data)
 # ************************************************************
 source("./data_analysis_sections/ML_variance_decomposition_enzymes.R")
 
+
+# ************************************************************
+# **************  Enzyme variance decomposition  conventional ***************
+# ************************************************************
+source("./data_analysis_sections/conventional_variance_decomposition_enzymes.R")
 
 
 
@@ -340,7 +344,7 @@ dev.off()
 ###########################################################################
 
 
-modeled_extrapolations <- temp_moist_season(temp = stan_data$temp+3,
+modeled_extrapolations <- temp_moist_season(temp = stan_data$temp+5,
                                         M = stan_data$M+0.05,
                                         day_year = stan_data$day_year,
                                         a = colMeans(post_bytreat_indA_Moy$a),
@@ -372,9 +376,13 @@ png("./Figures/scenario_extrapolation.png", height=1800, width = 2000, res = 300
 par(mar=c(12,4,2,2))
 bp <- boxplot(temp_effect*100 ~ stan_data$treatment, names = names_treats, las=2,
               col = palette_treat_simplified,
-              main = expression("Vulnerability to +3" * degree * " C"),
+              main = expression("Vulnerability to +5" * degree * " C"),
               ylab= "Increase in mineralization (% of current climate)",
               xlab="",
-              ylim=c(0,150))
+              ylim=c(0,200))
 add_shading_to_boxplot(bp, density = density_palette, angle = angle_palette)
 dev.off()
+
+
+
+source("./data_analysis_sections/site_local_data.R")

### 2024-10-22 data_analysis_sections/parametric_model_fitting.R
Fixed an error in the Ea timevariant plot (colors and sites were in the wrong order)

@@ -17,6 +17,7 @@ range(processed_data_filtered$CO2_flux_hour)
 stan_data <- list(N = length(processed_data_filtered$CO2_flux_norm),
                   Tr = length(levels(processed_data_filtered$treatment)),
                   treatment = as.numeric(processed_data_filtered$treatment),
+                  treatment_txt = (processed_data_filtered$treatment),
                   resp = processed_data_filtered$CO2_flux_norm,
                   temp = processed_data_filtered$T1_soil,
                   M = processed_data_filtered$Soil_moist,
@@ -24,6 +25,8 @@ stan_data <- list(N = length(processed_data_filtered$CO2_flux_norm),
                   plot_id = as.integer(processed_data_filtered$plot_id),
                   day_year = yday(as.Date(processed_data_filtered$date))
 )
+
+str(stan_data)
 
 stan_data_valid <- list(N = length(validation_data$CO2_flux_norm),
                         Tr = length(levels(validation_data$treatment)),
@@ -43,7 +46,7 @@ polygon(density(validation_data$CO2_flux_norm))
 
 
 # define the number of runs of the MCMC
-N_RUNS = 10000
+N_RUNS = 7000
 
 # Determine the number of cores available
 num_cores <- detectCores()
@@ -157,6 +160,7 @@ for(i in 1:stan_data$Tr){
        levels(processed_data_filtered$treatment)[i], col=palette_treat[i], cex=0.8)
 }
 
+
 A_densities_indA_Moy <- list()
 A_box_indA_Moy <- list()
 plot( density(post_bytreat_indA_Moy$A[,1]), xlim=c(range(post_bytreat_indA_Moy$A)[1]*0.9, range(post_bytreat_indA_Moy$A)[2]*1.1),
@@ -172,6 +176,7 @@ for(i in 1:stan_data$Pl){
 }
 dev.off()
 
+barplot(rep(1, length(palette_treat)), col = palette_treat, border = NA, space = 0)
 
 
 png("./Figures/posteriors_seasonality.png", height = 2000, width = 4000, res=350)

### 2024-10-22 data_analysis_sections/parametric_model_fitting_time_variant.R

@@ -1,0 +1,258 @@
+
+
+library(parallel)
+library(lubridate)
+
+# Set the number of cores to use
+options(mc.cores = parallel::detectCores())
+
+interactions_years = droplevels(interaction(processed_data_filtered$treatment, (year(processed_data_filtered$date)-2021)))
+interactions_years_val = droplevels(interaction(validation_data$treatment, (year(validation_data$date)-2021)))
+
+# Data# Data# Data list for Stan
+stan_data <- list(N = length(processed_data_filtered$CO2_flux_norm),
+                  Tr = length(levels(processed_data_filtered$treatment)),
+                  treatment = as.numeric(processed_data_filtered$treatment),
+                  resp = processed_data_filtered$CO2_flux_norm,
+                  temp = processed_data_filtered$T1_soil,
+                  M = processed_data_filtered$Soil_moist,
+                  Pl = length(levels(processed_data_filtered$plot_id)),
+                  plot_id = as.integer(processed_data_filtered$plot_id),
+                  day_year = yday(as.Date(processed_data_filtered$date)),
+                  elapsed_y = year(processed_data_filtered$date)-2021,
+                  treat_year = as.numeric(interactions_years),
+                  treat_year_N = length(levels(interactions_years))
+)
+
+str(stan_data)
+
+stan_data_valid <- list(N = length(validation_data$CO2_flux_norm),
+                        Tr = length(levels(validation_data$treatment)),
+                        treatment = as.numeric(validation_data$treatment),
+                        resp = validation_data$CO2_flux_norm,
+                        temp = validation_data$T1_soil,
+                        M = validation_data$Soil_moist,
+                        Pl = length(levels(validation_data$plot_id)),
+                        plot_id = as.integer(validation_data$plot_id),
+                        day_year = yday(as.Date(validation_data$date)),
[…]

### 2024-10-22 master_data_analysis.R

@@ -13,33 +13,9 @@ library(parallel)
 
 set.seed(123)
 
-#TODO Use basal area instead of DBH in the RF model
+#TODO color code in the boxplot and Ea, they don't match. Compare and find the error
 
-#DONE NORMALIZE THE TEMPERATURE SCENARIOS!!
-
-#DONE: E_o and E_a are mixed up in the manuscript
-
-#TODO plot SOC stocks by site
-#TODO plot the enzyme data by site (probably for supplement). Maybe ANOVA by site?
-
-#TODO: correlation. between E_a and enzyme activities
-
-#NOTE: CUE and aciivation energy: in reality they relate, modelwise no, it's a shortcoming of having a model not distinguishing for C quality
-
-#DONE normalization of respiration by SOC when data are there
-
-#TODO test variable E_a over time, remove seasonality scaling and use E_a by time
-
-#TODO convert volumetric water content to theta_s? When data come
-
-#DONE interpolated line of temperature by treatment (average)
-#DONE add the same graph for respiration
-#DONE try to make them side by side by site (3 plots)
-
-#DONE plot the reduction functions in appendix
-
-#DONE reorder the treatments with control first
-#DONE in the boxplots treatments control thinning clearcut with different shades, and slah with 45 degrees lines
+#TODO Q10
 
 
 
@@ -47,7 +23,7 @@ set.seed(123)
 
 # edaphics <- read.csv("edaphics.csv")
 #data <- read.csv("Database_co2_28_05_2024.csv")
-data <- read_xlsx("Database_co2_03_09_2024_format_Lorenzo.xlsx", sheet=1)
+data <- read_xlsx("Database_co2_18_10_2024_format_Lorenzo.xlsx", sheet=1)
 str(data)
 soil_data <- read_xlsx("Soil_data_Lorenzo_03_09_2024_enzymes.xlsx", sheet=1)
 
@@ -67,6 +43,7 @@ soil_data$date  <- as.Date(soil_data$date, format = "%d.%m.%Y")
 #rearrange the levels putting control first
 data$treatment <- factor(data$treatment, levels = c("control", "clear_cut_no_slash", "clear_cut_slash", "thinning_no_slash", "thinning_slash"))
 
+str(data)
 
 
 #####################
@@ -117,7 +94,6 @@ palette_plot <- c("#F4D166FF", "#F5CC63FF", "#F6C760FF", "#F7C25CFF", "#F8BD58FF
[…]

### 2024-10-23 Manuscript_draft/Draft.Rmd
Added descriptie stats about enzyme and microbial data
Fixed an error in the Ea timevariant plot (colors and sites were in the wrong order)
Small update in figures

@@
 
-## Parametric model
+## Parametric model, constant `E_a`
 We defined a parametric model to consider all the three effects on respiration we hypothesized.
@@
 The prior distribution for the term $ampl$ was assumed as $\sim \mathcal{N}(0, 0.5)$, so assuming as mean no seasonality effect, and for the term $peak$ was assumed as $\sim \mathcal{N}(175, 50)$, so assuming as mean for the peak the 24th of June (Midsummer day).
+Amplitude can be negative (this is just reverting the order of peaks and valleys of the sine function).
 
@@
 
+## Parametric model, time-variant `E_a`
+In order to test the hypothesis of an eventual variation of the activation energy of the decomposing material over time without any assumption on the function of variation, we fitted a version of the model where the $E_a$ parameter was stratified in order to let it vary independently for each treatment and year of sampling.
+
+```{r, echo=FALSE}
+# Load necessary package
+library(knitr)
+
+# Create the data frame
+df <- data.frame(
+  Parameter = c("A", "Ea", "a", "b", "amplitude", "peak_day"),
+  Description = c(
+    "Pre-exponential factor (log)", 
+    "Activation energy (Ea/R)", 
+    "Scaling of Moyano function (moisture, linear term)", 
+    "Scaling of Moyano function (moisture, quadratic term)", 
+    "Amplitude of seasonal sine wave", 
+    "Peak day of seasonal sine wave"
+  ),
+  `Stratification Level, const Eₐ` = c(
+    "Plot", 
+    "Treatment", 
+    "Treatment", 
+    "Treatment", 
+    "Treatment", 
+    "Treatment"
+  ),
+  `Stratification Level, var Eₐ` = c(
+    "Plot", 
+    "Treatment-year", 
+    "Treatment", 
+    "Treatment", 
+    "Treatment", 
+    "Treatment"
+  ),
+    `Prior Values` = c(
+    "normal(30, 100)", 
+    "normal(398.5, 50)", 
+    "normal(3.11, 0.1)", 
+    "normal(2.42, 0.1)", 
+    "normal(0, 0.25)", 
+    "normal(175, 50)"
+    )
+)
+
+# Generate the table using kable
+kable(df, format = "pipe", caption = "Model parameters stratification and priors")
+```
+
 ## Machine learning variance decomposition
@@
 
+## Machine learning variance decomposition on enzymes
+
 # Results
@@
 
+## Edaphic and biological variables
+![Fit of the two benchmark machine learning models, linear (a) and random forest (b)](../Figures/Appendix/enzymes_boxplots.png){width=80%}
+
 ## Benchmark models and parametric model fit
-
 ![Fit of the two benchmark machine learning models, linear (a) and random forest (b)](../Figures/fit_ML_validation.png){width=80%}
 
-![Fit of the parametric model](../Figures/fit_model_validation.png){width=60%}
+![Fit of the parametric model,  constant `E_a`](../Figures/fit_model_validation.png){width=60%}
 
-## Parametric model posteriors
+## Parametric model posteriors,  constant `E_a`
 
@@
 
-![Posteriors of the scaling parameter, varying as a function of single plots](../Figures/posteriors_boxplots_A.png){width=60%}
+## Parametric model posteriors,  time-variant `E_a` 
+![Posteriors of the `E_a` parameter varying as a function of treatments and sampling year](../Figures/timeseries_Ea_timevariant.png){width=60%}
 
@@
 # Appendix
+![Posteriors of the scaling parameter, varying as a function of single plots](../Figures/posteriors_boxplots_A.png){width=60%}
 
@@
 ![Seasonality scaling function](../Figures/Season.png){width=50%}
+
+```{r, echo=FALSE}
+# Load necessary libraries
+library(knitr)
+
+# Read the CSV file
+anova_results <- read.csv("../Tables/Enzyme_means.csv", stringsAsFactors = FALSE)
+
+# Display the table with HTML superscripts
+kable(anova_results, escape = FALSE)
+```

### 2024-10-23 master_data_analysis.R

@@ -13,11 +13,9 @@ library(parallel)
 
 set.seed(123)
 
-#TODO color code in the boxplot and Ea, they don't match. Compare and find the error
-
 #TODO Q10
 
-
+#TODO plotting enzymatic activities and other variables (boxplots)
 
 
 

### 2024-10-23 Manuscript_draft/Draft.Rmd
Done requested updates, text is now updated with the figures

@@
 # Abstract
-TEST CHANGE
 
@@
 \begin{aligned}
-Rh(t) \sim f(trees, DBH, BD, por, pH, P, N_{tot}, C{tot}, biom_{fungi}, biom_{bact}, \\
-biom_{act}, biom_{gram+}, biom_{gram-}, biom_{tot}, T(t), theta_v(t), treatment)
+Rh(t) \sim f(trees,  treatment, DBH, BA, pH, BD, 
+P, N, C_{stocks}, \text{fungi\_biomass}, 
+biom_{fungi}, biom_{bact},\\
+biom_{act},biom_{gram+}, biom_{gram-}, 
+biom_{tot}, \frac{biom_{fungi}}{biom_{bact}}, 
+T_{soil}, M_{soil}, month)
 \end{aligned}
 $$
-We included in this model tree density in plants ha$^{-1}$ ($trees$), average diameter at breast height ($DBH$), bulk density ($BD$), porosity ($por$), pH, phosophorous ($P$), total nitrogen ($N_{tot}$), total carbon ($C_{tot}$), fungal biomass ($biom_{fungi}$), bacterial biomass ($biom_{bact}$), actinomycetes biomass ($biom_{act}$), gram+ bacterial biomass ($biom_{gram+}$), gram- bacterial biomass ($biom_{gram-}$), total biomass ($biom_{tot}$), as well as variables mentioned above.
+
+We included in this model tree density in plants ha$^{-1}$ ($trees$), average diameter at breast height ($DBH$), basal area ($BA$),  pH, bulk density ($BD$), phosophorous ($P$), total nitrogen ($N}$), carbon stocks ($C_{stocks$), fungal biomass ($biom_{fungi}$), bacterial biomass ($biom_{bact}$), actinomycetes biomass ($biom_{act}$), gram+ bacterial biomass ($biom_{gram+}$), gram- bacterial biomass ($biom_{gram-}$), total biomass ($biom_{tot}$), fungi:bacteria ratio, soil temperature ($T_{soil}$), soil moisture  ($M_{soil}$) and month of the year.
 This model was tested only on a subset of the dataset (only the sites in Spain), for which all the additional variables were available.
@@
 ## Machine learning variance decomposition on enzymes
+$$
+\begin{aligned}
+Rh(t) \sim f(trees, treatment, BA, BD, \\
+\beta_{\text{gluco}}, \text{acid\_phos}, \beta_{\text{xylo}}, \text{chitise}, \\
+\text{cellobiohydrolase}, \beta_{\text{galacto}}, \alpha_{\text{gluco}}, \text{lipase}, \\
+\text{CO}_2\text{\_flux\_norm}, T_1\text{\_soil}, \text{Soil\_moist}, \text{month})
+\end{aligned}
+$$
 
@@
 ## Edaphic and biological variables
-![Fit of the two benchmark machine learning models, linear (a) and random forest (b)](../Figures/Appendix/enzymes_boxplots.png){width=80%}
+![Fit of the two benchmark machine learning models, linear (a) and random forest (b)](../Figures/enzymes_boxplots.png){width=80%}
+![Fit of the two benchmark machine learning models, linear (a) and random forest (b)](../Figures/edaphic_boxplots.png){width=80%}
 
@@
 
-![Posteriors of the parameters varying as a function of treatments](../Figures/posteriors_boxplots.png){width=60%}
+![Posteriors of the parameters varying as a function of treatments, `E_a` constant over time](../Figures/posteriors_boxplots.png){width=60%}
 
@@
 
-## Machine learning variance decomposition: predicting the whole dataset based on ecological variables
-The fitness of the detailed random forest model (fitted only on data from Spain) was the highest of all the models we tested, unsurprisingly given the huge amount of predictors compared with all other models.
-Soil temperature and moisture resulted the most important variables in predicting respiration from each site by far, followed by edaphic characteristics of the site (density and porosity), then by treatment, and then again by edaphic site characteristics (namely pH and total C).
-Only afterwards the first variable related to aboveground vegetation.
-Variables related to biomass, total and specific to microorganisms classes, come only later and did not seem particularly important.
-![Fit of the detailed random forest model (a) and variable importance of the same model (b)](../Figures/variance_decomposition.png){width=50%}
+## Machine learning variance decomposition: predicting respiration based on edaphic, ecological and enzymatic data
 
-## Machine learning variance decomposition: predicting a substet based on soil enzymatic data
+![Fit of the detailed random forest model (a) and variable importance of the same model (b) for edaphics and ecological variables Only the 20 variables explaining most variance are shown here.](../Figures/variance_decomposition_edaphics.png){width=50%}
 
-![Fit of the detailed random forest model (a) and variable importance of the same model (b)](../Figures/variance_decomposition_enzyme.png){width=50%}
+![Fit of the detailed random forest model (a) and variable importance of the same model (b) for enzymatic activities.
+Only the 20 variables explaining most variance are shown here.](../Figures/variance_decomposition_enzyme.png){width=50%}
 
@@
 # Appendix
-![Posteriors of the scaling parameter, varying as a function of single plots](../Figures/posteriors_boxplots_A.png){width=60%}
+![Posteriors of the scaling parameter, varying as a function of single plots](../Figures/Appendix/posteriors_boxplots_A.png){width=60%}
 
-![Posteriors of the parameters of the temperature function, plotted as density distributions](../Figures/posteriors_temp.png)
+![Posteriors of the parameters of the temperature function, plotted as density distributions](../Figures/Appendix/posteriors_temp.png)
 
-![Posteriors of the parameters of the moisture function, plotted as density distributions](../Figures/posteriors_seasonality.png)
+![Posteriors of the parameters of the moisture function, plotted as density distributions](../Figures/Appendix/posteriors_seasonality.png)
 
-![Posteriors of the parameters of the seasonality function, plotted as density distributions](../Figures/posteriors_moisture.png)
+![Posteriors of the parameters of the seasonality function, plotted as density distributions](../Figures/Appendix/posteriors_moisture.png)
 
@@
 
-![Temperature scaling function](../Figures/Temp.png){width=50%}
+![Temperature scaling function](../Figures/Appendix/Temp_function.png){width=50%}
 
-![Moisture scaling function](../Figures/Moist.png){width=50%}
+![Moisture scaling function](../Figures/Appendix/Moist_function.png){width=50%}
 
-![Seasonality scaling function](../Figures/Season.png){width=50%}
+![Seasonality scaling function](../Figures/Appendix/Season_function.png){width=50%}
 
 ```{r, echo=FALSE}
-# Load necessary libraries
 library(knitr)
+anova_results <- read.csv("../Tables/Enzyme_means.csv", stringsAsFactors = FALSE)
+kable(anova_results, escape = FALSE, caption = "Table 1: ANOVA Results for enzyme activities")
+```
 
-# Read the CSV file
-anova_results <- read.csv("../Tables/Enzyme_means.csv", stringsAsFactors = FALSE)
-
-# Display the table with HTML superscripts
-kable(anova_results, escape = FALSE)
+```{r, echo=FALSE}
+library(knitr)
+anova_results_edaph <- read.csv("../Tables/Edaphics_means.csv", stringsAsFactors = FALSE)
+kable(anova_results_edaph, escape = FALSE, caption = "Table 1: ANOVA Results for edaphic and ecological variables")
 ```

### 2024-10-23 data_analysis_sections/ML_variance_decomposition_edaphics.R

@@ -1,0 +1,77 @@
+
+names(soil_data)
+predictor_list = c("No_trees_ha", "Mean_DBH", "Basal_area", "treatment" ,
+                   "pH","bulk_den", "phosphorous","Ntot","C_stocks","fungi_biomass",
+                   "bacteria_biomass","actinobacteria_biomass","gram_pos_biomass","gram_neg_biomass","total_biomass","fungi_bact_rate","CO2_flux_norm",
+                   "T1_soil", "Soil_moist" ,"month" )
+
+# Create stratification group
+soil_strat_group <- interaction(soil_data$treatment, soil_data$site)
+
+# Create stratified partition
+train_indices <- createDataPartition(soil_strat_group, p = 0.8, list = FALSE)
+
+# Split the data into training and validation sets
+soil_data_train <- soil_data[train_indices, ]
+soil_data_valid <- soil_data[-train_indices, ]
+
+rf_soil_data_train <- na.omit(soil_data_train[,predictor_list])
+rf_soil_data_valid <- na.omit(soil_data_valid[,predictor_list])
+
+rf_soil_data_train$treatment <- as.factor(rf_soil_data_train$treatment)
+rf_soil_data_valid$treatment <- as.factor(rf_soil_data_valid$treatment)
+
+
+dim(rf_soil_data_train)
+dim(rf_soil_data_valid)
+
+
+
+# Define the grid of hyperparameters to search
+tune_grid <- expand.grid(
+  mtry = c(3, 4, 5, 6) # Adjust based on the number of predictors
+)
+
+# Define cross-validation method
+control <- trainControl(
+  method = "cv",
+  number = 5,
+  search = "grid"
+)
+
+# Perform the tuning
+set.seed(123) # For reproducibility
+tuned_rf <- train(
+  CO2_flux_norm ~ .,
+  data = rf_soil_data_train,
+  method = "rf",
+  trControl = control,
+  tuneGrid = tune_grid,
+  ntree = 500 # or another fixed value
+)
+
+
+rf_model_CO2_flux <- randomForest(CO2_flux_norm ~ . , data = rf_soil_data_train,
+                                  mtry = tuned_rf$bestTune$mtry)
+
+range_plot = range(c(predict(rf_model_CO2_flux, newdata = rf_soil_data_valid),
+                     rf_soil_data_valid$CO2_flux_norm))
+
[…]

### 2024-10-23 data_analysis_sections/ML_variance_decomposition_enzymes.R

@@ -62,18 +62,6 @@ rf_model_CO2_flux <- randomForest(CO2_flux_norm ~ . , data = rf_soil_data_train,
 range_plot = range(c(predict(rf_model_CO2_flux, newdata = rf_soil_data_valid),
                      rf_soil_data_valid$CO2_flux_norm))
 
-plot(predict(rf_model_CO2_flux, newdata = rf_soil_data_valid), rf_soil_data_valid$CO2_flux_norm, ylab="predicted", xlab="observed", main="RF model",
-     col=palette_treat[as.numeric(rf_soil_data_valid$treatment)], pch=as.numeric(rf_soil_data_valid$treatment), xlim=range_plot, ylim=range_plot)
-lm<-lm(rf_soil_data_valid$CO2_flux_norm ~ predict(rf_model_CO2_flux, newdata = rf_soil_data_valid))
-summary(lm)
-abline(a=0, b=1, col="black", lty=2)
-legend("topright", levels(rf_soil_data_valid$treatment), bty="n", pch=1:15, col = palette_treat)
-legend("bottomright", paste("R^2 =", round(summary(lm)$r.squared,3)), bty="n", pch=NA, col = NA)
-
-importance(rf_model_CO2_flux)
-varImpPlot(rf_model_CO2_flux)
-
-
 png("./Figures/variance_decomposition_enzyme.png", height = 3000, width = 1500, res = 300)
 par(mfrow=c(2,1))
 # the object range_plot comes from running the ML_model_benchmark.R file first
@@ -87,7 +75,7 @@ legend("bottomright", paste("R^2 =", round(summary(lm)$r.squared,3)), bty="n", p
 
 varimp_model_CO2_flux <- varImp(rf_model_CO2_flux)
 varimp_model_CO2_flux[order(varimp_model_CO2_flux$Overall),]
-par(mar=c(1,10,2,2))
+par(mar=c(3,10,2,2))
 barplot(varimp_model_CO2_flux[order(varimp_model_CO2_flux$Overall),], las=2,
         names.arg = rownames(varimp_model_CO2_flux)[order(varimp_model_CO2_flux$Overall)], horiz = T)
 dev.off()

### 2024-10-23 data_analysis_sections/conventional_variance_decomposition_enzymes.R

@@ -1,7 +1,7 @@
 
 library("multcompView")
 
-
+predictor_list
 
 # Function to calculate confidence intervals using quantiles
 calc_ci <- function(data) {
@@ -14,24 +14,34 @@ calc_ci <- function(data) {
 
 
 plot_names = colnames(soil_data)[30:44]
+# reorder the plots
+interaction_levels <- with(soil_data, interaction(treatment, site))
+reordered_interaction <- factor(interaction_levels, levels = c("control.france","thinning_slash.france","thinning_no_slash.france",
+                                                               "clear_cut_slash.france","clear_cut_no_slash.france",
+                                                               "control.romania","thinning_slash.romania", "thinning_no_slash.romania",
+                                                               "clear_cut_slash.romania","clear_cut_no_slash.romania",
+                                                               "control.spain","thinning_slash.spain", "thinning_no_slash.spain",
+                                                               "clear_cut_slash.spain","clear_cut_no_slash.spain"
+))
 
 enzyme_means_table = mat.or.vec( length(unique(interaction(soil_data$treatment, soil_data$site))), length(plot_names))
 colnames(enzyme_means_table) = plot_names
-rownames(enzyme_means_table) = levels(interaction(soil_data$treatment, soil_data$site))
+rownames(enzyme_means_table) = levels(reordered_interaction)
 
-png("./Figures/Appendix/enzymes_boxplots.png", height=4000, width = 5000, res = 300)
+png("./Figures/enzymes_boxplots.png", height=4000, width = 5000, res = 300)
 par(mar=c(14,5,1,1), mfrow=c(4,4))
 for (i in 1:length(plot_names)) {
 
   name = plot_names[i]
 
-  boxplot(soil_data[[name]] ~ interaction(soil_data$treatment, soil_data$site), xlab="", ylab="",
+
+  boxplot(soil_data[[name]] ~ reordered_interaction, xlab="", ylab="",
           col = palette_treat,
           las = 2,
           main = paste(name))
 
   # Perform ANOVA
-  anova_model <- aov(soil_data[[name]] ~ interaction(soil_data$treatment, soil_data$site))
+  anova_model <- aov(soil_data[[name]] ~ reordered_interaction)
 
[…]

### 2024-10-23 data_analysis_sections/parametric_model_fitting.R

@@ -5,12 +5,8 @@ library(parallel)
 # Set the number of cores to use
 options(mc.cores = parallel::detectCores())
 
-# test a bit of the data
 
-plot(density(processed_data_filtered$CO2_flux_norm))
-polygon(density(processed_data_filtered$CO2_flux_hour))
-range(processed_data_filtered$CO2_flux_norm)
-range(processed_data_filtered$CO2_flux_hour)
+Q10_range = range(processed_data_filtered$T1_soil)
 
 
 # Data list for Stan
@@ -23,7 +19,8 @@ stan_data <- list(N = length(processed_data_filtered$CO2_flux_norm),
                   M = processed_data_filtered$Soil_moist,
                   Pl = length(levels(processed_data_filtered$plot_id)),
                   plot_id = as.integer(processed_data_filtered$plot_id),
-                  day_year = yday(as.Date(processed_data_filtered$date))
+                  day_year = yday(as.Date(processed_data_filtered$date)),
+                  Q10_range = Q10_range
 )
 
 str(stan_data)
@@ -36,7 +33,8 @@ stan_data_valid <- list(N = length(validation_data$CO2_flux_norm),
                         M = validation_data$Soil_moist,
                         Pl = length(levels(validation_data$plot_id)),
                         plot_id = as.integer(validation_data$plot_id),
-                        day_year = yday(as.Date(validation_data$date))
+                        day_year = yday(as.Date(validation_data$date)),
+                        Q10_range = Q10_range
 )
 
 #quick check of the data...
@@ -82,6 +80,8 @@ mean(post_bytreat_indA_Moy$Ea)
 mean(post_bytreat_indA_Moy$A)
 mean(post_bytreat_indA_Moy$xi_temp)
 mean(post_bytreat_indA_Moy$xi_moist)
+
+
 
 
 posteriors_summary <- data.frame(
@@ -144,7 +144,7 @@ write.csv(posteriors_summary, file="./Tables/posteriors_summary.csv")
 
 
 
-png("./Figures/posteriors_temp.png", height = 2000, width = 4000, res=350)
+png("./Figures/Appendix/posteriors_temp.png", height = 2000, width = 4000, res=350)
 par(mfrow=c(1,2))
 E_0_densities_indA_Moy <- list()
[…]

### 2024-10-23 data_analysis_sections/parametric_model_fitting_time_variant.R

@@ -143,7 +143,7 @@ write.csv(posteriors_summary_timevar, file="./Tables/posteriors_summary_timevar.
 
 treat_year_factor <- droplevels(interaction(processed_data_filtered$treatment, (year(processed_data_filtered$date)-2021)))
 
-png("./Figures/posteriors_temp_timevar.png", height = 2000, width = 4000, res=350)
+png("./Figures/Appendix/posteriors_temp_timevar.png", height = 2000, width = 4000, res=350)
 par(mfrow=c(1,2))
 E_0_densities_indA_Moy_timevar <- list()
 E_0_box_indA_Moy_timevar <- list()
@@ -161,7 +161,7 @@ for(i in 1:stan_data$treat_year_N){
 dev.off()
 
 
-png("./Figures/posteriors_boxplots_Ea_timevariant.png", height = 1500, width = 3500, res= 300)
+png("./Figures/Appendix/posteriors_boxplots_Ea_timevariant.png", height = 1500, width = 3500, res= 300)
 par(mar=c(8,3,1.5,0))
 bp <- boxplot(E_0_box_indA_Moy_timevar,
               names = levels(treat_year_factor), las=2, main = "A", cex.axis =0.5)

### 2024-10-23 data_analysis_sections/site_local_data.R

@@ -1,5 +1,6 @@
 
 png("./Figures/SOC.png", height=1800, width = 2000, res = 300)
+par(mar=c(10,5,1,1))
 boxplot(soil_data$C_stocks ~ interaction(soil_data$site, soil_data$treatment),names = names_treats,
         col = palette_treat_simplified,
         main = "SOC content",
@@ -9,30 +10,30 @@ dev.off()
 
 
 
-# define the number of runs of the MCMC
-N_RUNS = 5000
-
-# Determine the number of cores available
-num_cores <- detectCores()
-
-# fitting the model (in multicore mode)
-start <- Sys.time()
-fit_bytreat_indA_indEa_Moy_sin <- stan(
-  file = 'temp_moist_season_model_ind_Ea.stan',
-  data = stan_data,
-  iter = N_RUNS,
-  chains = 4,
-  cores = num_cores-1,  # Use all available cores
-  control = list(adapt_delta = 0.99, max_treedepth = 15)
-)
-end <- Sys.time()
-tot_time <- end - start
-tot_time
-
-post_bytreat_indA_indEa_Moy <- extract(fit_bytreat_indA_indEa_Moy_sin)
-
-Ea_vector<- colMeans(post_bytreat_indA_indEa_Moy$Ea)
-
+# # define the number of runs of the MCMC
+# N_RUNS = 5000
+#
+# # Determine the number of cores available
+# num_cores <- detectCores()
+#
+# # fitting the model (in multicore mode)
+# start <- Sys.time()
+# fit_bytreat_indA_indEa_Moy_sin <- stan(
+#   file = 'temp_moist_season_model_ind_Ea.stan',
+#   data = stan_data,
+#   iter = N_RUNS,
+#   chains = 4,
+#   cores = num_cores-1,  # Use all available cores
+#   control = list(adapt_delta = 0.99, max_treedepth = 15)
+# )
+# end <- Sys.time()
+# tot_time <- end - start
+# tot_time
+#
+# post_bytreat_indA_indEa_Moy <- extract(fit_bytreat_indA_indEa_Moy_sin)
+#
+# Ea_vector<- colMeans(post_bytreat_indA_indEa_Moy$Ea)
+#
 
 
 str(soil_data)

### 2024-10-23 data_analysis_sections/site_micrometeorology.R

@@ -1,7 +1,7 @@
 
 ########### site micrometeorology
 
-png("./Figures/climatedata.png", height=2500, width = 2000, res=300)
+png("./Figures/Appendix/climatedata.png", height=2500, width = 2000, res=300)
 par(mfrow=c(3,2), mar=c(2,5,1,1))
 
 data_meteo = data.frame(Date = processed_data_filtered_preprocess$date,
@@ -154,7 +154,7 @@ legend("topright", paste("Average moisture"), pch=NA, bty="n")
 
 plot(average_resp[average_resp$Treat == treats[1],]$DateNumeric, pch=NA, xlab = "Date", ylab = expression(Respiration ~ (g ~ C ~ kg^{-1})),
      average_resp[average_resp$Treat == treats[1],]$AverageResp, main = "", axes = FALSE,
-     xlim = range(extended_date_range), col = custom_palette_temp, ylim=c(0,0.85))
+     xlim = range(extended_date_range), col = custom_palette_temp, ylim=c(0,1))
 axis(1, at = seq(min(extended_date_range), max(extended_date_range), length.out = 10),
      labels = as.Date(seq(min(extended_date_range), max(extended_date_range), length.out = 10), origin = "1970-01-01"))
 axis(2)

### 2024-10-23 functions_plotting.R

@@ -1,6 +1,6 @@
 
 
-png("./Figures/Temp.png", width = 2000, height = 2000, res=300)
+png("./Figures/Appendix/Temp_function.png", width = 2000, height = 2000, res=300)
 temp = seq(0,45)
 xi_temp = exp(-311.9924/ ((temp+ 273.15) - 227.13));
 
@@ -10,7 +10,7 @@ plot(temp, xi_temp, type = "l", main = "Loyd-Taylor",
 dev.off()
 
 
-png("./Figures/Moist.png", width = 2000, height = 2000, res=300)
+png("./Figures/Appendix/Moist_function.png", width = 2000, height = 2000, res=300)
 
 ### From Sierra SoilR
 fW.Moyano <- function(theta, a = 3.11, b = 2.42) {
@@ -26,7 +26,7 @@ plot(th, xi, type = "l", main = "Moyano 2013",
 dev.off()
 
 
-png("./Figures/Season.png", width = 2000, height = 2000, res=300)
+png("./Figures/Appendix/Season_function.png", width = 2000, height = 2000, res=300)
 
 amplitude=0.1183339
 day_year = seq(1:365)

### 2024-10-23 master_data_analysis.R

@@ -208,6 +208,7 @@ source("./data_analysis_sections/ML_model_benchmarks.R")
 source("./data_analysis_sections/parametric_model_fitting_time_variant.R")
 source("./data_analysis_sections/parametric_model_fitting.R")
 
+source("./data_analysis_sections/Q10_plotting.R")
 
 
 ## Working on the residuals

### 2024-10-23 Manuscript_draft/Draft.Rmd
Q10 figure in the draft

@@
 # Appendix
+![Posteriors of the scaling parameter, varying as a function of single plots](../Figures/posteriors_boxplots_q10.png){width=60%}
+
 ![Posteriors of the scaling parameter, varying as a function of single plots](../Figures/Appendix/posteriors_boxplots_A.png){width=60%}

### 2024-10-23 Manuscript_draft/Draft.Rmd
new update of the database and updated figures. Leticia please check I am not making mistakes with the vectors, like wrong labels...

@@
 library(knitr)
+posteriors <- read.csv("../Tables/posteriors_summary.csv", stringsAsFactors = FALSE)
+kable(posteriors, escape = FALSE, caption = "Table 1: posterior probability distributions of the prameters, constant Ea")
+```
+
+```{r, echo=FALSE}
+library(knitr)
 anova_results <- read.csv("../Tables/Enzyme_means.csv", stringsAsFactors = FALSE)
-kable(anova_results, escape = FALSE, caption = "Table 1: ANOVA Results for enzyme activities")
+kable(anova_results, escape = FALSE, caption = "Table 2: ANOVA Results for enzyme activities")
 ```
@@
 anova_results_edaph <- read.csv("../Tables/Edaphics_means.csv", stringsAsFactors = FALSE)
-kable(anova_results_edaph, escape = FALSE, caption = "Table 1: ANOVA Results for edaphic and ecological variables")
+kable(anova_results_edaph, escape = FALSE, caption = "Table3: ANOVA Results for edaphic and ecological variables")
 ```

### 2024-10-23 data_analysis_sections/parametric_model_fitting.R

@@ -5,9 +5,22 @@ library(parallel)
 # Set the number of cores to use
 options(mc.cores = parallel::detectCores())
 
-
+#Q10 range, the two temperatures where to calculate Q10
 Q10_range = range(processed_data_filtered$T1_soil)
 
+#reordering the treatments for plotting uniformity
+processed_data_filtered$treatment <- factor(processed_data_filtered$treatment, levels = c("control.france","thinning_slash.france","thinning_no_slash.france",
+                                                               "clear_cut_slash.france","clear_cut_no_slash.france",
+                                                               "control.romania","thinning_slash.romania", "thinning_no_slash.romania",
+                                                               "clear_cut_slash.romania","clear_cut_no_slash.romania",
+                                                               "control.spain","thinning_slash.spain", "thinning_no_slash.spain",
+                                                               "clear_cut_slash.spain","clear_cut_no_slash.spain"))
+validation_data$treatment <- factor(validation_data$treatment, levels = c("control.france","thinning_slash.france","thinning_no_slash.france",
+                                                                                          "clear_cut_slash.france","clear_cut_no_slash.france",
+                                                                                          "control.romania","thinning_slash.romania", "thinning_no_slash.romania",
+                                                                                          "clear_cut_slash.romania","clear_cut_no_slash.romania",
+                                                                                          "control.spain","thinning_slash.spain", "thinning_no_slash.spain",
+                                                                                          "clear_cut_slash.spain","clear_cut_no_slash.spain"))
 
 # Data list for Stan
 stan_data <- list(N = length(processed_data_filtered$CO2_flux_norm),
@@ -246,13 +259,13 @@ dev.off()
 
 levels(processed_data_filtered$treatment)
 
-names_treats <- c( "control France", "CC no-slash France", "CC slash France", " thin no-slash France", " thin slash France",
-                   "control Romania", "CC no-slash Romania", "CC slash Romania", " thin no-slash Romania"," thin slash Romania",
-                   "control Spain", "CC no-slash Spain", "CC slash Spain", " thin no-slash Spain", " thin slash Spain")
+names_treats <- c( "control France", "thin slash France"," thin no-slash France",  "CC slash France", "CC no-slash France",
[…]

### 2024-10-24 Manuscript_draft/Draft.Rmd
Inlclusion of enzymes in M&M

@@
 
-## Soil physical-chemical analyses
-Chemical analyses were measured on 5 soil cores of 4 cm diameter separated 0.5 m to a depth of 10 cm along the 2m monitoring transect.
+## Soil physical-chemical and enzymatic analyses
+Chemical analyses and enzymes were measured on 5 soil cores of 4 cm diameter separated 0.5 m to a depth of 10 cm along the 2m monitoring transect.
 The five soil cores were mixed into a composite sample per plot.
@@
 Soil C stocks (g m2) were calculated by multiplying C concentration by bulk density and sampling depth.
+The activity of extracellular enzymes involved in SOM decomposition was measured through enzyme assays developed by Štursová & Baldrian (2011).
+Briefly, the activity of cellobiohydrolase, β-galactosidase, β-glucosidase, β-xylosidase, lipase, chitinase, α-glucosidase and acid-phosphatase  were measured at pH 5.0 in 1:100 (w/v) sample slurries using methylumbelliferol-based substrates, with an excitation wavelength of 355 nm and an emission wavelength of 460 nm.
+Product quantification was based on standard curves with a range of 4-methylumbelliferol concentrations in the same sample slurry.
 

### 2024-10-24 Manuscript_draft/Draft.Rmd
Minor edits on the text

@@
 
-Here, we hypothesized that logging of trees will decrease the respiration rates because of the strong decrease in vegetation activity and microbiota associated.
+Here, we hypothesized that:
+(1) logging of trees will decrease the respiration rates because of the strong decrease in vegetation activity and microbiota associated.
 In turn, addition of slash will provide labile organic matter to decompose increasing the respiration rates.
-We also hypothesized that the variation in soil respiration would be mainly dependent on microclimatic variables (soil temperature and moisture), especially in the management treatments where vegetation was removed, since heterotrophic respiration will prevail over autotrophic respiration.
-Moreover, the sensitivity of soil respiration to microclimatic conditions will depend on the type of climate/forest, being the Mediterranean the most sensitive as the microbial communities will be better adapted to the strong seasonal changes in water availability.
+(2) the variation in soil respiration would be mainly dependent on microclimatic variables (soil temperature and moisture), especially in the management treatments where vegetation was removed, since heterotrophic respiration will prevail over autotrophic respiration, and where the litter layer (slash) was removed
+(3) the sensitivity of soil respiration to microclimatic conditions will also depend on the type of climate/forest, being the Mediterranean the less sensitive as the microbial communities will be better adapted to the strong seasonal changes in water availability.
+(4) the sensitivity of respiration to microclimate will decrease as the time passes since the disturbance because of new fresh C input from the growth of understory vegetation
 
-To test these hypotheses, we have monthly recorded data on soil respiration and microclimate in three replicated Quercus forests across Europe subjected to different forest management strategies during three years.
-We have analysed the sensitivity of soil respiration to microclimate (i.e., soil temperature and soil moisture) by constructing a mechanistic model…
+To test these hypotheses, we have recorded data on soil respiration and microclimate every month in three replicated Quercus forests across Europe subjected to different forest management strategies during three years.
+We have analysed the sensitivity of soil respiration to microclimate (i.e., soil temperature and soil moisture) by constructing a mechanistic parametric model and compared to a non-parametric model (i.e., random forest).
 
@@
 In order to ensure a robust validation, all models have been validated with a split approach, keeping two completely independent data sets (80% of the points for training and 20% validation).
-Given the size of the dataset, we considered this approach more reliable than any crossvalidation approach.
+Given the size of the dataset, we considered this approach more reliable than any cross-validation approach.
 
@@
 For the temperature scaling we assumed that the activation energy term $E_a$ (which defines roughly the slope of the relationship) is driven by treatments ($j$) while the linear rescaling term $A_{k}$ is driven by the substrate at the single plot level ($k$).
-This because substrate (organic matter) amount is usually highly variable in soil even at small spatial scales, while its quality, which defines its temperature sensitivity, is assumed to follow the inputs quality which are relatively uniform over each treatments.
+This is because substrate (organic matter) amount is usually highly variable in soil even at small spatial scales, while its quality, which defines its temperature sensitivity, is assumed to follow the inputs quality which are relatively uniform over each treatments.
 The effect of the rescaling term $A_{k}$ is included in the function for temperature scaling to keep the convention of the original formulation of the equation, but it could be brought outside since it just a linear interaction.
@@
 Both the terms $ampl_j$ and $peak_j$ are described by Bayesian priors. Variance is by treatment.
-We assumed that the variability in seaslonalityis driven by treatments ($j$). 
+We assumed that the variability in seasonality is driven by treatments ($j$). 
 The prior distribution for the term $ampl$ was assumed as $\sim \mathcal{N}(0, 0.5)$, so assuming as mean no seasonality effect, and for the term $peak$ was assumed as $\sim \mathcal{N}(175, 50)$, so assuming as mean for the peak the 24th of June (Midsummer day).

### 2024-10-25 Manuscript_draft/Draft.Rmd
Caption figures

@@
 ## Sites microclimate
-![Climatic parameters and respirations measured at the sites (mean values after loess interpolation with span = 0.5).
+![Climatic parameters (soil temperature and soil moisture) and respiration measured at the study sites (mean values per month and treatment after loess interpolation with span = 0.5).
 Please notice thot not every time series covers the same period of time.
@@
 ## Edaphic and biological variables
-![Fit of the two benchmark machine learning models, linear (a) and random forest (b)](../Figures/enzymes_boxplots.png){width=80%}
-![Fit of the two benchmark machine learning models, linear (a) and random forest (b)](../Figures/edaphic_boxplots.png){width=80%}
+![Boxplots of soil microbial biomass and enzymatic activities per disturbance treatment and study site.
+Red shades = France, green shades = Romania, blue shades = Spain](../Figures/enzymes_boxplots.png){width=80%}
+![Boxplots of stand characteristics and soil physical-chemical properties per disturbance treatment and study site.
+Red shades = France, green shades = Romania, blue shades = Spain](../Figures/edaphic_boxplots.png){width=80%}
 

### 2024-11-04 Manuscript_draft/Draft.Rmd
Results on Ea and Q10

@@
 ## Parametric model posteriors, constant `E_a` 
+The activation energy (Ea) and temperature sensitivity (Q10) values of SOM mineralization were consistently higher in Romania, followed by France and Spain.
+In general, the activation energy and Q10 values decreased with slash addition due to higher SOM quality compared with treatments in which slash was removed.
+This was especially evident in Spain for both thinning and clear-cut and in the clear treatment in France.
+The activation energy and Q10 in the treatments where slash was added did not differ from controls (Figure x, xa).
+These results were consistent when the activation energy was analysed as a function of time since disturbance and the activation energy systematically decreased in all treatments and all countries with time since disturbance (Figure x).
+The photosynthetic activity did not modify the sensitivity of soil respiration to temperature (peak and amplitude factors from the sine-function not affected) (Figure xb, c).
+However, the amplitude of the sine-function tended to be higher in France (Figure xc).
+The parameters of soil moisture function (a and b) did not change the response of the model either (Figure xd, e).
 

### 2024-11-05 Manuscript_draft/Draft.Rmd
Results section

@@
 
-## Sites microclimate
+## Characterization of study sites
+Tree density in France was very high compared to Spain and especially to Romania where the tree density was very low (Figure x).
+In contrast, the DBH was the lowest in France and the highest in Romania (Figure x).
+Spain showed intermediate values of both tree density and DBH (Figure x).
+For soil variables, in the untouched forest (control), the bulk density was the highest in Spain and the lowest in France.
+Spain had the highest pH while France and Romania had similar ranges of pH.
+The highest significant levels of C and N were found in Spain while the lowest C content was found in Romania.
+France in turn presented doubled C:N rates than the other two sites (Figure x).
+Total microbial biomass, fungal biomass and bacterial biomass peaked in France, followed by Spain and it was the lowest in Romania.
+Similarly, in general, activities of most enzymes were higher in France compared with Spain and Romania (Figure x).
+
+##Effects of management disturbance treatments on soil physical-chemical and biological properties
+The disturbance treatments barely affected the soil properties in Romania (Figure x).
+In France, the bulk density increased as the intensity of the disturbance did (i.e., from thinning with slash to clear-cut no slash) and in Spain it significantly increased in all the disturbance treatments compared with control (Figure x).
+The pH only significantly increased in both clear-cut treatments in Spain.
+Moreover, in Spain the N content decreased in the thinning and clear-cut without slash (Figure x).
+In France, compared with controls, C content increased in all the disturbance treatment, but in the clear-cut without slash that decreased (Figure x).
+In Spain, C content only decreased in the thinning-without slash treatment (Figure x).
+Similar to soil properties, microbial biomass, fungal biomass and bacterial biomass did not change with the disturbance treatment in Romania.
+In Spain, these variables significantly decreased in the disturbance treatments where the slash was removed and in France only in the clear-cut treatment where the slash was removed (Figure x).
+The disturbance treatments barely affected enzymatic activities in Romania and Spain, but generally decreased in clear-cut where slash was removed in France (Figure x).
+
+![Boxplots of soil microbial biomass and enzymatic activities per disturbance treatment and study site.
+Red shades = France, green shades = Romania, blue shades = Spain](../Figures/enzymes_boxplots.png){width=80%}
+![Boxplots of stand characteristics and soil physical-chemical properties per disturbance treatment and study site.
+Red shades = France, green shades = Romania, blue shades = Spain](../Figures/edaphic_boxplots.png){width=80%}
+
+## Effects of management disturbance treatments on soil microclimate and respiration
+
 ![Climatic parameters (soil temperature and soil moisture) and respiration measured at the study sites (mean values per month and treatment after loess interpolation with span = 0.5).
@@
 
-## Edaphic and biological variables
-![Boxplots of soil microbial biomass and enzymatic activities per disturbance treatment and study site.
-Red shades = France, green shades = Romania, blue shades = Spain](../Figures/enzymes_boxplots.png){width=80%}
-![Boxplots of stand characteristics and soil physical-chemical properties per disturbance treatment and study site.
-Red shades = France, green shades = Romania, blue shades = Spain](../Figures/edaphic_boxplots.png){width=80%}
+As expected, in Spain the temperature variation was the lowest compared to the other two sites.
+In general, control plots showed the lowest thermal amplitude in all sites (i.e., with higher temperatures in winter and lower in summer).
+In Spain, the soil temperature increased as the intensity of the tree removal, mainly in summer, but the addition of slash buffered these increases.
+In France, the main effect on soil temperature was the tree removal, especially in summer and this was independent on the addition of slash.
+Soil temperature was not affected by disturbance and post-disturbance treatments in Romania.
+The observed effects of the disturbance and post-disturbance treatments on soil temperature were decreasing with time.
+Soil moisture fluctuated inversely with temperature.
+In Spain, soil moisture was higher in the clear-cut with and without slash addition, except in the driest periods (from April to August).
+Similarly, in Romania, clear-cut increased soil moisture independently of slash addition.
+Contrarily, in France, soil moisture tended to decrease in the clear-cut treatment with and without slash (Figure x).
+Both in Spain and France, removal of slash decreased respiration rates, independently of tree cutting intensity, and thinning coupled with slash addition kept the respiration rates as the same level as controls.
+In Romania, effects of the disturbances were lower, however clear-cut increased respiration rates, especially when slash was added to the soil.
+At the end of the monitoring period, three years for Spain and two years for France and Romania after disturbance, respiration rates increased in the clear-cut with slash addition in the three study sites (Figure x).
 

### 2024-11-06 data_analysis_sections/ML_variance_decomposition_edaphics.R
Fixed figures (had just to re-run)

@@ -22,6 +22,11 @@ rf_soil_data_train$treatment <- as.factor(rf_soil_data_train$treatment)
 rf_soil_data_valid$treatment <- as.factor(rf_soil_data_valid$treatment)
 
 
+rf_soil_data_train_T1 <- na.omit(soil_data_train[soil_data_train$time == "T1",predictor_list])
+rf_soil_data_valid_T1 <- na.omit(soil_data_valid[soil_data_valid$time == "T1",predictor_list])
+
+
+
 dim(rf_soil_data_train)
 dim(rf_soil_data_valid)
 
@@ -54,6 +59,17 @@ tuned_rf <- train(
 rf_model_CO2_flux <- randomForest(CO2_flux_norm ~ . , data = rf_soil_data_train,
                                   mtry = tuned_rf$bestTune$mtry)
 
+
+rf_model_CO2_flux_T1 <- randomForest(CO2_flux_norm ~ . , data = rf_soil_data_train_T1,
+                                  mtry = tuned_rf$bestTune$mtry)
+varImpPlot(rf_model_CO2_flux_T1)
+
+plot(predict(rf_model_CO2_flux_T1, newdata = rf_soil_data_valid_T1), rf_soil_data_valid_T1$CO2_flux_norm)
+plot(predict(rf_model_CO2_flux_T1), rf_soil_data_train_T1$CO2_flux_norm)
+summary(lm(predict(rf_model_CO2_flux_T1) ~ rf_soil_data_train_T1$CO2_flux_norm))
+
+
+
 range_plot = range(c(predict(rf_model_CO2_flux, newdata = rf_soil_data_valid),
                      rf_soil_data_valid$CO2_flux_norm))
 
@@ -74,4 +90,6 @@ par(mar=c(3,10,2,2))
 barplot(varimp_model_CO2_flux[order(varimp_model_CO2_flux$Overall),], las=2,
         names.arg = rownames(varimp_model_CO2_flux)[order(varimp_model_CO2_flux$Overall)], horiz = T)
 dev.off()
+
+
 

### 2024-11-06 data_analysis_sections/Q10_explanatory_hypotheses.R

@@ -1,0 +1,67 @@
+
+
+
+
+
+#the levels at which we need to collapse the datasets:
+soil_data_aggregation_level = droplevels(interaction(soil_data$plot_ID, soil_data$site))
+processed_data_filtered$plot_id
+
+
+dim(post_bytreat_indA_Moy$q10)
+
+# means of Q10
+Q10_averages <- tapply(colMeans(post_bytreat_indA_Moy$q10), processed_data_filtered$plot_id, mean, na.rm = TRUE)
+
+
+# means of all other variables
+library(dplyr)
+
+# Combine the soil_data with the aggregation level
+soil_data <- soil_data %>%
+  mutate(aggregation_level = soil_data_aggregation_level)  # Add aggregation level as a new column
+
+# Aggregate by the aggregation level
+aggregated_soil_data <- soil_data %>%
+  group_by(aggregation_level) %>%
+  summarise(across(where(is.numeric), mean, na.rm = TRUE))  # Calculate mean for numeric columns
+
+aggregated_soil_data = cbind(aggregated_soil_data, Q10_averages)
+
+
+
+# correlation matrix
+names(aggregated_soil_data)
+
+
+selected_columns <- c(
+  "year", "No_trees", "No_trees_ha", "Basal_area", "Mean_DBH","bulk_den",
+  "pH", "phosphorous", "Ntot", "Ctot",
+  "C_stocks", "fungi_biomass", "bacteria_biomass", "actinobacteria_biomass",
+  "gram_pos_biomass", "gram_neg_biomass", "total_biomass", "fungi_bact_rate",
+  "beta_gluco", "acid_phos", "beta_xylo", "chitise",
+  "cellobiohydrolase", "beta_galacto", "alpha_gluco", "lipase",
+  "CO2_flux_hour", #"CO2_flux_norm",
+  "T1_soil", "Soil_moist",
+  "Q10_averages")
+
+# Filter out columns that have all NA values or only one unique value
+non_zero_sd_columns <- aggregated_soil_data[, selected_columns][, sapply(aggregated_soil_data[, selected_columns], function(x) {
+  sd(x, na.rm = TRUE) > 0 && length(unique(na.omit(x))) > 1
+})]
+
+# Calculate correlation on columns with non-zero standard deviation
+correlation_matrix <- cor(aggregated_soil_data[, selected_columns], use = "pairwise.complete.obs")
[…]

### 2024-11-06 data_analysis_sections/conventional_variance_decomposition_enzymes.R

@@ -189,3 +189,17 @@ dev.off()
 
 write.csv(edaphic_means_table, "./Tables/Edaphics_means.csv")
 
+
+
+
+
+
+
+
+
+
+##### mixed models decomposition
+
+names(soil_data)
+
+

### 2024-11-06 master_data_analysis.R

@@ -13,7 +13,6 @@ library(parallel)
 
 set.seed(123)
 
-#TODO Q10 is not correct as range
 
 
 

### 2024-11-06 data_analysis_sections/Q10_explanatory_hypotheses.R
added PLS variance decomposition of Q10

@@ -28,6 +28,16 @@ aggregated_soil_data <- soil_data %>%
 
 aggregated_soil_data = cbind(aggregated_soil_data, Q10_averages)
 
+# Step 2: Extract categorical variables as separate vectors
+site_vector <- soil_data %>%
+  group_by(aggregation_level) %>%
+  summarise(site = first(site)) %>%
+  pull(site)
+
+treatment_vector <- soil_data %>%
+  group_by(aggregation_level) %>%
+  summarise(treatment = first(treatment)) %>%
+  pull(treatment)
 
 
 # correlation matrix
@@ -35,8 +45,8 @@ names(aggregated_soil_data)
 
 
 selected_columns <- c(
-  "year", "No_trees", "No_trees_ha", "Basal_area", "Mean_DBH","bulk_den",
-  "pH", "phosphorous", "Ntot", "Ctot",
+  "year", "No_trees", "No_trees_ha", "Basal_area", "bulk_den", #"Mean_DBH",  removed becuase many NAs in ROmania
+  "pH",  "Ntot", "Ctot", #"phosphorous", removed becuase many NAs in Romania
   "C_stocks", "fungi_biomass", "bacteria_biomass", "actinobacteria_biomass",
   "gram_pos_biomass", "gram_neg_biomass", "total_biomass", "fungi_bact_rate",
   "beta_gluco", "acid_phos", "beta_xylo", "chitise",
@@ -45,13 +55,17 @@ selected_columns <- c(
   "T1_soil", "Soil_moist",
   "Q10_averages")
 
-# Filter out columns that have all NA values or only one unique value
-non_zero_sd_columns <- aggregated_soil_data[, selected_columns][, sapply(aggregated_soil_data[, selected_columns], function(x) {
-  sd(x, na.rm = TRUE) > 0 && length(unique(na.omit(x))) > 1
-})]
+### removing 19.romania because it misses soil temp and moist!
+reduced_dataset = aggregated_soil_data[ rownames(aggregated_soil_data) != "19.romania", selected_columns]
+reduced_dataset_sites = aggregated_soil_data[ rownames(aggregated_soil_data) != "19.romania", "aggregation_level"]
+negative_coords <- which( reduced_dataset< 0, arr.ind = TRUE)
+reduced_dataset[negative_coords] = 0 # there are some negs in the enzymes
+
+
[…]

### 2024-11-06 master_data_analysis.R

@@ -60,7 +60,9 @@ palette_plot <-  c(paletteer::paletteer_c("ggthemes::Orange-Gold", n=length(uniq
 palette_treat <- c("#F4D166FF", "#F6A143FF", "#EE711DFF", "#CD4E22FF", "#9E3A26FF",
                    "#B3E0A6FF", "#7FC572FF", "#5DA554FF", "#378748FF", "#24693DFF",
                    "#BCE4D8FF", "#81C3CBFF", "#46A1B8FF", "#347C9FFF", "#2C5985FF")
-
+palette_site  <- c("#F6A143FF",
+                   "#7FC572FF",
+                   "#46A1B8FF")
 # palette_plot
 palette_plot <- c("#F4D166FF", "#F5CC63FF", "#F6C760FF", "#F7C25CFF", "#F8BD58FF",
                   "#F8B855FF", "#F8B352FF", "#F8AE4FFF", "#F7A94BFF", "#F7A447FF",

### 2024-11-06 data_analysis_sections/Q10_explanatory_hypotheses.R
minor update to the PLS figure

@@ -114,6 +114,6 @@ legend("bottomright", paste("Explained variance (2 comps) = ", round(explained_v
 
 par(mar = c(12, 4, 2, 1))  # Adjust 'mar' to control the space around each plot
 abs_coef <- abs(coef(pls_model))
-barplot(as.vector(abs_coef), main = "", las = 2, col = "steelblue", ylab = "Regression coefficient", names.arg = rownames((abs_coef)))
+barplot(abs_coef[order(abs_coef)], main = "", las = 2, col = "steelblue", ylab = "Regression coefficient", names.arg = rownames(abs_coef)[order(abs_coef)])
 dev.off()
 

### 2024-11-06 data_analysis_sections/Q10_explanatory_hypotheses.R
added mixed model analysis

@@ -40,12 +40,15 @@ treatment_vector <- soil_data %>%
   pull(treatment)
 
 
+treatment_vector = treatment_vector[ rownames(aggregated_soil_data) != "19.romania"]
+site_vector = site_vector[ rownames(aggregated_soil_data) != "19.romania"]
+
 # correlation matrix
 names(aggregated_soil_data)
 
 
 selected_columns <- c(
-  "year", "No_trees", "No_trees_ha", "Basal_area", "bulk_den", #"Mean_DBH",  removed becuase many NAs in ROmania
+ "No_trees_ha", "Basal_area", "bulk_den", #"Mean_DBH",  removed becuase many NAs in ROmania "year",
   "pH",  "Ntot", "Ctot", #"phosphorous", removed becuase many NAs in Romania
   "C_stocks", "fungi_biomass", "bacteria_biomass", "actinobacteria_biomass",
   "gram_pos_biomass", "gram_neg_biomass", "total_biomass", "fungi_bact_rate",
@@ -57,12 +60,12 @@ selected_columns <- c(
 
 ### removing 19.romania because it misses soil temp and moist!
 reduced_dataset = aggregated_soil_data[ rownames(aggregated_soil_data) != "19.romania", selected_columns]
-reduced_dataset_sites = aggregated_soil_data[ rownames(aggregated_soil_data) != "19.romania", "aggregation_level"]
+reduced_dataset_sites = aggregated_soil_data[rownames(aggregated_soil_data) != "19.romania", "aggregation_level"]
 negative_coords <- which( reduced_dataset< 0, arr.ind = TRUE)
 reduced_dataset[negative_coords] = 0 # there are some negs in the enzymes
+aggregated_soil_data$year
 
-
-data_scaled <- scale(aggregated_soil_data[ rownames(aggregated_soil_data) != "19.romania", selected_columns])
+data_scaled <- scale(reduced_dataset)
 
 # Calculate correlation on columns with non-zero standard deviation
 correlation_matrix <- cor(data_scaled, use = "pairwise.complete.obs")
@@ -80,6 +83,14 @@ heatmap(correlation_matrix,
         main = "")
 dev.off()
 
+
+############### mixed model analysis
+library(lme4)
+reduced_dataset_mixed = cbind(as.data.frame(data_scaled), treatment_vector)
[…]

### 2024-11-06 data_analysis_sections/Q10_explanatory_hypotheses.R
relative importance of mixed model (package MuMIn) added

@@ -92,10 +92,25 @@ summary(mixed_model)
 
 summary(lm(predict(mixed_model) ~ reduced_dataset_mixed$Q10_averages))
 
+# Calculate relative importance for fixed effects
+library(MuMIn)
+
+mixed_model_ml <- lmer(Q10_averages ~ No_trees_ha + fungi_bact_rate + C_stocks + T1_soil + Soil_moist + (1 | treatment_vector),
+                       data = reduced_dataset_mixed, REML = FALSE, na.action = na.fail)
+
+dredge_results <- dredge(mixed_model_ml)
+rel_importance <- model.avg(get.models(dredge_results, subset = TRUE))
+str(sw(rel_importance))
+
+png("./Figures/mixed_model_relimp.png", height=1400, width = 1400, res=300)
+par(mar = c(9, 4, 2, 1))  # Adjust 'mar' to control the space around each plot
+barplot(as.numeric(sw(rel_importance)), names.arg = names(sw(rel_importance)),  col = "steelblue", ylab = "Relative importance", las=2)
+dev.off()
 
 
+##########
 
-############### PLS analysis
+##### PLS analysis
 library(pls)
 pls_model <- plsr(Q10_averages ~ ., data = as.data.frame(data_scaled), ncomp = 5)
 summary(pls_model)

### 2024-11-12 data_analysis_sections/Q10_explanatory_hypotheses.R
Minor modification in the code

@@ -108,6 +108,8 @@ barplot(as.numeric(sw(rel_importance)), names.arg = names(sw(rel_importance)),  
 dev.off()
 
 
+
+
 ##########
 
 ##### PLS analysis

### 2024-11-12 Manuscript_draft/Draft.Rmd
Minor update on the variance decomposition

@@
 
-## Machine learning variance decomposition: predicting respiration based on edaphic, ecological and enzymatic data
+## Variance decomposition: predicting respiration based on edaphic, ecological and enzymatic data
 
@@
 Only the 20 variables explaining most variance are shown here.](../Figures/variance_decomposition_enzyme.png){width=50%}
+
+## Variance decomposition: predicting Q10 based on edaphic, ecological and enzymatic data
+
+![.](../Figures/mixed_model_relimp.png){width=50%}
+
+![.](../Figures/Q10_PLS_biplot.png){width=50%}
 
@@
 # Appendix
-![Posteriors of the scaling parameter, varying as a function of single plots](../Figures/posteriors_boxplots_q10.png){width=60%}
+![.](../Figures/posteriors_boxplots_q10.png){width=60%}
 

### 2024-11-12 data_analysis_sections/Q10_explanatory_hypotheses.R

@@ -54,7 +54,7 @@ selected_columns <- c(
   "gram_pos_biomass", "gram_neg_biomass", "total_biomass", "fungi_bact_rate",
   "beta_gluco", "acid_phos", "beta_xylo", "chitise",
   "cellobiohydrolase", "beta_galacto", "alpha_gluco", "lipase",
-  "CO2_flux_hour", #"CO2_flux_norm",
+  "CO2_flux_norm",   #"CO2_flux_hour",
   "T1_soil", "Soil_moist",
   "Q10_averages")
 
@@ -83,14 +83,18 @@ heatmap(correlation_matrix,
         main = "")
 dev.off()
 
+# ---------------------------------------------
+#              Mixed models
+# ---------------------------------------------
 
-############### mixed model analysis
+############### mixed model with C
+
 library(lme4)
 reduced_dataset_mixed = cbind(as.data.frame(data_scaled), treatment_vector)
 mixed_model <- lmer(Q10_averages ~ No_trees_ha + fungi_bact_rate + C_stocks + T1_soil + Soil_moist + treatment_vector + (1 | site_vector), data = reduced_dataset_mixed)
 summary(mixed_model)
 
-summary(lm(predict(mixed_model) ~ reduced_dataset_mixed$Q10_averages))
+mixed_model_lm <- summary(lm(predict(mixed_model) ~ reduced_dataset_mixed$Q10_averages))
 
 # Calculate relative importance for fixed effects
 library(MuMIn)
@@ -105,7 +109,31 @@ str(sw(rel_importance))
 png("./Figures/mixed_model_relimp.png", height=1400, width = 1400, res=300)
 par(mar = c(9, 4, 2, 1))  # Adjust 'mar' to control the space around each plot
 barplot(as.numeric(sw(rel_importance)), names.arg = names(sw(rel_importance)),  col = "steelblue", ylab = "Relative importance", las=2)
+legend("topright", legend = list(bquote(R^2 == .(round(mixed_model_lm$r.squared, 2)))), bty="n")
 dev.off()
+
+
+
+############### mixed model with N
+
+mixed_model_N <- lmer(Q10_averages ~ No_trees_ha + fungi_bact_rate + Ntot + T1_soil + Soil_moist + treatment_vector + (1 | site_vector), data = reduced_dataset_mixed)
+summary(mixed_model_N)
+
[…]

### 2024-11-12 Manuscript_draft/Draft.Rmd
new draft.rmd

--- 2024-11-05 Manuscript_draft/Draft.Rmd
@@
 We have analysed the sensitivity of soil respiration to microclimate (i.e., soil temperature and soil moisture) by constructing a mechanistic parametric model and compared to a non-parametric model (i.e., random forest).
-
+We have also explored the main drivers of variation in Q10 across different forests and forest managements disturbances.
 # Material and methods
@@
 
+## Temperature sensitivity (Q10)
+
 ## Machine learning variance decomposition on enzymes
@@
 ## Characterization of study sites
-Tree density in France was very high compared to Spain and especially to Romania where the tree density was very low (Figure x).
-In contrast, the DBH was the lowest in France and the highest in Romania (Figure x).
-Spain showed intermediate values of both tree density and DBH (Figure x).
+Tree density in France was very high compared to Spain and especially to Romania where the tree density was very low (Figure 1).
+In contrast, the DBH was the lowest in France and the highest in Romania (Figure 1).
+Spain showed intermediate values of both tree density and DBH (Figure 1).
 For soil variables, in the untouched forest (control), the bulk density was the highest in Spain and the lowest in France.
@@
 The highest significant levels of C and N were found in Spain while the lowest C content was found in Romania.
-France in turn presented doubled C:N rates than the other two sites (Figure x).
+France in turn presented doubled C:N rates than the other two sites (Figure 1).
 Total microbial biomass, fungal biomass and bacterial biomass peaked in France, followed by Spain and it was the lowest in Romania.
-Similarly, in general, activities of most enzymes were higher in France compared with Spain and Romania (Figure x).
+Similarly, in general, activities of most enzymes were higher in France compared with Spain and Romania (Figure 2).
 
-##Effects of management disturbance treatments on soil physical-chemical and biological properties
-The disturbance treatments barely affected the soil properties in Romania (Figure x).
-In France, the bulk density increased as the intensity of the disturbance did (i.e., from thinning with slash to clear-cut no slash) and in Spain it significantly increased in all the disturbance treatments compared with control (Figure x).
+![Figure1.
+Boxplots of stand characteristics and soil physical-chemical properties per disturbance treatment and study site.
+Red shades = France, green shades = Romania, blue shades = Spain](../Figures/edaphic_boxplots.png){width=80%}
+
+## Effects of management disturbance treatments on soil physical-chemical and biological properties
+The disturbance treatments barely affected the soil properties in Romania (Figure 1).
+In France, the bulk density increased as the intensity of the disturbance did (i.e., from thinning with slash to clear-cut no slash) and in Spain it significantly increased in all the disturbance treatments compared with control (Figure 1).
 The pH only significantly increased in both clear-cut treatments in Spain.
-Moreover, in Spain the N content decreased in the thinning and clear-cut without slash (Figure x).
-In France, compared with controls, C content increased in all the disturbance treatment, but in the clear-cut without slash that decreased (Figure x).
-In Spain, C content only decreased in the thinning-without slash treatment (Figure x).
+Moreover, in Spain the N content decreased in the thinning and clear-cut without slash (Figure 1).
+In France, compared with controls, C content increased in all the disturbance treatment, but in the clear-cut without slash that decreased (Figure 1).
+In Spain, C content only decreased in the thinning-without slash treatment (Figure 1).
 Similar to soil properties, microbial biomass, fungal biomass and bacterial biomass did not change with the disturbance treatment in Romania.
-In Spain, these variables significantly decreased in the disturbance treatments where the slash was removed and in France only in the clear-cut treatment where the slash was removed (Figure x).
-The disturbance treatments barely affected enzymatic activities in Romania and Spain, but generally decreased in clear-cut where slash was removed in France (Figure x).
+In Spain, these variables significantly decreased in the disturbance treatments where the slash was removed and in France only in the clear-cut treatment where the slash was removed (Figure 2).
+The disturbance treatments barely affected enzymatic activities in Romania and Spain, but generally decreased in clear-cut where slash was removed in France (Figure 2).
 
-![Boxplots of soil microbial biomass and enzymatic activities per disturbance treatment and study site.
+![Figure2.
+Boxplots of soil microbial biomass and enzymatic activities per disturbance treatment and study site.
 Red shades = France, green shades = Romania, blue shades = Spain](../Figures/enzymes_boxplots.png){width=80%}
-![Boxplots of stand characteristics and soil physical-chemical properties per disturbance treatment and study site.
-Red shades = France, green shades = Romania, blue shades = Spain](../Figures/edaphic_boxplots.png){width=80%}
 
@@
 
-![Climatic parameters (soil temperature and soil moisture) and respiration measured at the study sites (mean values per month and treatment after loess interpolation with span = 0.5).
-Please notice thot not every time series covers the same period of time.
-Red shades = France, green shades = Romania, blue shades = Spain](../Figures/climatedata_means.png){width=60%}
-
-As expected, in Spain the temperature variation was the lowest compared to the other two sites.
+As expected, in Spain the temperature variation was the lowest compared to the other two sites (Figure 3a).
 In general, control plots showed the lowest thermal amplitude in all sites (i.e., with higher temperatures in winter and lower in summer).
@@
 Similarly, in Romania, clear-cut increased soil moisture independently of slash addition.
-Contrarily, in France, soil moisture tended to decrease in the clear-cut treatment with and without slash (Figure x).
+Contrarily, in France, soil moisture tended to decrease in the clear-cut treatment with and without slash (Figure 3b).
 Both in Spain and France, removal of slash decreased respiration rates, independently of tree cutting intensity, and thinning coupled with slash addition kept the respiration rates as the same level as controls.
 In Romania, effects of the disturbances were lower, however clear-cut increased respiration rates, especially when slash was added to the soil.
-At the end of the monitoring period, three years for Spain and two years for France and Romania after disturbance, respiration rates increased in the clear-cut with slash addition in the three study sites (Figure x).
+At the end of the monitoring period, three years for Spain and two years for France and Romania after disturbance, respiration rates increased in the clear-cut with slash addition in the three study sites (Figure 3c).
+
+![Figure 3.
+Climatic parameters (soil temperature and soil moisture) and respiration measured at the study sites (mean values per month and treatment after loess interpolation with span = 0.5).
+Please notice thot not every time series covers the same period of time.
+Red shades = France, green shades = Romania, blue shades = Spain](../Figures/climatedata_means.png){width=60%}
 
 ## Benchmark models and parametric model fit
-![Fit of the two benchmark machine learning models, linear (a) and random forest (b)](../Figures/fit_ML_validation.png){width=80%}
+![Figure 4. Fit of the two benchmark machine learning models, linear (a) and random forest (b)](../Figures/fit_ML_validation.png){width=80%}
 
-![Fit of the parametric model,  constant `E_a`](../Figures/fit_model_validation.png){width=60%}
+![Figure 5. Fit of the parametric model,  constant `E_a`](../Figures/fit_model_validation.png){width=60%}
 
-## Parametric model posteriors, constant `E_a` 
-The activation energy (Ea) and temperature sensitivity (Q10) values of SOM mineralization were consistently higher in Romania, followed by France and Spain.
+## Temperature sensitivity (Q10) and parametric model posteriors 
+The temperature sensitivity (Q10) values of SOM mineralization were consistently higher in Romania (2.9 in controls), followed by France (2.7) and Spain (2.5) (Figure S4).
+The activation energy (Ea) followed the same pattern as Q10 (Figure 6a, Figure S4).
 In general, the activation energy and Q10 values decreased with slash addition due to higher SOM quality compared with treatments in which slash was removed.
 This was especially evident in Spain for both thinning and clear-cut and in the clear treatment in France.
-The activation energy and Q10 in the treatments where slash was added did not differ from controls (Figure x, xa).
-These results were consistent when the activation energy was analysed as a function of time since disturbance and the activation energy systematically decreased in all treatments and all countries with time since disturbance (Figure x).
-The photosynthetic activity did not modify the sensitivity of soil respiration to temperature (peak and amplitude factors from the sine-function not affected) (Figure xb, c).
-However, the amplitude of the sine-function tended to be higher in France (Figure xc).
-The parameters of soil moisture function (a and b) did not change the response of the model either (Figure xd, e).
+The activation energy and Q10 in the treatments where slash was added did not differ from controls (Figure 6a, Figure S4).
+These results were consistent when the activation energy was analysed as a function of time since disturbance and the activation energy systematically decreased in all treatments and all countries with time since disturbance (Figure 7).
+The photosynthetic activity did not modify the sensitivity of soil respiration to temperature (peak and amplitude factors from the sine-function not affected) (Figure 6b, c).
+However, the amplitude of the sine-function tended to be higher in France (Figure 6c).
+The parameters of soil moisture function (a and b) did not change the response of the model either (Figure 6d, e).
 
-![Posteriors of the parameters varying as a function of treatments, `E_a` constant over time](../Figures/posteriors_boxplots.png){width=60%}
+![Figure 6.
+Posteriors of the parameters from the parametric model varying as a function of treatments: (a) activation energy(Ea) for temperature, (b) peak and (c) amplitude parameters from the seasonality sine-function, (d) and (e) correspond to a and b parameters from the soil moisture function, all testing the effects on respiration.
+Ea constant over time.
+Red shades = France, green shades = Romania, blue shades = Spain](../Figures/posteriors_boxplots.png){width=60%}
 
-## Parametric model posteriors,  time-variant `E_a` 
-![Posteriors of the `E_a` parameter varying as a function of treatments and sampling year](../Figures/timeseries_Ea_timevariant.png){width=60%}
+![Figure 7. Posteriors of the activation energy (Ea) parameter varying as a function of treatments and sampling year per country](../Figures/timeseries_Ea_timevariant.png){width=60%}
+
+The variables that most variance explained of Q10 were N content and fungal to bacteria ratio and then number of trees, beta-xylosidases and soil temperature (Figure x).
+
+![Figure 8.
+Principal component regression between edaphic and ecological variables and temperature sensitivity (Q10) analysed by Partial least squares (PLS) regression (upper pannel).
+Variance decomposition of main variables explaining Q10 (lower pannel)](../Figures/Q10_PLS_biplot.png){width=60%}
 
@@
 
-## Role of temperature and moisture in conditioning decomposition
-Temperature and moisture seem the most important components in explaining the variance of the soil respiration.
-The seasonality scaling, although apparently differentiating slightly between sites, presents always very low amplitude posteriors, weighting on average only a 5% of the total measured respiration.
+## Role of temperature and moisture in decomposition
+Our model showed that soil temperature and moisture were able to explained more than 50% of variance in soil respiration.
+Because phenologial patterns in deciduous species can affect sensitivity of soil respiration to temperature (Curiel-Yuste et al 2004), in this model we have also included a seasonality scaling.
+The seasonal scaling although apparently differentiated slightly between sites, it always presented very low amplitude posteriors, weighting on average only a 5% of the total measured respiration.
 It was more important in France (between 4.5% and 12.2%) and much less in Spain and Romania (between 2.0% and 3.2%).
-This seasonality might represent many different processes, but the most likely are seasonal processes connected with the above-ground vegetation and their inputs to the soil, which in turn would control microbial communities.
+This is probably related to higher photosynthetic activity variation of Quercus pubescens growing in a Mediterranean climate in France.
+This would indicate higher changes in root respiration and proliferation that would in turn affect microbial communities.
+Because in the three study sites there were deciduous Quercus species, another explanation is that soil respiration could have been also affected by the deciduousness of the understorey vegetation (Curiel-Yuste 2004).
 
-## Activation energy of different organic material
+## Forest management effect on activation energy and sensitivity to temperature
+Our results confirmed the hypothesis 2, since addition of slash decreased the activation energy probably related to a reduction in chemical recalcitrance.
+This supports the carbon quality-temperature hypothesis stating that the biochemically recalcitrant C pool is more sensitive to changes in temperature than the labile C pool (Bossata and Agren 1999).
+We also observed a positive correlation between Q10 and activation energy agreeing with Davidson & Janssens, 2006 and Sierra, 2012 that the higher the activation energy the higher is the relative temperature sensitivity.
 
-## Role of edaphic, ecological and microbiological variables in conditioning decomposition
-Any other variable than soil temperature and moisture did not seem to play a pivotal role in explaining the variance of the soil respiration.
+In line with our hypothesis x, we have observed a general negative trend of activation energy with time in all management disturbance treatments and all sites.
+This result also gives support to the C quality temperature hypothesis since fresh inputs from the new vegetation low down the temperature sensitivity of C mineralization as these inputs are more labile to decomposed (Lefevre et al 2014).
+Still, the growth of new vegetation, within three years from disturbance, did not compensate for the decreases of activation energies in the plots where removal of slash had an effect.
+Additionally, we cannot rule out the fact that other mechanisms could have contributed to low activation energies with time such as differences between years in primary production (i.e., contribution of autotrophic respiration) or changes in substrate and enzyme diffusivity driven by differences in soil moisture between years (Davinson and Jansen 2022).
+In line with this, for example, in the control plots from Spain the activation energy decreased from year zero (2021) to year two (2023) and then it increased a bit while in disturbed plots was always descendent.
+Since changes in activation energy due to the growth of new vegetation is not expected in undisturbed control plots and given that 2022 and 2023 were extremely dry in Spain, probably drought exerted a constrained access of enzymes to substrates (Davinson and Jansen 2022).
+This suggest that drought effect may have similar impacts on temperature sensitivity as chemical recalcitrance (Levegre et al 2014).
+
+## Role of edaphic, ecological and microbiological variables in temperature sensitivity
+The Q10 values obtained in our study were comparable to the Q10 values measured in the field by other studies (ref).
+Xu et al 2015 observed the largest relative frequency of Q10 values across forest in China was within the range of 2.0–3.0.
+Fierer et al.
+(2006) observed Q10 values ranging from 2.2 to 4.6 (average of 3.0) at continental-scale in an incubation experiment.
+
+Hydrologic leaching of dissolved and particulate carbon compounds can also be important outputs of C from the system (Davinson and Jansen 2006), that could explained the low carbon content in Romania (sandy soil with low retention capacity).
+Longer time of frozen soils could reduce the diffusion of substrates and extracellular enzymes within the soil (Davinson and Jansen 2006), given higher values of Q10 not related to recalcitrance.
+
+Our analyses revealed the three variables that explained most of the variation of Q10 were related to i) the quality of substrate (i.e., nitrogen content as limiting nutrient in forest), ii) the microorganisms, in particular, fungal to bateria ratio and iii) the primary productivity (that is, the number of trees).
 
@@
 # Appendix
-![Posteriors of the scaling parameter, varying as a function of single plots](../Figures/posteriors_boxplots_q10.png){width=60%}
 
-![Posteriors of the scaling parameter, varying as a function of single plots](../Figures/Appendix/posteriors_boxplots_A.png){width=60%}
+![Figure S1. Temperature scaling function](../Figures/Appendix/Temp_function.png){width=50%}
 
-![Posteriors of the parameters of the temperature function, plotted as density distributions](../Figures/Appendix/posteriors_temp.png)
+![Figure S2. Moisture scaling function](../Figures/Appendix/Moist_function.png){width=50%}
 
-![Posteriors of the parameters of the moisture function, plotted as density distributions](../Figures/Appendix/posteriors_seasonality.png)
+![Figure S3. Seasonality scaling function](../Figures/Appendix/Season_function.png){width=50%}
 
-![Posteriors of the parameters of the seasonality function, plotted as density distributions](../Figures/Appendix/posteriors_moisture.png)
+![Figure S4.
+Boxplots of temperature sensitivity (Q10) of soil respiration per disturbance treatment and study site.
+Red shades = France, green shades = Romania, blue shades = Spain](../Figures/posteriors_boxplots_q10.png){width=60%}
 
-![Ratio (expressed as %) between the mean amplitude parameter and the mean measured respiration by treatment](../Figures/Amplitude_ratio.png)
+![Figure S5. Posteriors of the scaling parameter, varying as a function of single plots](../Figures/Appendix/posteriors_boxplots_A.png){width=60%}
 
-![Temperature scaling function](../Figures/Appendix/Temp_function.png){width=50%}
+![Figure S6. Posteriors of the parameters of the temperature function, plotted as density distributions](../Figures/Appendix/posteriors_temp.png)
 
-![Moisture scaling function](../Figures/Appendix/Moist_function.png){width=50%}
+![Figure S7. Posteriors of the parameters of the moisture function, plotted as density distributions](../Figures/Appendix/posteriors_seasonality.png)
 
-![Seasonality scaling function](../Figures/Appendix/Season_function.png){width=50%}
+![Figure S8. Posteriors of the parameters of the seasonality function, plotted as density distributions](../Figures/Appendix/posteriors_moisture.png)
+
+![Figure S9. Ratio (expressed as %) between the mean amplitude parameter and the mean measured respiration by treatment](../Figures/Amplitude_ratio.png)
 

### 2024-11-12 data_analysis_sections/parametric_model_fitting_time_variant.R
Updated runs (new data files)

@@ -1,5 +1,5 @@
 
-
+gc()
 library(parallel)
 library(lubridate)
 
@@ -255,5 +255,6 @@ dev.off()
 
 
 #to leave memory for following tasks
-# rm(fit_bytreat_indA_Moy_sin_timevar)
-# rm(post_bytreat_indA_Moy_timevar)
+rm(fit_bytreat_indA_Moy_sin_timevar)
+rm(post_bytreat_indA_Moy_timevar)
+gc()

### 2024-11-13 Manuscript_draft/Draft.Rmd
Hypotheses

@@
 (1) logging of trees will decrease the respiration rates because of the strong decrease in vegetation activity and microbiota associated.
-In turn, addition of slash will provide labile organic matter to decompose increasing the respiration rates.
-(2) the variation in soil respiration would be mainly dependent on microclimatic variables (soil temperature and moisture), especially in the management treatments where vegetation was removed, since heterotrophic respiration will prevail over autotrophic respiration, and where the litter layer (slash) was removed
-(3) the sensitivity of soil respiration to microclimatic conditions will also depend on the type of climate/forest, being the Mediterranean the less sensitive as the microbial communities will be better adapted to the strong seasonal changes in water availability.
-(4) the sensitivity of respiration to microclimate will decrease as the time passes since the disturbance because of new fresh C input from the growth of understory vegetation
+This decrease will be compensated by addition of slash which provides labile organic matter, increasing the heterotrophic respiration.
+(2) soil respiration would be more dependent on microclimatic variables (soil temperature and moisture), in the management disturbance treatments where the vegetation was removed and where the litter layer (slash) was removed, by depletion of labile organic matter.
+Likewise, the sensitivity of respiration to microclimate will decrease as the time passes since the disturbance because of new fresh C input from the growth of understory vegetation
+(4) the sensitivity of soil respiration to microclimatic conditions will also depend on the type of climate/forest, being the Mediterranean the less sensitive as the microbial communities will be better adapted to the strong seasonal changes in water availability.
 
@@
 We have also explored the main drivers of variation in Q10 across different forests and forest managements disturbances.
+
 # Material and methods

### 2024-11-13 Manuscript_draft/Draft.Rmd
small changes

@@
 Likewise, the sensitivity of respiration to microclimate will decrease as the time passes since the disturbance because of new fresh C input from the growth of understory vegetation
-(4) the sensitivity of soil respiration to microclimatic conditions will also depend on the type of climate/forest, being the Mediterranean the less sensitive as the microbial communities will be better adapted to the strong seasonal changes in water availability.
+(3) the sensitivity of soil respiration to microclimatic conditions will also depend on the type of climate/forest, being the Mediterranean the less sensitive as the microbial communities will be better adapted to the strong seasonal changes in water availability.
 
@@
 # Discussion
+General paragraph
 
-## Role of temperature and moisture in decomposition
+Hydrologic leaching of dissolved and particulate carbon compounds can also be important outputs of C from the system (Davinson and Jansen 2006), that could explained the low carbon content in Romania (sandy soil with low retention capacity).
+
+### Role of temperature and moisture in decomposition (model description)
 Our model showed that soil temperature and moisture were able to explained more than 50% of variance in soil respiration.
@@
 ## Forest management effect on activation energy and sensitivity to temperature
-Our results confirmed the hypothesis 2, since addition of slash decreased the activation energy probably related to a reduction in chemical recalcitrance.
-This supports the carbon quality-temperature hypothesis stating that the biochemically recalcitrant C pool is more sensitive to changes in temperature than the labile C pool (Bossata and Agren 1999).
-We also observed a positive correlation between Q10 and activation energy agreeing with Davidson & Janssens, 2006 and Sierra, 2012 that the higher the activation energy the higher is the relative temperature sensitivity.
+Generally, our results confirmed the hypothesis 1, since slash addition increased the respiration rates compared to the treatments where the slash was removed.
+Addition of slash created an artificial litter layer that buffered the effects over soil microclimatic conditions, and provided new habitat and food availability for soil organisms (ref).
+These results indicate that increases in heterotrophic respiration due to due to an increment in organic matter availability by slash addition compensated for the decreases in autotrophic respiration and root-associated microbiota due to tree logging and soil disturbance (Borkhuu et al.
+2015).
+Moreover, the strength of the response was also dependent on the forest, being more evident in Spain and France where the tree density was much higher than Romania.
+In Romania such a low density of trees together with a dense understory of grasses in all treatments might explain the lack of response and indicate high degree of resilience to perturbations.
 
-In line with our hypothesis x, we have observed a general negative trend of activation energy with time in all management disturbance treatments and all sites.
+Our results also confirmed the hypothesis 2, since addition of slash decreased the activation energy of SOM decomposition.
+We argue that this result is probably related to a reduction in chemical recalcitrance, supporting the carbon quality-temperature hypothesis stating that the biochemically recalcitrant C pool is more sensitive to changes in temperature than the labile C pool (Bossata and Agren 1999).
+Moreover, the activation energy correlated positively with the Q10, agreeing with Davidson & Janssens, 2006 and Sierra, 2012 that the higher the activation energy the higher is the relative temperature sensitivity.
+We have also observed a general negative trend of activation energy with time in all management disturbance treatments and all sites.
 This result also gives support to the C quality temperature hypothesis since fresh inputs from the new vegetation low down the temperature sensitivity of C mineralization as these inputs are more labile to decomposed (Lefevre et al 2014).
-Still, the growth of new vegetation, within three years from disturbance, did not compensate for the decreases of activation energies in the plots where removal of slash had an effect.
-Additionally, we cannot rule out the fact that other mechanisms could have contributed to low activation energies with time such as differences between years in primary production (i.e., contribution of autotrophic respiration) or changes in substrate and enzyme diffusivity driven by differences in soil moisture between years (Davinson and Jansen 2022).
-In line with this, for example, in the control plots from Spain the activation energy decreased from year zero (2021) to year two (2023) and then it increased a bit while in disturbed plots was always descendent.
+Still, the growth of new vegetation, within three years from disturbance, did not compensate for the decreases in activation energies in the plots where removal of slash had an effect.
+Additionally, we cannot rule out the fact that other mechanisms could have contributed to low down activation energies with time such as differences between years in primary production (i.e., contribution of autotrophic respiration) or changes in substrate and enzyme diffusivity driven by differences in soil moisture between years (Davinson and Jansen 2022).
+In line with this, for example, in the control plots from Spain the activation energy decreased from year zero (2021) to year two (2023) and then it increased a bit, while in disturbed plots it was always descendant.
 Since changes in activation energy due to the growth of new vegetation is not expected in undisturbed control plots and given that 2022 and 2023 were extremely dry in Spain, probably drought exerted a constrained access of enzymes to substrates (Davinson and Jansen 2022).
@@
 (2006) observed Q10 values ranging from 2.2 to 4.6 (average of 3.0) at continental-scale in an incubation experiment.
-
-Hydrologic leaching of dissolved and particulate carbon compounds can also be important outputs of C from the system (Davinson and Jansen 2006), that could explained the low carbon content in Romania (sandy soil with low retention capacity).
-Longer time of frozen soils could reduce the diffusion of substrates and extracellular enzymes within the soil (Davinson and Jansen 2006), given higher values of Q10 not related to recalcitrance.
+Still, we observed a large variation of Q10 depending on the forest, being the highest in Romania.
+This is in agreement with the known higher temperature sensitivity in colder regions (Koven et al 2017).
+The reason could be related to longer time of frozen soils that could reduce the diffusion of substrates and extracellular enzymes within the soil (Davinson and Jansen 2006), giving higher values of Q10 not related to recalcitrance.
 

### 2024-11-14 Manuscript_draft/Draft.Rmd
small changes in discussion

@@
 ## Role of edaphic, ecological and microbiological variables in temperature sensitivity
-The Q10 values obtained in our study were comparable to the Q10 values measured in the field by other studies (ref).
+The Q10 values obtained in our study were comparable to the Q10 values measured by other studies.
 Xu et al 2015 observed the largest relative frequency of Q10 values across forest in China was within the range of 2.0–3.0.
@@
 This is in agreement with the known higher temperature sensitivity in colder regions (Koven et al 2017).
-The reason could be related to longer time of frozen soils that could reduce the diffusion of substrates and extracellular enzymes within the soil (Davinson and Jansen 2006), giving higher values of Q10 not related to recalcitrance.
+Longer time of frozen soils in Romania could have reduced the diffusion of substrates and extracellular enzymes within the soil (Davinson and Jansen 2006), giving higher values of Q10 not related to recalcitrance.
 
-Our analyses revealed the three variables that explained most of the variation of Q10 were related to i) the quality of substrate (i.e., nitrogen content as limiting nutrient in forest), ii) the microorganisms, in particular, fungal to bateria ratio and iii) the primary productivity (that is, the number of trees).
+Our analyses revealed the three variables that explained most of the variation of Q10 were related to i) the quality of substrate (i.e., nitrogen content as limiting nutrient in forest), ii) the microorganisms, in particular, fungal to bacteria ratio and iii) the soil temperature.
+Our results indicate that SOM decomposition not only depends on the substrate biochemistry but also on the ability of the existing microbial community to decompose the available substrate.
+This is in partial agreement with a recent global scale study stating that soil microbiomes are the main players governing the responses of soil heterotrophic respiration to temperature, and secondly substrate quantity and biochemical recalcitrance (Sáez-Sandino et al 2023).
+Since total microbial biomass did not correlated with total nitrogen, the observed effect of nitrogen might be related to an increased in mineral nitrogen.
+Compared to controls, fungal and bacteria biomass decreased in both thinning and clear-cut indicating a stronger decrease in ectomycorrhizal fungi compared to root-fungi associated bacteria.
+In forest, the role of fungi, especially of ectomycorrhizal fungi in decomposing SOM while mining for nutrients.
+The stronger decrease in ectomycorrhizal fungi, might have increased sensitivity to temperature linked to reduced fungal activity.
+The decrease This could   The higher increase in fungal biomass compared to bacteria in the treatments with slash indicate a faster colonization of saprotrophic fungi in litter.
+An probably temperatre sensitivity is controlled by a reduced number of taxa with a high abundance (Sáez-Sandino et al 2023).
+Diversity metrics would give more insights about that.
 

### 2024-11-14 Manuscript_draft/Draft.Rmd
disc

@@
 Our analyses revealed the three variables that explained most of the variation of Q10 were related to i) the quality of substrate (i.e., nitrogen content as limiting nutrient in forest), ii) the microorganisms, in particular, fungal to bacteria ratio and iii) the soil temperature.
-Our results indicate that SOM decomposition not only depends on the substrate biochemistry but also on the ability of the existing microbial community to decompose the available substrate.
-This is in partial agreement with a recent global scale study stating that soil microbiomes are the main players governing the responses of soil heterotrophic respiration to temperature, and secondly substrate quantity and biochemical recalcitrance (Sáez-Sandino et al 2023).
-Since total microbial biomass did not correlated with total nitrogen, the observed effect of nitrogen might be related to an increased in mineral nitrogen.
+Our results indicate that SOM decomposition not only depends on the substrate biochemistry but also on the ability of the microbial community to decompose the available substrate.
+This is in partial agreement with a recent global-scale study stating that soil microbiomes are the main players governing the responses of soil heterotrophic respiration to temperature, and secondly substrate quantity and biochemical recalcitrance (Sáez-Sandino et al 2023).
+Since total microbial biomass did not correlate with total nitrogen, the observed effect of nitrogen might be related to an increased in mineral nitrogen.
+This would alleviate the nitrogen limitation normally present in forests and favouring decomposition.
+
 Compared to controls, fungal and bacteria biomass decreased in both thinning and clear-cut indicating a stronger decrease in ectomycorrhizal fungi compared to root-fungi associated bacteria.

### 2024-11-15 master_data_analysis.R
Updated the analysis of soil variables after new file

@@ -21,7 +21,8 @@ set.seed(123)
 #data <- read.csv("Database_co2_28_05_2024.csv")
 data <- read_xlsx("Database_co2_18_10_2024_format_Lorenzo.xlsx", sheet=1)
 str(data)
-soil_data <- read_xlsx("Soil_data_Lorenzo_03_09_2024_enzymes.xlsx", sheet=1)
+#soil_data <- read_xlsx("Soil_data_Lorenzo_03_09_2024_enzymes.xlsx", sheet=1)
+soil_data <- read_xlsx("Soil_data_Lorenzo_14_11_2024_enzymes.xlsx", sheet=1)
 
 
 data$treatment <- as.factor(data$treatment)

### 2024-11-15 master_data_analysis.R
Updated also the Q10 analisys part

@@ -210,6 +210,15 @@ source("./data_analysis_sections/parametric_model_fitting_time_variant.R")
 source("./data_analysis_sections/parametric_model_fitting.R")
 
 
+
+# ************************************************************
+# ************** Testing Q10 explanatories ******************
+# ************************************************************
+source("./data_analysis_sections/Q10_explanatory_hypotheses.R")
+
+
+
+
 ## Working on the residuals
 
 residuals_indA_moy <- predicted_means_indA_Moy - processed_data_filtered$CO2_flux_hour

### 2024-11-15 Manuscript_draft/Draft.Rmd
discussion

--- 2024-11-13 Manuscript_draft/Draft.Rmd
@@
 Addition of slash created an artificial litter layer that buffered the effects over soil microclimatic conditions, and provided new habitat and food availability for soil organisms (ref).
-These results indicate that increases in heterotrophic respiration due to due to an increment in organic matter availability by slash addition compensated for the decreases in autotrophic respiration and root-associated microbiota due to tree logging and soil disturbance (Borkhuu et al.
+These results indicate that increases in heterotrophic respiration due to an increment in organic matter availability by slash addition compensated for the decreases in autotrophic respiration and root-associated microbiota due to tree logging and soil disturbance (Borkhuu et al.
 2015).
@@
 This is in agreement with the known higher temperature sensitivity in colder regions (Koven et al 2017).
-The reason could be related to longer time of frozen soils that could reduce the diffusion of substrates and extracellular enzymes within the soil (Davinson and Jansen 2006), giving higher values of Q10 not related to recalcitrance.
+Longer time of frozen soils in Romania could have reduced the diffusion of substrates and extracellular enzymes within the soil (Davinson and Jansen 2006), giving higher values of Q10 not related to recalcitrance.
 
-Our analyses revealed the three variables that explained most of the variation of Q10 were related to i) the quality of substrate (i.e., nitrogen content as limiting nutrient in forest), ii) the microorganisms, in particular, fungal to bateria ratio and iii) the primary productivity (that is, the number of trees).
+Our analyses revealed the three variables that explained most of the variation of Q10 were related to i) the quality/quantity of substrate (i.e., nitrogen content), ii) the microorganisms, in particular, fungal to bacteria ratio and iii) the soil temperature.
+Nitrogen fertilization experiments have shown that nitrogen availability decreases the Q10 (Mo et al 2008; Zhang et al 2019), probably related to changes in microbial carbon use efficiency (Manzoni et al 2012).
+In our experiment, it seems that addition of slash have stimulated decomposition of labile C which increased N availability (Ghosh et al., 2019; Shahbaz et al.,2017), and thus decreased the temperature sensitivity.
+Our results indicate that SOM decomposition not only depends on the substrate biochemistry but also on the ability of the microbial community to decompose the available substrate.
+This is in agreement with a recent global-scale study stating that soil microorganism are the main players governing the responses of soil heterotrophic respiration to temperature, and secondly substrate quantity and biochemical recalcitrance (Sáez-Sandino et al 2023).
+
+We did not measure chemical recalcitrance per se, but our results suggest that microbial communities might have a key role in determining Q10 in forest ecosystems.
+Lower Q10 values associated to the addition of fresh plant material as slash was related to higher fungal to bateria rates.
+Although both fungi and bacteria have the capacity to decompose litter components (Stursova et al.
+2012), fungi has a major role in the decomposition of plant biomass due to their mycelial growth, the ability to translocate nutrients and the possession of a strong enzymatic capacity (Schneider et al.
+2012, Tlaskal et al 2016).
+In contrast, bacteria seems to preferentially gain carbon from live or death fungal biomass and increase at later stages of litter decomposition (Tlaskal et al 2016, Lindahl et al 2010).
+Since ectomycorrhizal fungal mycelium comprise most of the fungal biomass in forests (Wallander et al 2001), a strong reduction in ectomycorrhizal fungal biomass is expected in all the tree removal disturbance treatments, thus decreasing the fungal to bacteria ratio.
+As mentioned, this decrease in ectomycorrhizal fungal biomass might be compensated by a rapid growth of saprotrophic fungal species where slash is added.
+Community data would give more insights about that.
+Surprinsingly, hydrolitic enzymes were not the main variables affecting Q10.
+This effect might be masked by the positive correlation of fungal and bacteria biomass with enzymes.
+Still, most enzymes correlated negatively with Q10.
+The exception was beta-xylosidase, that correlated positively with Q10 and explained a high proportion of variance.
+The only microbial group that strongly correlated with beta-xylosidase was the actinobacteria and problably the relationship is determined by the exceptionally high activity of these enzyme in Romania, compared with other sites.
+This result is consistent with previous studies showing that members of the phylum Actinobacteria are oligotrophic micoorganisms with the ability to compete for resources in a resource-limited environment (Trivedi et al 2013; Bao et al 2021).
 

### 2024-11-19 data_analysis_sections/Q10_explanatory_hypotheses.R
Added P in the Q10 explanation

@@ -49,7 +49,7 @@ names(aggregated_soil_data)
 
 selected_columns <- c(
  "No_trees_ha", "Basal_area", "bulk_den", #"Mean_DBH",  removed becuase many NAs in ROmania "year",
-  "pH",  "Ntot", "Ctot", #"phosphorous", removed becuase many NAs in Romania
+  "pH",  "Ntot", "Ctot", "phosphorous", #removed becuase many NAs in Romania
   "C_stocks", "fungi_biomass", "bacteria_biomass", "actinobacteria_biomass",
   "gram_pos_biomass", "gram_neg_biomass", "total_biomass", "fungi_bact_rate",
   "beta_gluco", "acid_phos", "beta_xylo", "chitise",

### 2024-11-19 Manuscript_draft/Draft.Rmd
Small changes in discussion

@@
 Our model showed that soil temperature and moisture were able to explained more than 50% of variance in soil respiration.
+MORE REFS RELATED TO MODELS (Nissan et al 2023,etc.).
 Because phenologial patterns in deciduous species can affect sensitivity of soil respiration to temperature (Curiel-Yuste et al 2004), in this model we have also included a seasonality scaling.
@@
 Generally, our results confirmed the hypothesis 1, since slash addition increased the respiration rates compared to the treatments where the slash was removed.
-Addition of slash created an artificial litter layer that buffered the effects over soil microclimatic conditions, and provided new habitat and food availability for soil organisms (ref).
+Addition of slash created an artificial litter layer that buffered the effects over soil microclimatic conditions, and provided new habitat and food availability for soil organisms (Pérez-Izquierdo et al.
+in prep).
 These results indicate that increases in heterotrophic respiration due to an increment in organic matter availability by slash addition compensated for the decreases in autotrophic respiration and root-associated microbiota due to tree logging and soil disturbance (Borkhuu et al.
@@
 We argue that this result is probably related to a reduction in chemical recalcitrance, supporting the carbon quality-temperature hypothesis stating that the biochemically recalcitrant C pool is more sensitive to changes in temperature than the labile C pool (Bossata and Agren 1999).
-Moreover, the activation energy correlated positively with the Q10, agreeing with Davidson & Janssens, 2006 and Sierra, 2012 that the higher the activation energy the higher is the relative temperature sensitivity.
+Moreover, the activation energy correlated positively with the Q10, agreeing with Davidson & Janssens (2006) and Sierra (2012) that the higher the activation energy the higher is the relative temperature sensitivity.
 We have also observed a general negative trend of activation energy with time in all management disturbance treatments and all sites.
@@
 In our experiment, it seems that addition of slash have stimulated decomposition of labile C which increased N availability (Ghosh et al., 2019; Shahbaz et al.,2017), and thus decreased the temperature sensitivity.
-Our results indicate that SOM decomposition not only depends on the substrate biochemistry but also on the ability of the microbial community to decompose the available substrate.
+We did not measure chemical recalcitrance per se, but our results suggest that SOM decomposition not only depends on the substrate biochemistry but also on the ability of the microbial community to decompose the available substrate, having a key role in determining Q10 in forest ecosystems.
 This is in agreement with a recent global-scale study stating that soil microorganism are the main players governing the responses of soil heterotrophic respiration to temperature, and secondly substrate quantity and biochemical recalcitrance (Sáez-Sandino et al 2023).
 
-We did not measure chemical recalcitrance per se, but our results suggest that microbial communities might have a key role in determining Q10 in forest ecosystems.
 Lower Q10 values associated to the addition of fresh plant material as slash was related to higher fungal to bateria rates.
@@
 Community data would give more insights about that.
-Surprinsingly, hydrolitic enzymes were not the main variables affecting Q10.
+Surprisingly, hydrolitic enzymes were not the main variables affecting Q10.
 This effect might be masked by the positive correlation of fungal and bacteria biomass with enzymes.

### 2024-11-20 Manuscript_draft/Draft.Rmd
Some updates to the text

@@
 
-## Data pretreatment
-
 ## Validation
@@
 Given the size of the dataset, we considered this approach more reliable than any cross-validation approach.
+
+## Parametric modeling framework
+The parametric models were calibrated in a stratified Bayesian framework, where we assigned independent parameters depending on different groping of the data.
+All subsequent data analysis (model statistics, parameter distributions) have been done working on the whole posterior MCMC (Markov chain Monte Carlo) for each model.
+The model was written in Stan [@carpenter2017stan; @standev2018stancore], run from its R interface [@Rstan].
+The likelihood function was defined as:  
+$$
+    Rh_{meas} \sim \mathcal{N}(Rh_{sim}, \sigma)
+$$
+Where the parameter $\sigma$ was defined as the variance of the whole dataset (0.27).
+All models were explaining the measured respiration normalized by SOC stocks in each experimental plot.
 
@@
 Where the index $j$ corresponds to the treatments and the index $k$ corresponds to the plots (minimal statistical unit), while $(t)$ denotes the variation over time of each forcing variable (temperature, moisture, or days of the year).
-
-## Model calibration
-The parametric model was calibrated in a stratified Bayesian framework, where we assigned independent parameters depending on different groping of the data.   
-The model was written in Stan [@carpenter2017stan; @standev2018stancore], run from its R interface [@Rstan].
-The likelihood function was defined as:  
-$$
-    Rh_{meas} \sim \mathcal{N}(Rh_{sim}, \sigma)
-$$
-Where the parameter $\sigma$ was defined as the variance of the whole dataset (0.27).
 
@@
 ### Seasonality
-In order to correct for eventual seasonality patters, we introduce also a seasonality scaling.
+In order to detect and correct for eventual seasonality patters, we introduced also a seasonality scaling.
 The seasonality scaling functions assume a sine function with period = 365 days, with both amplitude $ampl_j$ (the magnitude) and peak day $peak_j$ (when the function is at its only annual peak) varying as a function of treatments:
@@
 ## Parametric model, time-variant `E_a`
-In order to test the hypothesis of an eventual variation of the activation energy of the decomposing material over time without any assumption on the function of variation, we fitted a version of the model where the $E_a$ parameter was stratified in order to let it vary independently for each treatment and year of sampling.
+In order to test the hypothesis of an eventual variation of the activation energy of the decomposing material over time without any assumption on the function of variation, we fitted a version of the model where the $E_a$ parameter was stratified in order to let it vary independently for each treatment ($j$) and year of sampling ($y$).
+Each sampling year corresponds to multiple measurements (which is necessary to obtain enough information to constrain the temperature function).
+
+$$
+\xi_{temp(t,j)} = A_{k} \cdot e^{\frac{-Ea_{j,y}}{(T(t)+273.15)-T_0)}}
+$$
 
@@
   ),
-  `Stratification Level, const Eₐ` = c(
+  `Stratification Level, time const Eₐ` = c(
     "Plot", 
@@
   ),
-  `Stratification Level, var Eₐ` = c(
+  `Stratification Level, time var Eₐ` = c(
     "Plot", 
@@
 ## Machine learning variance decomposition
-On a more limited subset (only Spain), we tested a more extensive random forest model in order to test the predictive power of other variables on the respiration. The model had structure as:  
+We also tested a more extensive random forest model in order to test the predictive power of other variables on the respiration. The model had structure as:  
 $$
@@
 
-## Temperature sensitivity (Q10)
-
-## Machine learning variance decomposition on enzymes
+A similar approach was used to assess the predictive power of enzyme activities on respiration:
 $$
@@
 $$
+
+## Temperature sensitivity (Q10)
+The Q10 was calculated from the time-invariant process-based model based on each model iteration (so for each element of each MCMC) according to the function:
+$$
+Q_{10} = \left( \frac{R_2}{R_1} \right)^{\frac{10}{T_2 - T_1}}
+$$
+The temperature range was chosen between 15&deg;C.
+and 25&deg;C, while the relative respiration was selected from the predictions from each model realization.
+Since produced a vector of estimated Q10 values as long as each MCMC, we could infer from it also the uncertainty estimates for the Q10 values, which are proportional to the uncertainty estimated by each MCMC.
+
+## Temperature sensitivity relationships with ecological variables
+We considered the relationships between the Q10 (calculated as above), as an indicator of temperature sensitivity, and other variables in the dataset.
+For this we considered the means of the Q10 across each MCMC.
+We first assessed the interactions with a partial least square (PLS) regression (R package `pls`, @Seasholtz1992):
+
+$$
+Q_{10\_\text{averages}} \sim f(
+\text{No\_trees\_ha}, BA, BD, \text{pH}, pH,\\
+C_{tot}, P,  C_{stocks}, \text{fungi\_biomass}, \text{bacteria\_biomass}, \text{actinobacteria\_biomass}, 
+\text{gram\_pos\_biomass}, \text{gram\_neg\_biomass}, \text{total\_biomass}, \text{fungi\_bact\_rate}, \\
+\text{beta\_gluco}, \text{acid\_phos}, \text{beta\_xylo}, \text{chitise}, \text{cellobiohydrolase}, 
+\text{beta\_galacto}, \text{alpha\_gluco}, \text{lipase}, \text{CO2\_flux\_norm}, \text{T1\_soil}, \text{Soil\_moist})
+$$
+
+The importance of predictor variables in the Partial Least Squares (PLS) model was assessed by examining the values of the regression coefficients derived from the model.
+
+We then explored more specific relationships with a linear mixed-effects model (R package `lmer`):
+
+$$
+Q_{10\_\text{averages}} = \beta_0 + \beta_1 \cdot \text{No\_trees\_ha} + \beta_2 \cdot \text{fungi\_bact\_rate} + \\
+\beta_3 \cdot C_{stocks} + \beta_4 \cdot T_{soil} + \beta_5 \cdot  M_{soil} + 
+\beta_6 \cdot treatment + site + \epsilon
+$$
+
+Where:
+
+- \( \beta_0 \): Intercept.
+- \( \beta_1, \beta_2, \dots, \beta_6 \): fixed effect coefficients for the predictors.
+- \( u_{\text{site\_vector}} \): random effect for the site (assumes \( u \sim \mathcal{N}(0, \sigma_u^2) \)).
+- \( \epsilon \): Residual error term (assumes \( \epsilon \sim \mathcal{N}(0, \sigma_\epsilon^2) \)).
 
@@
 ## Benchmark models and parametric model fit
+The two benchmark models were used to set expectations on a possible lower (linear model) and upper (random forest model) bound of the predictive capacity of the data used for the process-based model.
+As expected, the linear model had a much lower predictive capacity than the random forest model, but both showed that the data contained relevant information to predict soil respiration.
 ![Figure 4. Fit of the two benchmark machine learning models, linear (a) and random forest (b)](../Figures/fit_ML_validation.png){width=80%}
+
+The process-based model (constant $E_a$ over time) had, again as expected, a predictive power in between the two (but closer to the random forest model.)
 
@@
 ## Temperature sensitivity (Q10) and parametric model posteriors 
-The temperature sensitivity (Q10) values of SOM mineralization were consistently higher in Romania (2.9 in controls), followed by France (2.7) and Spain (2.5) (Figure S4).
-The activation energy (Ea) followed the same pattern as Q10 (Figure 6a, Figure S4).
+Most of the between treatment and between sites variance was explained by the activation energy parameters of the model.
+The temperature sensitivity (Q10) values of SOM mineralization were consistently higher in Romania (2.9 in controls), followed by France (2.7) and Spain (2.5) (Figure S4), as well as the activation energy (Ea) parameter wiich as expected followed the same pattern as Q10 (Figure 6a, Figure S4).
 In general, the activation energy and Q10 values decreased with slash addition due to higher SOM quality compared with treatments in which slash was removed.
@@
 These results were consistent when the activation energy was analysed as a function of time since disturbance and the activation energy systematically decreased in all treatments and all countries with time since disturbance (Figure 7).
-The photosynthetic activity did not modify the sensitivity of soil respiration to temperature (peak and amplitude factors from the sine-function not affected) (Figure 6b, c).
+Unaccounted processes related to photosynthetic activity did not seem to have a substantial impact on the variance of respiration, since posterior values for the amplitude parameter of the sine function were close to zero (Figure 6b, c).
 However, the amplitude of the sine-function tended to be higher in France (Figure 6c).
-The parameters of soil moisture function (a and b) did not change the response of the model either (Figure 6d, e).
+The parameters of soil moisture function (a and b) did not explain a big part of the treatment and site related variance either (Figure 6d, e), since posteriors were substantially indistinguishable.
+The scaling parameter $A$ (Figure S5) still seems needed to explain part of the variance in the measured respiration, even after normalizing the respiration for initial C stocks.
 
@@
 
-![Figure 7. Posteriors of the activation energy (Ea) parameter varying as a function of treatments and sampling year per country](../Figures/timeseries_Ea_timevariant.png){width=60%}
+![Figure 7. Posteriors of the activation energy (Ea) parameter varying as a function of treatments and sampling year per country](../Figures/timeseries_Ea_timevariant.png){width=90%}
 
@@
 ![Figure 8.
-Principal component regression between edaphic and ecological variables and temperature sensitivity (Q10) analysed by Partial least squares (PLS) regression (upper pannel).
-Variance decomposition of main variables explaining Q10 (lower pannel)](../Figures/Q10_PLS_biplot.png){width=60%}
+partial least squares regression between edaphic and ecological variables and temperature sensitivity (Q10) (upper pannel).
+Variance decomposition of main variables explaining Q10 (lower pannel) based on the values of the regression coefficients](../Figures/Q10_PLS_biplot.png){width=60%}
 
@@
 ## Model extrapolations
-
+Vulnerability to warming, expressed as an increase in mineralization in response to a 5&deg;C increase in average temperature, being a direct modeled consequence of temperature sensitivity, followed the pattern of the temperature sensitivity parameter.
 ![Sensitivity of the organic matter in different treatments to a 5&deg;C increase in temperature, expressed as increase in mineralization from the baseline temperature](../Figures/scenario_extrapolation.png){width=60%}
@@
 # Discussion
-General paragraph
+## General paragraph
 
@@
 
-### Role of temperature and moisture in decomposition (model description)
-Our model showed that soil temperature and moisture were able to explained more than 50% of variance in soil respiration.
+The scaling parameter $A$ was still needed to explain some of the variance in respiration measured at the level of single experimental plots even after normalization by SOC stocks.
+This is likely because of the high variability in SOC content of soils
+
+## Role of temperature and moisture in decomposition 
+The process-based model showed that processes related to soil temperature and moisture were able to explained more than 50% of variance in soil respiration.
 MORE REFS RELATED TO MODELS (Nissan et al 2023,etc.).
-Because phenologial patterns in deciduous species can affect sensitivity of soil respiration to temperature (Curiel-Yuste et al 2004), in this model we have also included a seasonality scaling.
-The seasonal scaling although apparently differentiated slightly between sites, it always presented very low amplitude posteriors, weighting on average only a 5% of the total measured respiration.
+
+### Temperature interactions
+Temperature-related processes, described mainly by the activation energy parameter which determines the temperature sensitivity of the decomposition process, seemed to be the most influential in explaining the between-treatment and between-site variance in decomposition.
+
+### Relationships between organic matter charachteristics, inputs of C to the soil, and temperature sensitivity
+
+### Moisture interactions
+Moisture-related processes, described by the two parameters of the temperature scaling function, did not seem to explain much of the  between-treatment and between-site variability variance in decomposition, since posteriors were largely not differentiated by site and treatment.
+
+### Seasonal patterms
+Because phenological patterns in deciduous species can affect sensitivity of soil respiration to temperature (Curiel-Yuste et al 2004), in this model we have also included a seasonality scaling.
+The seasonal scaling although did not seem to differentiate between sites in the posteriors of both parameters (amplitude and yday of the year of peak), and it always presented very low amplitude posteriors, weighting on average only a 5% of the total measured respiration.
 It was more important in France (between 4.5% and 12.2%) and much less in Spain and Romania (between 2.0% and 3.2%).

### 2024-11-20 Manuscript_draft/Draft.Rmd
Other changes in the draft

@@
 The scaling parameter $A$ was still needed to explain some of the variance in respiration measured at the level of single experimental plots even after normalization by SOC stocks.
-This is likely because of the high variability in SOC content of soils
+This is likely because of the high variability in SOC content of soils, which, particularly in forest soils, can be affected by local processes at a very small spatial scale.
+For example small depressions and local micromorphology, root concentration, localized events such as branch or tree falls, and so on, can substantially affect SOC at a metric or submetric scale.
+This could easily cause discrepancies between the SOC stocks at the site of each respiration measuring collar and the SOC stocks measured by pooling at the level of each experimental plot.
+Since respiration depends first of all on the amoutn of substrate, local stocks can heavily affect it.
+By allowing the $A$ parameter to vary at the level of the experimental plot we could allow for the needed freedom for our parametric models to account for this effect (at the cost of an increased variance of the posterior, but reducing eventual biases).
 
 ## Role of temperature and moisture in decomposition 
-The process-based model showed that processes related to soil temperature and moisture were able to explained more than 50% of variance in soil respiration.
+The process-based model showed that processes related to soil temperature and moisture were able to explain more than 50% of variance in soil respiration.
 MORE REFS RELATED TO MODELS (Nissan et al 2023,etc.).
@@
 Temperature-related processes, described mainly by the activation energy parameter which determines the temperature sensitivity of the decomposition process, seemed to be the most influential in explaining the between-treatment and between-site variance in decomposition.
-
-### Relationships between organic matter charachteristics, inputs of C to the soil, and temperature sensitivity
 
@@
 
-Our results also confirmed the hypothesis 2, since addition of slash decreased the activation energy of SOM decomposition.
-We argue that this result is probably related to a reduction in chemical recalcitrance, supporting the carbon quality-temperature hypothesis stating that the biochemically recalcitrant C pool is more sensitive to changes in temperature than the labile C pool (Bossata and Agren 1999).
-Moreover, the activation energy correlated positively with the Q10, agreeing with Davidson & Janssens (2006) and Sierra (2012) that the higher the activation energy the higher is the relative temperature sensitivity.
+Our results also confirmed the hypothesis 2, since addition of slash in general decreased the activation energy of SOM decomposition, particularly when in a clearc cut (which is already adding fresh inputs to the soil with residuals like leaves and branches falling to the ground as well as roots that start to decompose).
+We argue that this result is probably related to a reduction in chemical recalcitrance due to the increased amount of fresh C inputs, supporting the carbon quality-temperature hypothesis stating that the biochemically recalcitrant C pool is more sensitive to changes in temperature than the labile C pool [@Bosatta.1999], confirmed by other direct measurements [@Lefevre.2014].
+A more recalcitrant substrate likely implies more biochemical steps for its decomposition, each step subjected to the same thermodynamic constraints.
+More steps in the overall oxydative reaction leading also implies a higher chance of one being more limited than the others due to the enzyme characteristics.
+
 We have also observed a general negative trend of activation energy with time in all management disturbance treatments and all sites.
@@
 Still, the growth of new vegetation, within three years from disturbance, did not compensate for the decreases in activation energies in the plots where removal of slash had an effect.
-Additionally, we cannot rule out the fact that other mechanisms could have contributed to low down activation energies with time such as differences between years in primary production (i.e., contribution of autotrophic respiration) or changes in substrate and enzyme diffusivity driven by differences in soil moisture between years (Davinson and Jansen 2022).
+Additionally, we cannot rule out the fact that other mechanisms could have contributed to low down activation energies with time such as differences between years in primary production (i.e., contribution of autotrophic respiration) or changes in substrate and enzyme diffusivity driven by differences in soil moisture between years (Davidson and Jansen 2022).
 In line with this, for example, in the control plots from Spain the activation energy decreased from year zero (2021) to year two (2023) and then it increased a bit, while in disturbed plots it was always descendant.
 Since changes in activation energy due to the growth of new vegetation is not expected in undisturbed control plots and given that 2022 and 2023 were extremely dry in Spain, probably drought exerted a constrained access of enzymes to substrates (Davinson and Jansen 2022).
-This suggest that drought effect may have similar impacts on temperature sensitivity as chemical recalcitrance (Levegre et al 2014).
+This suggest that drought effect may have similar impacts on temperature sensitivity as chemical recalcitrance.
 

### 2024-11-26 Manuscript_draft/Draft.Rmd
Modified the draft slightly, removed the first year on the variable E_0 over time, still units to do

@@
 ```
+
+# Todo
+Remove the first year in Spain, add in M&M why :done, short discussion in M&M, should prob read more about moisture and temperature sensitivity interactions
+Add units
+E_a is not activation energy, add a paragraph about in in the methodology: done
+Change E_a to E_0: done
+Add in discussion something about why we chose this model: done, kind of, just two lines in M&M saying that we chose these because they were readily available, used often in soil science and fitting the purpose.
+I actually had no particular reason to go for Lloyd-Taylor over Arrhenius, and for Moyano it was just that it was readily available to me.
 
@@
 
-## Parametric model, constant `E_a`
+## Parametric model, constant `E_0`
 We defined a parametric model to consider all the three effects on respiration we hypothesized.
@@
 The prior distribution for the term $a$ was assumed as $\sim \mathcal{N}(3.11, 0.1)$ and for the term $b$ was assumed as $\sim \mathcal{N}(2.42, 0.1)$, utilizing as mean values the values in @SierraMueller and setting up conservatives standard deviations.
+The choice of this particular function among others was motivated by its availability and relatively wide usage in soil science, given that due to the calibration process any function with a similar shape would have served the same purpose in this context.
 
@@
 $$
-\xi_{temp(t,j)} = A_{k} \cdot e^{\frac{-Ea_j}{(T(t)+273.15)-T_0)}}
+\xi_{temp(t,j)} = A_{k} \cdot e^{\frac{-E_0_j}{(T(t)+273.15)-T_0)}}
 $$
-The two parameters $A_{k}$ and $E_a$ are represented by Bayesian priors.  
-For the temperature scaling we assumed that the activation energy term $E_a$ (which defines roughly the slope of the relationship) is driven by treatments ($j$) while the linear rescaling term $A_{k}$ is driven by the substrate at the single plot level ($k$).
+The two parameters $A_{k}$ and $E_0$ are represented by Bayesian priors.  
+The $E_0$ parameter in Lloyd and Taylor is a semi-empirical parameter describing biological temperature sensitivity.
+it is very similar to $E_a$ in Arrhenius, but the latter has a clear physical interpretation tied to reaction energy barriers.
+The meaning of the  $E_0$ parameter in Lloyd and Taylor describes the same process, but it's not directly comparable in absolute terms with the $E_a$ parameter in Arrhenius and they are not strictly linearly proportional (they would be when fitting both to the same process, such as in this case, because of the calibration, but still they hjave different physical meanings).
+The choice of this particular function over an Arrhenius relationship was motivated by it being relatively common in soil science studies, and fitting our purpose in the same way.
+For the temperature scaling we assumed that the activation energy term $E_0$ (which defines roughly the slope of the relationship) is driven by treatments ($j$) while the linear rescaling term $A_{k}$ is driven by the substrate at the single plot level ($k$).
 This is because substrate (organic matter) amount is usually highly variable in soil even at small spatial scales, while its quality, which defines its temperature sensitivity, is assumed to follow the inputs quality which are relatively uniform over each treatments.
 The effect of the rescaling term $A_{k}$ is included in the function for temperature scaling to keep the convention of the original formulation of the equation, but it could be brought outside since it just a linear interaction.
-The prior distribution for the term $A$ was assumed as $\sim \mathcal{N}(300, 100)$ and for the term $E_a$ was assumed as $\sim \mathcal{N}(398.5, 50)$, utilizing as mean values the values  in @Lloyd.1994l68 and setting up conservatives standard deviations.
+The prior distribution for the term $A$ was assumed as $\sim \mathcal{N}(300, 100)$ and for the term $E_0$ was assumed as $\sim \mathcal{N}(398.5, 50)$, utilizing as mean values the values  in @Lloyd.1994l68 and setting up conservatives standard deviations.
 
@@
 
-## Parametric model, time-variant `E_a`
-In order to test the hypothesis of an eventual variation of the activation energy of the decomposing material over time without any assumption on the function of variation, we fitted a version of the model where the $E_a$ parameter was stratified in order to let it vary independently for each treatment ($j$) and year of sampling ($y$).
+## Parametric model, time-variant `E_0`
+In order to test the hypothesis of an eventual variation of the activation energy of the decomposing material over time without any assumption on the function of variation, we fitted a version of the model where the $E_0$ parameter was stratified in order to let it vary independently for each treatment ($j$) and year of sampling ($y$).
 Each sampling year corresponds to multiple measurements (which is necessary to obtain enough information to constrain the temperature function).
@@
 $$
+
+In this case we excluded the first year of sampling from the Spain site because this series is incomplete and was performed only in the autumn months, and this might introduce some biases in the analysis.
 
@@
 
-The process-based model (constant $E_a$ over time) had, again as expected, a predictive power in between the two (but closer to the random forest model.)
+The process-based model (constant $E_0$ over time) had, again as expected, a predictive power in between the two (but closer to the random forest model.)
 
-![Figure 5. Fit of the parametric model,  constant `E_a`](../Figures/fit_model_validation.png){width=60%}
+![Figure 5. Fit of the parametric model,  constant `E_0`](../Figures/fit_model_validation.png){width=60%}
 
@@
 
-### Seasonal patterms
+### Seasonal patterns
 Because phenological patterns in deciduous species can affect sensitivity of soil respiration to temperature (Curiel-Yuste et al 2004), in this model we have also included a seasonality scaling.
-The seasonal scaling although did not seem to differentiate between sites in the posteriors of both parameters (amplitude and yday of the year of peak), and it always presented very low amplitude posteriors, weighting on average only a 5% of the total measured respiration.
+The seasonal scaling although did not seem to differentiate between sites in the posteriors of both parameters (amplitude and day of the year of peak), and it always presented very low amplitude posteriors, weighting on average only a 5% of the total measured respiration.
 It was more important in France (between 4.5% and 12.2%) and much less in Spain and Romania (between 2.0% and 3.2%).
@@
 
-Our results also confirmed the hypothesis 2, since addition of slash in general decreased the activation energy of SOM decomposition, particularly when in a clearc cut (which is already adding fresh inputs to the soil with residuals like leaves and branches falling to the ground as well as roots that start to decompose).
+Our results also confirmed the hypothesis 2, since addition of slash in general decreased the activation energy of SOM decomposition, particularly when in a clearcut (which is already adding fresh inputs to the soil with residuals like leaves and branches falling to the ground as well as roots that start to decompose).
 We argue that this result is probably related to a reduction in chemical recalcitrance due to the increased amount of fresh C inputs, supporting the carbon quality-temperature hypothesis stating that the biochemically recalcitrant C pool is more sensitive to changes in temperature than the labile C pool [@Bosatta.1999], confirmed by other direct measurements [@Lefevre.2014].
@@
 In line with this, for example, in the control plots from Spain the activation energy decreased from year zero (2021) to year two (2023) and then it increased a bit, while in disturbed plots it was always descendant.
-Since changes in activation energy due to the growth of new vegetation is not expected in undisturbed control plots and given that 2022 and 2023 were extremely dry in Spain, probably drought exerted a constrained access of enzymes to substrates (Davinson and Jansen 2022).
+Since changes in activation energy due to the growth of new vegetation is not expected in undisturbed control plots and given that 2022 and 2023 were extremely dry in Spain, probably drought exerted a constrained access of enzymes to substrates (Davidson and Jansen 2022).
 This suggest that drought effect may have similar impacts on temperature sensitivity as chemical recalcitrance.

### 2024-11-26 Presentations/whattodo.Rmd
Modified the draft slightly, removed the first year on the variable E_0 over time, still units to do

---
title: "What to do?"
author: "Lorenzo Menichetti"
date: "2024-11-20"
output:
  beamer_presentation: default
  ioslides_presentation: default
  slidy_presentation: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## What we have done so far 1
You know already I fitted with a stratified Bayesian structure a model combining temperature, moisture and seasonality to explain respiration. Very roughly:  
$$
Rh(t) =  \xi_{temp(t,j, k)} \cdot \xi_{moist(t,j)} + \xi_{sin(t,j)}
$$

Where the index $j$ corresponds to the treatments and the index $k$ corresponds to the plots (minimal statistical unit), while $(t)$ denotes the variation over time of each forcing variable (temperature, moisture, or days of the year). These are the stratification factors.

## Why Bayesian?
It is crucial to fit all these scaling functions at the same time because:

* they are clearly interacting (linearly) in the model
* there are correlations. Moisture is ALSO dependent on temperature, if we fit in sequence, first temperature and then a moisture scaling for the remaining variance, part of this interaction might spill in the temperature scaling, which will describe also part of the variance due to moisture

## Why stratified?
With stratified sampling I can assign different freedom to each parameter, tuning them based on our hypotheses. For example I can assign the scaling to vary by single experimental unit (because of local SOC variance), but then activation energy to a site x treatment factor, assuming it is influenced by treatments.  
This produces a relatively robust estimate of all the parameters involved with all their interactions, and with their uncertainty (which is crucial for robustness).

## What we have done so far 2: activation energy stratified by treatment x site
  ![](../Figures/posteriors_boxplots.png){width=70%}

## What we have done so far 3 - bonus: time as independent stratification factor
  ![](../Figures/timeseries_Ea_timevariant.png){width=80%}

This more to show what we can do... I don't think this is necessarily useful for you. It might.

## Why Arrhenius (type)? A critique of Tumi et al. 2011.
Because I believe in it. And in general I do not believe in any model. But Arrhenius theory is very well tested, that's something we know.  
We can safely assume that temperature interactions follow that parametric function, in the range of temperatures we consider, and this is information.

Using an empirical function, like the Gaussian in Tuomi, discards uselessly information we have. Besides opening for artifacts (because the Gaussian is symmetric, and it is totally out of reality for the descending part).  
Tomi et al. is based on a strict interpretation of Bayes' factor, but the fitness of their function was the same than a LLoyd-Taylor. It was just "better" in terms of Bayes' factor, which is also influenced by their prior choice (and they used very broad uniforms).

## Moisture
I used Moyano et al., but Boris has a much better function (more elegant, more process-based and more compact than a polynomial fit), here I suggest we join forces.

## What could you do with it?
Easiest would be that:

* first I fit for you a very similar model, giving you estimates of temperature and mositure parameters for each treatment (or treatment and site, using data close to the microbial sampling times) estimated based on temperature, moisture and seasonality correction. You get full MCMCs, so with a nonparametric error model information attached.
* then you could develop a function based on these data relating both temperature and moisture params to diversity or key species. Like this we also test where to add that variance, does it make sense to assume an activation energy varying with diversity, or maybe the moisture scaling varies?
* then we could even implement those functions in the same model and see what changes, if anything

## Any other function would work for me
Even if I would keep clear from empirical functions (like Tuomi et al. 2011), I don't have particular preferences. Do you have some better function? We use it.  
Do you want to use a reversible denaturation for Temperature, and a diffusion-based one for moisture? Well, that would be wonderful, maybe an overkill and I like overkills. I like the elegance of the Daniels' temperature model, and even if I don't understand diffusion-based models I would be very happy to step up my game and use them (and Marleen knows them).

My suggestion is just to rely on a Bayesian stratified sampling and to explore all the model parameter space, using normal priors and constraining them with as much information as we can find in the literature (uniforms are nice to be on the safe side but I suggest we live dangerously).

## Where are the data?
I am not sure. I am working on this, I think we might have the dataset with the respiration measurements in 1, maybe 2, maybe 3 months max, connected with the dataset from Prague.   
We won't have SOC stocks there, just respiration,temperature and sometimes moisture. Once these data are ready I might manage to fit the model in 1-2 months, hopefully.

Moisture is missing in some cases, it's not there so often.
I think we can still do the same work, just using a calobrated, inferred, moisture scaling for the sites where it is missing, this will just increase the uncertainty in those sites. In the case of our calibration it wasn't so crucial.

### 2024-11-26 data_analysis_sections/parametric_model_fitting_time_variant.R

@@ -238,13 +238,13 @@ for (j in 1:length(sites)) {
     treat_data <- data_site[data_site$treat == treatments[i], ]
 
     # Add the confidence interval bands (polygon)
-    polygon(c(treat_data$year, rev(treat_data$year)),
-            c(treat_data$CI_Upper, rev(treat_data$CI_Lower)),
+    polygon(c(treat_data[treat_data$year != 0,]$year, rev(treat_data[treat_data$year != 0,]$year)),
+            c(treat_data[treat_data$year != 0,]$CI_Upper, rev(treat_data[treat_data$year != 0,]$CI_Lower)),
             col = add.alpha(palette_treat[i+color_index], 0.25), border = NA)
 
     # Add the points and lines
-    points(treat_data$year, treat_data$Mean, col = palette_treat[i+color_index], pch = i)
-    lines(treat_data$year, treat_data$Mean, col = palette_treat[i+color_index])
+    points(treat_data[treat_data$year != 0,]$year, treat_data[treat_data$year != 0,]$Mean, col = palette_treat[i+color_index], pch = i)
+    lines(treat_data[treat_data$year != 0,]$year, treat_data[treat_data$year != 0,]$Mean, col = palette_treat[i+color_index])
   }
   color_index = color_index + 4
 

### 2024-11-26 data_analysis_sections/parametric_model_fitting.R
Updated version of the plots (new data)

@@ -181,7 +181,7 @@ par(mfrow=c(1,2))
 E_0_densities_indA_Moy <- list()
 E_0_box_indA_Moy <- list()
 plot( density(post_bytreat_indA_Moy$Ea[,1]), xlim=c(range(post_bytreat_indA_Moy$Ea)[1]*0.9, range(post_bytreat_indA_Moy$Ea)[2]*1.1),
-      ylim=c(0,0.07), xlab = expression('E'[a]), col=NA, main="")
+      ylim=c(0,0.07), xlab = expression('E'[0]), col=NA, main="")
 for(i in 1:stan_data$Tr){
   E_0_densities_indA_Moy[[i]] <- density(post_bytreat_indA_Moy$Ea[,i])
   E_0_box_indA_Moy[[i]] <- E_0_densities_indA_Moy[[i]]$x

### 2024-11-26 data_analysis_sections/parametric_model_fitting_time_variant.R

@@ -46,7 +46,7 @@ boxplot(stan_data$resp ~ stan_data$treat_year)
 
 
 # define the number of runs of the MCMC
-N_RUNS = 5000
+N_RUNS = 6000
 
 # Determine the number of cores available
 num_cores <- detectCores()
@@ -230,7 +230,7 @@ for (j in 1:length(sites)) {
 
   # Create an empty plot
   plot(data_site$year, data_site$Mean, type = "n", ylim = yrange, xlim=c(0,3),
-       xlab = "Year", ylab = expression('E'[a]), main = paste(sites_plot_names[j]))
+       xlab = "Year", ylab = expression('E'[0]), main = paste(sites_plot_names[j]))
 
   # Loop over each treatment to add points, lines, and CI bands
   treatments <- unique(data_site$treat)

### 2024-11-26 master_data_analysis.R

@@ -19,7 +19,7 @@ set.seed(123)
 
 # edaphics <- read.csv("edaphics.csv")
 #data <- read.csv("Database_co2_28_05_2024.csv")
-data <- read_xlsx("Database_co2_18_10_2024_format_Lorenzo.xlsx", sheet=1)
+data <- read_xlsx("Database_co2_26_11_2024_format_Lorenzo.xlsx", sheet=1)
 str(data)
 #soil_data <- read_xlsx("Soil_data_Lorenzo_03_09_2024_enzymes.xlsx", sheet=1)
 soil_data <- read_xlsx("Soil_data_Lorenzo_14_11_2024_enzymes.xlsx", sheet=1)

### 2024-12-09 Manuscript_draft/Draft.Rmd
Introduction

@@
 
-Forest management have the potential to affect respiration rates by modifying the microclimate.
-Removal of trees by forest logging increases the exposure of soil to temperature fluctuations and alter the soil water content either by decreasing water through soil evaporation or by increasing it by reducing the transpiration of water by trees (Zhang et al 2018).
-The degree of the response of soil will be dependent on the severity of the forest intervention.
-Thus, more severe forestry intervention such as clear-cuts may increase soil moisture, enhance heat and frost damage (Kovaks et al 2019; Zhang et al 2018) and it seems to reduce respiration rates (Yang et el 2022; Zhang et al 2018).
-Less severe interventions such as thinning, have shown to increase respiration in broadleaf forests in a meta-analysis (Zhang et al 2018).
-Negative effects of forest management on soil could be buffered by the addition of forest residues on soil (i.e., slash addition), which can improve the microclimatic conditions, nutrient fertility and biodiversity (ref).
-We do not know whether the response of soil CO2 effluxes to the microclimatic conditions (i.e.
-soil temperature and moisture) differs among different forest management strategies.
-Understanding the sensitivity of respiration to microclimate under different forest management strategies is crucial in a context of climate-smart forestry (CSF) development.
+Forest disturbances have the potential to affect respiration rates and hence carbon cycling by modifying the microclimate as well as the decomposer communities (ref).
+Removal of trees by forest logging increases the exposure of soil to temperature fluctuations and alter the soil water content either by decreasing water through soil evaporation or by increasing it due to a reduction of the transpiration of water by trees (Zhang et al 2018).
+The degree of the response of soil is dependent on the severity of the forest disturbance.
+Thus, more severe disturbances such as clear-cuts seems to increase soil moisture, enhance heat and frost damage (Kovaks et al 2019; Zhang et al 2018) and reduce respiration rates (Yang et el 2022; Zhang et al 2018).
+Less severe disturbances such as thinning, have shown to increase respiration in broadleaf forests in a meta-analysis (Zhang et al 2018).
+Autotrophic respiration is very sensitive to the decrease in photosynthetic activity due to the loss of trees (Levy-Varon et al.
+2012, 2014;Moore et al.
+2013; Borkhuu et al.
+2015).
+Instead, increases in heterotrophic respiration due to the increases of saprotrophic activity after the disturabcne
+
+Negative effects of forest disturbances on soil could be buffered by the addition of forest residues on soil (i.e., slash addition), which can improve the microclimatic conditions, nutrient fertility and biodiversity (Pérez-Izquierdo et al.
+2025).
+There is a lack of knowledge about the response of soil CO2 fluxes to the microclimatic conditions (i.e.
+soil temperature and moisture) under different forest disturbances and if this response can be modulated by effective post-disturbance management strategies.
+Understanding the sensitivity of respiration to microclimate under different forest disturbances is crucial in a context of climate-change-mediated increased disturbances.
 

### 2024-12-09 Manuscript_draft/Draft.Rmd
Remove all the management terms

@@
 # Introduction
-The soil is one of the largest terrestrial reservoirs of carbon, containing approximately three times more carbon than the atmosphere or the vegetation on Earth (IPCC, 2021).
-Soil carbon is released to the atmosphere as carbon dioxide (CO2) through decomposition of organic matter by microbes (heterotrophic respiration) and root respiration (autotrophic respiration).
-This process (i.e., soil respiration) constitutes the main carbon efflux into the atmosphere (ref).
-Under the new climate change scenarios increases of CO2 efflux from soils are expected (Schlesinger and Andrews, 2000), this can imply a positive feedback to future warming if stored carbon is mobilized and carbon input to soil does not increase (Lamberty&Thomson, 2010).
+The soil is one of the largest terrestrial reservoirs of carbon (C), containing approximately three times more C than the atmosphere or the vegetation on Earth (IPCC, 2021).
+Soil C is released to the atmosphere as carbon dioxide (CO2) through decomposition of organic matter by microbes (heterotrophic respiration) and root respiration (autotrophic respiration).
+This process (i.e., soil respiration) constitutes the main C efflux into the atmosphere (ref).
+Under the new climate change scenarios increases of CO2 efflux from soils are expected (Schlesinger and Andrews, 2000), this can imply a positive feedback to future warming if stored C is mobilized and C input to soil does not increase (Lamberty&Thomson, 2010).
 Because soil respiration is highly sensitive to soil temperature and soil moisture (Nissan et al.
@@
 
-Forest disturbances have the potential to affect respiration rates and hence carbon cycling by modifying the microclimate as well as the decomposer communities (ref).
+Forest disturbances have the potential to affect respiration rates and hence C cycling by modifying the microclimate, nutrients as well as the decomposer communities (ref).
 Removal of trees by forest logging increases the exposure of soil to temperature fluctuations and alter the soil water content either by decreasing water through soil evaporation or by increasing it due to a reduction of the transpiration of water by trees (Zhang et al 2018).
@@
 Less severe disturbances such as thinning, have shown to increase respiration in broadleaf forests in a meta-analysis (Zhang et al 2018).
-Autotrophic respiration is very sensitive to the decrease in photosynthetic activity due to the loss of trees (Levy-Varon et al.
+After a disturbance, imbalances between autotrophic and heterotrophic respiration might happen (ref).
+The autotrophic respiration is very sensitive to the decrease in photosynthetic activity due to the loss of trees (Levy-Varon et al.
 2012, 2014;Moore et al.
@@
 2015).
-Instead, increases in heterotrophic respiration due to the increases of saprotrophic activity after the disturabcne
+Instead, increases in heterotrophic respiration due to the increases in saprotrophic activity after the disturbance are normal (ref).
 
@@
 2025).
+Moreover, addition of slash in turn might decrease the temperature sensitivity of soil respiration, making it less vulnerable to climate-induced temperature raise.
 There is a lack of knowledge about the response of soil CO2 fluxes to the microclimatic conditions (i.e.
@@
 Here, we hypothesized that:
-(1) logging of trees will decrease the respiration rates because of the strong decrease in vegetation activity and microbiota associated.
+(1) canopy disturbances will decrease the respiration rates because of the strong decrease in vegetation activity and microbiota associated.
 This decrease will be compensated by addition of slash which provides labile organic matter, increasing the heterotrophic respiration.
-(2) soil respiration would be more dependent on microclimatic variables (soil temperature and moisture), in the management disturbance treatments where the vegetation was removed and where the litter layer (slash) was removed, by depletion of labile organic matter.
+(2) soil respiration would be more dependent on microclimatic variables (soil temperature and moisture), in the disturbance treatments where the vegetation was removed and where the litter layer (slash) was removed, by depletion of labile organic matter.
 Likewise, the sensitivity of respiration to microclimate will decrease as the time passes since the disturbance because of new fresh C input from the growth of understory vegetation
@@
 
-To test these hypotheses, we have recorded data on soil respiration and microclimate every month in three replicated Quercus forests across Europe subjected to different forest management strategies during three years.
+To test these hypotheses, we have recorded data on soil respiration and microclimate every month in three replicated Quercus forests across Europe subjected to different intensities of forest canopy disturbances during three years.
 We have analysed the sensitivity of soil respiration to microclimate (i.e., soil temperature and soil moisture) by constructing a mechanistic parametric model and compared to a non-parametric model (i.e., random forest).
-We have also explored the main drivers of variation in Q10 across different forests and forest managements disturbances.
+We have also explored the main drivers of variation in Q10 across different forests and forest canopy disturbances.
 
@@
 ## Study sites
-The study sites are located in Gamiz, Northern Spain (42°49'2.32"N; 2°37'10.67"W), Saint-Christol-d’Albion, South of France (xxx) and Dumbravita, East Romania (42°49'2.32"N; 2°37'10.67"W).
+The study sites are located in Gamiz, Northern Spain (42°49'2.32"N; 2°37'10.67"W), Saint-Christol-d’Albion, South of France (44°02'58.0"N 5°32'35.8"E) and Dumbravita, East Romania (42°49'2.32"N; 2°37'10.67"W).
 According to Köppen-Geiger climatic classification, the climate in Gamiz is Oceanic, with cold and rainy winters and warm summers.
@@
 In each site, three stands of 0.12 ha each and replicated in four blocks were delimited.
-The stands corresponded to a control and two canopy disturbance treatments, i.e., 50% thinning and clear-cut (Fig.
+The stands corresponded to a control and two canopy disturbance treatments, i.e., 50% canopy removal (CD50) and 100 % canopy removal (CD100)  (Fig.
 S1).
-The stands from disturbed treatments (thinning and clear-cut) were in turn divided in two: on half of the stands the slash resulting from the logging (leaves and twigs) was crashed and added on top of the soil and on the other half, the woody debris resulting from logging was removed.
-This made a total of 5 treatments (control, clear-cut with and without slash, thinning with and without slash) per block (Fig.
+The stands from disturbed treatments (CD50 and CD100) were in turn divided in two: on half of the stands the slash resulting from the logging (leaves and twigs) was crashed and added on top of the soil and on the other half, the woody debris resulting from logging was removed.
+This made a total of 5 treatments (control, CD50 with and without slash, CD100 with and without slash) per block (Fig.
 S1).
-In each of the 5 treatments, two plots were established per block making a total of 40 plots (4 blocks x 5 treatments x 2 replicates per block; Fig S1).
+In each of the 5 treatments, two plots were established making a total of 40 plots (4 blocks x 5 treatments x 2 replicates; Fig S1).
 In turn, in each plot a transect of 2m length was established for monitoring (Fig.
@@
 ## Soil respiration 
-To measure soil CO2 efflux (i.e., soil respiration), three rings of 10 cm diameter and 8 cm high were put along the 2 m transect in each of the monitoring plots (Fig.
+To measure soil CO2 efflux (i.e., soil respiration), three rings of 10 cm diameter and approximately 8 cm high were put along the 2 m transect in each of the monitoring plots (Fig.
 S1).
@@
 ## Forest stand characteristics
-Within each of the control and thinned stands (n=8), we recorded the number of trees to calculate tree density (No.
+Within each of the control and CD50 stands (n=8), we recorded the number of trees to calculate tree density (No.
 tree per ha) and measured tree diameter at breast height (DBH; 1.3 m) to calculate basal area.
@@
 The pH was measured in distilled water 1:10 (w/v).
-The carbon (C) and nitrogen (N) content of the soil was measured using an elemental analyser: C content was measured by sulfochromic oxidation (ISO 14235), and N content was estimated by sulfuric acid mineralization with the addition of selenium and sodium sulphate and conversion to ammonium ions (ISO 11261), which were measured with the Segmented Flow Analyzer (SFA) Skalar.
+The C and nitrogen (N) content of the soil was measured using an elemental analyser: C content was measured by sulfochromic oxidation (ISO 14235), and N content was estimated by sulfuric acid mineralization with the addition of selenium and sodium sulphate and conversion to ammonium ions (ISO 11261), which were measured with the Segmented Flow Analyzer (SFA) Skalar.
 Available phosphorus (P) was determined after extraction with malachite green according to the protocol by Ohno & Zibilske (1991).
@@
 
-We included in this model tree density in plants ha$^{-1}$ ($trees$), average diameter at breast height ($DBH$), basal area ($BA$),  pH, bulk density ($BD$), phosophorous ($P$), total nitrogen ($N}$), carbon stocks ($C_{stocks$), fungal biomass ($biom_{fungi}$), bacterial biomass ($biom_{bact}$), actinomycetes biomass ($biom_{act}$), gram+ bacterial biomass ($biom_{gram+}$), gram- bacterial biomass ($biom_{gram-}$), total biomass ($biom_{tot}$), fungi:bacteria ratio, soil temperature ($T_{soil}$), soil moisture  ($M_{soil}$) and month of the year.
+We included in this model tree density in plants ha$^{-1}$ ($trees$), average diameter at breast height ($DBH$), basal area ($BA$),  pH, bulk density ($BD$), phosophorous ($P$), total nitrogen ($N}$), C stocks ($C_{stocks$), fungal biomass ($biom_{fungi}$), bacterial biomass ($biom_{bact}$), actinomycetes biomass ($biom_{act}$), gram+ bacterial biomass ($biom_{gram+}$), gram- bacterial biomass ($biom_{gram-}$), total biomass ($biom_{tot}$), fungi:bacteria ratio, soil temperature ($T_{soil}$), soil moisture  ($M_{soil}$) and month of the year.
 This model was tested only on a subset of the dataset (only the sites in Spain), for which all the additional variables were available.
@@
 Spain showed intermediate values of both tree density and DBH (Figure 1).
-For soil variables, in the untouched forest (control), the bulk density was the highest in Spain and the lowest in France.
+For soil variables, in the untouched forest (i.e., control), the bulk density was the highest in Spain and the lowest in France.
 Spain had the highest pH while France and Romania had similar ranges of pH.
@@
 Boxplots of stand characteristics and soil physical-chemical properties per disturbance treatment and study site.
-Red shades = France, green shades = Romania, blue shades = Spain](../Figures/edaphic_boxplots.png){width=80%}
+Red shades = France, green shades = Romania, blue shades = Spain.
+CD50 = 50% canopy disturbance; CD100 = 100% canopy disturbance; +S = with slash; -S = without slash.](../Figures/edaphic_boxplots.png){width=80%}
 
@@
 The disturbance treatments barely affected the soil properties in Romania (Figure 1).
-In France, the bulk density increased as the intensity of the disturbance did (i.e., from thinning with slash to clear-cut no slash) and in Spain it significantly increased in all the disturbance treatments compared with control (Figure 1).
-The pH only significantly increased in both clear-cut treatments in Spain.
-Moreover, in Spain the N content decreased in the thinning and clear-cut without slash (Figure 1).
-In France, compared with controls, C content increased in all the disturbance treatment, but in the clear-cut without slash that decreased (Figure 1).
-In Spain, C content only decreased in the thinning-without slash treatment (Figure 1).
-Similar to soil properties, microbial biomass, fungal biomass and bacterial biomass did not change with the disturbance treatment in Romania.
-In Spain, these variables significantly decreased in the disturbance treatments where the slash was removed and in France only in the clear-cut treatment where the slash was removed (Figure 2).
-The disturbance treatments barely affected enzymatic activities in Romania and Spain, but generally decreased in clear-cut where slash was removed in France (Figure 2).
+In France, the bulk density increased as the intensity of the disturbance did (i.e., from CD50 with slash to CD100 no slash) and in Spain it significantly increased in all the disturbance treatments compared with control (Figure 1).
+The pH only significantly increased in both CD100 treatments in Spain.
+Moreover, in Spain the N content decreased in the CD50 and CD100 without slash (Figure 1).
+In France, compared with controls, C content increased in all the disturbance treatment, but in the CD100 without slash that decreased (Figure 1).
+In Spain, C content only decreased in the CD50 without slash treatment (Figure 1).
+Similar to soil properties, microbial biomass, fungal biomass and bacterial biomass did not change with the disturbance treatments in Romania.
+In Spain, these variables significantly decreased in the disturbance treatments where the slash was removed and in France only in the CD100 treatment where the slash was removed (Figure 2).
+The disturbance treatments barely affected enzymatic activities in Romania and Spain, but generally decreased in the CD100 where slash was removed in France (Figure 2).
 
@@
 Boxplots of soil microbial biomass and enzymatic activities per disturbance treatment and study site.
-Red shades = France, green shades = Romania, blue shades = Spain](../Figures/enzymes_boxplots.png){width=80%}
+Red shades = France, green shades = Romania, blue shades = Spain.
+CD50 = 50% canopy disturbance; CD100 = 100% canopy disturbance; +S = with slash; -S = without slash.](../Figures/enzymes_boxplots.png){width=80%}
 
@@
 In general, control plots showed the lowest thermal amplitude in all sites (i.e., with higher temperatures in winter and lower in summer).
-In Spain, the soil temperature increased as the intensity of the tree removal, mainly in summer, but the addition of slash buffered these increases.
-In France, the main effect on soil temperature was the tree removal, especially in summer and this was independent on the addition of slash.
+In Spain, the soil temperature increased as the intensity of the canopy disturbance, mainly in summer, but the addition of slash buffered these increases.
+In France, the main effect on soil temperature was the canopy disturbance, especially in summer and this was independent on the addition of slash.
 Soil temperature was not affected by disturbance and post-disturbance treatments in Romania.
@@
 Soil moisture fluctuated inversely with temperature.
-In Spain, soil moisture was higher in the clear-cut with and without slash addition, except in the driest periods (from April to August).
-Similarly, in Romania, clear-cut increased soil moisture independently of slash addition.
-Contrarily, in France, soil moisture tended to decrease in the clear-cut treatment with and without slash (Figure 3b).
-Both in Spain and France, removal of slash decreased respiration rates, independently of tree cutting intensity, and thinning coupled with slash addition kept the respiration rates as the same level as controls.
-In Romania, effects of the disturbances were lower, however clear-cut increased respiration rates, especially when slash was added to the soil.
-At the end of the monitoring period, three years for Spain and two years for France and Romania after disturbance, respiration rates increased in the clear-cut with slash addition in the three study sites (Figure 3c).
+In Spain, soil moisture was higher in the CD100 with and without slash addition, except in the driest periods (from April to August).
+Similarly, in Romania, CD100 increased soil moisture independently of slash addition.
+Contrarily, in France, soil moisture tended to decrease in the CD100 treatment with and without slash (Figure 3b).
+Both in Spain and France, removal of slash decreased respiration rates, independently of canopy disturbance intensity, and CD50 coupled with slash addition kept the respiration rates as the same level as controls.
+In Romania, effects of the disturbances were lower, however CD100 increased respiration rates, especially when slash was added to the soil.
+At the end of the monitoring period, three years for Spain and two years for France and Romania after disturbance, respiration rates increased in the CD100 with slash addition in the three study sites (Figure 3c).
 
@@
 
-![Figure 5. Fit of the parametric model,  constant `E_0`](../Figures/fit_model_validation.png){width=60%}
+![Figure 5.
+Fit of the parametric model,  constant `E_0`.
+Red shades = France, green shades = Romania, blue shades = Spain.
+CD50 = 50% canopy disturbance; CD100 = 100% canopy disturbance; +S = with slash; -S = without slash.](../Figures/fit_model_validation.png){width=60%}
 
@@
 Most of the between treatment and between sites variance was explained by the activation energy parameters of the model.
-The temperature sensitivity (Q10) values of SOM mineralization were consistently higher in Romania (2.9 in controls), followed by France (2.7) and Spain (2.5) (Figure S4), as well as the activation energy (Ea) parameter wiich as expected followed the same pattern as Q10 (Figure 6a, Figure S4).
-In general, the activation energy and Q10 values decreased with slash addition due to higher SOM quality compared with treatments in which slash was removed.
-This was especially evident in Spain for both thinning and clear-cut and in the clear treatment in France.
+The temperature sensitivity (Q10) values of SOM mineralization were consistently higher in Romania (2.9 in controls), followed by France (2.7) and Spain (2.5) (Figure S4), as well as the activation energy (Ea) parameter which as expected followed the same pattern as Q10 (Figure 6a, Figure S4).
+In general, the activation energy and Q10 values decreased with slash addition compared with treatments in which slash was removed.
+This was especially evident in Spain for both CD50 and CD100 and in the CD100 treatment in France.
 The activation energy and Q10 in the treatments where slash was added did not differ from controls (Figure 6a, Figure S4).
-These results were consistent when the activation energy was analysed as a function of time since disturbance and the activation energy systematically decreased in all treatments and all countries with time since disturbance (Figure 7).
+These results were consistent when the activation energy was analysed as a function of time since disturbance, and the activation energy systematically decreased in all disturbance treatments and all countries with time since disturbance (Figure 7).
 Unaccounted processes related to photosynthetic activity did not seem to have a substantial impact on the variance of respiration, since posterior values for the amplitude parameter of the sine function were close to zero (Figure 6b, c).
@@
 
-Hydrologic leaching of dissolved and particulate carbon compounds can also be important outputs of C from the system (Davinson and Jansen 2006), that could explained the low carbon content in Romania (sandy soil with low retention capacity).
+Hydrologic leaching of dissolved and particulate C compounds can also be important outputs of C from the system (Davinson and Jansen 2006), that could explained the low C content in Romania (sandy soil with low retention capacity).
 
@@
 This is likely because of the high variability in SOC content of soils, which, particularly in forest soils, can be affected by local processes at a very small spatial scale.
-For example small depressions and local micromorphology, root concentration, localized events such as branch or tree falls, and so on, can substantially affect SOC at a metric or submetric scale.
+For example, small depressions and local micromorphology, root concentration, localized events such as branch or tree falls, and so on, can substantially affect SOC at a metric or submetric scale.
 This could easily cause discrepancies between the SOC stocks at the site of each respiration measuring collar and the SOC stocks measured by pooling at the level of each experimental plot.
-Since respiration depends first of all on the amoutn of substrate, local stocks can heavily affect it.
+Since respiration depends first of all on the amount of substrate, local stocks can heavily affect it.
 By allowing the $A$ parameter to vary at the level of the experimental plot we could allow for the needed freedom for our parametric models to account for this effect (at the cost of an increased variance of the posterior, but reducing eventual biases).
@@
 in prep).
-These results indicate that increases in heterotrophic respiration due to an increment in organic matter availability by slash addition compensated for the decreases in autotrophic respiration and root-associated microbiota due to tree logging and soil disturbance (Borkhuu et al.
+These results indicate that increases in heterotrophic respiration due to an increment in organic matter availability by slash addition compensated for the decreases in autotrophic respiration and root-associated microbiota due to the canopy and soil disturbance (Borkhuu et al.
 2015).
@@
 
-Our results also confirmed the hypothesis 2, since addition of slash in general decreased the activation energy of SOM decomposition, particularly when in a clearcut (which is already adding fresh inputs to the soil with residuals like leaves and branches falling to the ground as well as roots that start to decompose).
-We argue that this result is probably related to a reduction in chemical recalcitrance due to the increased amount of fresh C inputs, supporting the carbon quality-temperature hypothesis stating that the biochemically recalcitrant C pool is more sensitive to changes in temperature than the labile C pool [@Bosatta.1999], confirmed by other direct measurements [@Lefevre.2014].
+Our results also confirmed the hypothesis 2, since addition of slash in general decreased the activation energy of SOM decomposition, particularly in CD100 (which is already adding fresh inputs to the soil with residuals like leaves and branches falling to the ground as well as roots and mycelium that start to decompose).
+We argue that this result is probably related to a reduction in chemical recalcitrance due to the increased amount of fresh C inputs, supporting the C quality-temperature hypothesis stating that the biochemically recalcitrant C pool is more sensitive to changes in temperature than the labile C pool [@Bosatta.1999], confirmed by other direct measurements [@Lefevre.2014].
 A more recalcitrant substrate likely implies more biochemical steps for its decomposition, each step subjected to the same thermodynamic constraints.
@@
 
-We have also observed a general negative trend of activation energy with time in all management disturbance treatments and all sites.
+We have also observed a general negative trend of activation energy with time in all disturbance treatments and all sites.
 This result also gives support to the C quality temperature hypothesis since fresh inputs from the new vegetation low down the temperature sensitivity of C mineralization as these inputs are more labile to decomposed (Lefevre et al 2014).
@@
 Our analyses revealed the three variables that explained most of the variation of Q10 were related to i) the quality/quantity of substrate (i.e., nitrogen content), ii) the microorganisms, in particular, fungal to bacteria ratio and iii) the soil temperature.
-Nitrogen fertilization experiments have shown that nitrogen availability decreases the Q10 (Mo et al 2008; Zhang et al 2019), probably related to changes in microbial carbon use efficiency (Manzoni et al 2012).
+Nitrogen fertilization experiments have shown that nitrogen availability decreases the Q10 (Mo et al 2008; Zhang et al 2019), probably related to changes in microbial C use efficiency (Manzoni et al 2012).
 In our experiment, it seems that addition of slash have stimulated decomposition of labile C which increased N availability (Ghosh et al., 2019; Shahbaz et al.,2017), and thus decreased the temperature sensitivity.
@@
 2012, Tlaskal et al 2016).
-In contrast, bacteria seems to preferentially gain carbon from live or death fungal biomass and increase at later stages of litter decomposition (Tlaskal et al 2016, Lindahl et al 2010).
-Since ectomycorrhizal fungal mycelium comprise most of the fungal biomass in forests (Wallander et al 2001), a strong reduction in ectomycorrhizal fungal biomass is expected in all the tree removal disturbance treatments, thus decreasing the fungal to bacteria ratio.
+In contrast, bacteria seems to preferentially gain C from live or death fungal biomass and increase at later stages of litter decomposition (Tlaskal et al 2016, Lindahl et al 2010).
+Since ectomycorrhizal fungal mycelium comprise most of the fungal biomass in forests (Wallander et al 2001), a strong reduction in ectomycorrhizal fungal biomass is expected in all the canopy disturbance treatments, thus decreasing the fungal to bacteria ratio.
 As mentioned, this decrease in ectomycorrhizal fungal biomass might be compensated by a rapid growth of saprotrophic fungal species where slash is added.

### 2024-12-10 Manuscript_draft/Draft.Rmd
small things

@@
 Because soil respiration is highly sensitive to soil temperature and soil moisture (Nissan et al.
-2022), understanding the responses of soil respiration to these microclimatic conditions, which are affected by global change factors as climate change and forest management, is critical for predicting future atmospheric CO2 concentrations.
+2022), understanding the responses of soil respiration to these microclimatic conditions, which are affected by global change factors as climate change and forest disturbances, is critical for predicting future atmospheric CO2 concentrations.
 Temperature affects respiration rates positively (Lloyd and Taylor, 1994), by increasing the metabolic rates of soil microorganisms and enzyme efficiency (ref).
